### 2018-07-01 paper/main.tex
check in a bare-bones introduction outline for the paper, noting the length limit, submission date, etc.

% 10+2 Due Fri 28 Aug anywhere-on-earth
%
% ICSE 2019 Technical Track submission must not exceed 10 pages,
% including all text, figures, tables, and appendices; two additional pages
% containing only references are permitted.
%
% ICSE 2019 Technical Track will employ a double-blind review process.
% Thus, no submission may reveal its authors’ identities. 
%
% https://2019.icse-conferences.org/track/icse-2019-Technical-Papers#Call-for-Papers

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{textcomp}
\usepackage{xcolor}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

% 
\title{Blaming the Typeless: \\ Scalable, Human-Centric Python Localization}

\iffalse

\author{\IEEEauthorblockN{Benjamin Cosman}
\IEEEauthorblockA{\textit{UC San Digeo}\\
blcosman@eng.ecsd.edu}
\and
\IEEEauthorblockN{Leon Medvinsky}
\IEEEauthorblockA{\textit{UC San Digeo}\\
lmedvinsky@eng.ecsd.edu}
\and
\IEEEauthorblockN{Ranjit Jhala} 
\IEEEauthorblockA{\textit{UC San Digeo}\\
jhala@cs.ecsd.edu}
\and
\IEEEauthorblockN{Westley Weimer} 
\IEEEauthorblockA{\textit{University of Michigan}\\
weimerw@umich.edu}
}

\fi

\author{\IEEEauthorblockN{\emph{submitted for double-blind review}}}

\maketitle

\begin{abstract}
Abstract 
\end{abstract}

\section{Introduction}

\emph{Placeholder Outline.} This text is an informal placeholder outline
only.

\emph{Identify an important problem.} Support the claim that Python fault
localization for possibly-ill-typed student fragments is an important
problem. Likely steps: fault localization is important, Python is
important, incomplete code is important. First, software engineering is
important and expensive; testing, maintenance, and debugging are the
dominant activities in SE; localization and triage are key steps. Second, 
Python is increasingly used for pedagogy and industrial deployment, here
are some GitHub numbers about the rate of growth in Python projects
compared to C or Java, etc. Third, IDE tool use requires operating on code
fragments, here are some numbers from PythonTutor~\cite{Guo2013-vu}, generality, etc.

\emph{Identify the properties of a good solution.} Pick your favorite
three. Some candidates: accuracy (cite Parnin/Orso paper an indicate that
it must be found within top three, etc., since humans will not read long
lists), generality (must apply to student code and incomplete fragments), 
scalability (must apply to hundreds of thousands of fragments, must present
real-time responses suitable for web/IDE use), agreement with subjective
judgments (must handle ``ambiguity'' the same way humans would when you
don't know whether to blame the use or the definition, etc.), efficiency
(must operate using a small number of dynamic runs, unlike Tarantula,
etc.).

\emph{Show that the current state of the art is inadequate.} For each $x$ of
\{manual debugging, Python's default error messages,
Mycroft/Sherrloc/Tarantula/whatever\}, show that $x$ lacks at least one of the
desired properties listed above.

\emph{Insights and special sauce.} What are the two or three good ideas
that we bring together to solve this problem? Some candidates: the use of
modern machine learning techniques (deep learning, etc.); the use of
tree-based contextual features to hit the sweet spot between syntax and semantics;
special handling of unsatisfied use-def type constraints (TBA); etc.

\emph{What is our proposal?} We propose XYZ, an algorithm to do ABC. It
takes as input DEF. It has stages P, Q and R. It produces as output GHI. It
provides guarantees JKL.

\emph{How will we evaluate it?} For each of the properties of a good
solution listed above, indicate how we will evaluate our contribution. What
evidence will we provide? Typically, for each property $x$, we support our
claim to $x$ with either a formal proof (e.g., our algorithm has $x$ by
construction), an empirical evaluation on software artifacts, or an
empirical evaluation on human subjects. In each case, indicate both the
evaluation metrics (e.g., if we care about efficiency, is it measured in
lines-per-second? if we care about precision, what is our non-standard
top-three metric) and success criteria (e.g., how will we define success?
what is the baseline we want to beat). For empirical evaluations, briefly
and implicitly argue that the results are likely to generalize by giving
the size of the sampled set.

\emph{Introduction wrap-up.} The contributions of this paper are as follows
...

\section{Motivating Example}

\section{Overview}

\section{Approach}

\section{Evaluation}

\subsection{Methodology}

\section{Related Work}

\section{Conclusion}

\bibliographystyle{abbrv} 
\bibliography{nanomaly,sw}

\end{document}

### 2018-07-12 src/NanoML/Classify.hs
feature generation

@@ -24,49 +24,6 @@ data Literal
   | LS [String]
   deriving (Show, Eq)
 
--- foldExpr :: (Monoid a, Show b) => (Expr b -> a -> a) -> a -> Expr b -> a
--- foldExpr f z = go
---   where
---   go e = f e $ case e of
---     Var {} -> z
---     Int {} -> z
---     LongInt {} -> z
---     Float {} -> z
---     Imaginary {} -> z
---     Bool {} -> z
---     None {} -> z
---     Ellipsis _ -> z
---     ByteStrings {} -> z
---     Strings {} -> z
---     UnicodeStrings {} -> z
---     Call e args _ -> mconcat $ map go (e:(argExpr <$> args))
---     Subscript x y _ -> mappend (go x) (go y)
---     SlicedExpr e ss _ -> go e <> mconcat (map go $ concat (foo <$> ss))
---     CondExpr t b f _ -> go b <> go t <> go f
---     BinaryOp _ x y _ -> mappend (go x) (go y)
---     UnaryOp _ x _ -> go x
---     -- Dot
---     Lambda _ e _ -> go e
---     Tuple es _ -> mconcat (map go es)
---     -- Yield
---     -- Generator
---     -- ListComp
---     List es _ -> mconcat (map go es)
---     -- Dictionary
---     -- DictComp
---     Set es _ -> mconcat (map go es)
---     -- SetComp
---     -- Starred
---     Paren e _ -> go e
---     StringConversion e _ -> go e
---     e -> error $ "unhandled case of foldExpr: " ++ (show e)
-
--- foo :: Slice a -> [Expr a]
--- foo (SliceProper l u Nothing _) = catMaybes [l, u]
--- foo (SliceProper l u (Just s) _) = catMaybes [l, u, s]
--- foo (SliceExpr e _) = [e]
--- foo (SliceEllipsis _) = []
-
 diff :: ExprSpan -> ExprSpan -> Set SrcSpan
 diff e1 e2 = case (e1, e2) of
   (Lambda {}, _)
@@ -80,16 +37,16 @@ diff e1 e2 = case (e1, e2) of
   (Var x _, Var y _)
     | void x == void y
       -> mempty
-  (Int i1 s1 _, Int i2 s2 _) --TODO could the string versions matter?
+  (Int i1 _ _, Int i2 _ _)
     | i1 == i2
       -> mempty
[…]

### 2018-07-13 paper/main.tex
add pipeline outline

@@
 
+\emph{(Data pipeline summary)} Our raw data consist of every Python 3 program
+that a user executed on PythonTutor.com (not in "live" mode) during 2017, which
+comes to around 5 million programs. Each program which crashes is paired with
+the next program (by the same user) that does not crash. We discard pairs where the
+difference between crashing and fixed versions is too high (more than a stddev
+above average).
+
+For each statement and expression in each
+program, we compute features including the following:
+
+\begin{itemize}
+
+\item \emph{Static / syntactic features}
+\begin{enumerate}
+    \item What kind of statement/expression it is, e.g. assignment / return
+     / import (for statements) or variable / literal / application (expressions).
+    \item Size (the number of AST nodes in the subtree rooted at this node)
+\end{enumerate}
+
+\item \emph{Dynamic features}
+We run each program through an instrumented interpreter (i.e. a modified PythonTutor
+backend, which itself is based on the BDB debugging library). This lets us
+compute the following features:
+\begin{enumerate}
+    \item What type(s) does the expression have at runtime?
+    \item Is it part of the error slice?
+\end{enumerate}
+(Since this interpreter works at the granularity of lines and we want to work with
+expressions as well, we first convert each program to ANF.)
+
+We compute the error slice by running the program and building a dependency
+graph where node A depends on node B if either B defines a variable that A uses,
+or B affects control flow allowing A to run.
+
+We also check what exception is thrown (TypeError, KeyError, etc), and add that
+as a feature to all vectors derived from that program.
+
+\item \emph{Contextual features} After all other features are computed so we have a
+vector $v_e$ representing expression $e$, we set $v_e^{+context} = v_e \circ
+v_p \circ v_{c_1} \circ v_{c_2} \circ v_{c_3}$ where $p$ is the expression's parent
+in the AST and $c_i$ are its children. This allows us to recover some of the
+program structure that would otherwise be completely lost when we convert the
+structured program into an unstructured bag of feature vectors.
+
+\item \emph{Label} Each feature vector is labeled by whether that expression/statement
+changed between the crashing and fixed versions. (This is our proxy for whether
+that node is to blame for the crash).
+
+\end{itemize}
+
+Each vector is then one-hot encoded (e.g. feature "Type : Int" becomes set
+of features "Type-Int : 1", "Type-Bool : 0", "Type-Str : 0", etc.) for use with
+standard machine learning tools. The models we train (inc. decision trees/forests,
+MLP, GBM) can then rank expressions of a test program in order of how likely they
+are to be the source of the bug, so we can compute Top-k accuracies: how often
+is the true program change located in the model's first k outputs. We report Top-1
+and Top-3...
+
 \section{Approach}

### 2018-07-23 paper/main.tex
paper outline

@@
 \emph{(Data pipeline summary)} Our raw data consist of every Python 3 program
-that a user executed on PythonTutor.com (not in "live" mode) during 2017, which
-comes to around 5 million programs. Each program which crashes is paired with
-the next program (by the same user) that does not crash. We discard pairs where the
+that a user executed on PythonTutor.com (not in "live" mode) during 2017, other
+than those with syntax errors or undefined variables. Each program which crashes
+(throws an uncaught Python exception) is paired with
+the next program (by the same user) that does not crash, under the assumption
+that the latter is the fixed version of the former. We discard pairs where the
 difference between crashing and fixed versions is too high (more than a stddev
-above average).
+above average), since these are most likely to be violations of that
+assumption (i.e. the program that does not crash is unrelated to the crashing
+program).
 
@@
 \begin{enumerate}
-    \item What type(s) does the expression have at runtime?
+    \item What type(s) does the expression have at runtime? (The values of this feature
+    can be a type like int, or the special values "unknown" if the expression is
+    never successfully evaluated and "multiple" if it changes type.)
     \item Is it part of the error slice?
+    \item Is it the node at which the program actually crashes?
 \end{enumerate}
@@
 
-We also check what exception is thrown (TypeError, KeyError, etc), and add that
-as a feature to all vectors derived from that program.
+We also check what uncaught exception is thrown by the program (TypeError,
+KeyError, etc), and add that as a constant feature to all vectors derived from that program.
 

### 2018-07-26 pipeline/make_trace/make_trace.py
minor slicer edits

@@ -637,7 +637,7 @@ def slice(source, ri, line=None, debug=False, tr=None, raw=False):
             print('Exception at line ' + str(step_to_line[exception_step]))
     elif not line:
         if raw:
-            return None, [], None, UD_1
+            return None
         else:
             return None, None, None
 

### 2018-07-26 pipeline/make_trace/process_raw.py

@@ -62,10 +62,10 @@ def process_one(string):
     ri = json.dumps(obj['raw_input']) if 'raw_input' in obj else '[]'
 
     (type_info, result) = type_and_slice(source, ri)
-    exceptionSpanANF, slice_spansANF, exceptionMsg, UD_1ANF = result
-    if exceptionSpanANF == None:
-        exceptionSpan, slice_spans, UD_1 = exceptionSpanANF, slice_spansANF, None
+    if result == None:
+        exceptionSpan, slice_spans, exceptionMsg, UD_1 = None, [], None, None
     else:
+        exceptionSpanANF, slice_spansANF, exceptionMsg, UD_1ANF = result
         exceptionSpan, slice_spans, UD_1 = unANF(source, exceptionSpanANF, slice_spansANF, UD_1ANF)
     return (exceptionSpan, type_info, slice_spans, exceptionMsg, UD_1, UD_1ANF)
 

### 2018-07-28 pipeline/make_trace/process_raw.py
fix bug

@@ -63,7 +63,7 @@ def process_one(string):
 
     (type_info, result) = type_and_slice(source, ri)
     if result == None:
-        exceptionSpan, slice_spans, exceptionMsg, UD_1 = None, [], None, None
+        exceptionSpan, slice_spans, exceptionMsg, UD_1, UD_1ANF = None, [], None, None, None
     else:
         exceptionSpanANF, slice_spansANF, exceptionMsg, UD_1ANF = result
         exceptionSpan, slice_spans, UD_1 = unANF(source, exceptionSpanANF, slice_spansANF, UD_1ANF)

### 2018-07-30 paper/main.tex
Some slight restructuring. Next: time to think about the evaluation.

@@
 
-\section{Overview}
+\section{Overview and Approach}
 
-\emph{(Data pipeline summary)} Our raw data consist of every Python 3 program
-that a user executed on PythonTutor.com (not in "live" mode) during 2017, other
-than those with syntax errors or undefined variables. Each program which crashes
-(throws an uncaught Python exception) is paired with
-the next program (by the same user) that does not crash, under the assumption
-that the latter is the fixed version of the former. We discard pairs where the
-difference between crashing and fixed versions is too high (more than a stddev
-above average), since these are most likely to be violations of that
-assumption (i.e. the program that does not crash is unrelated to the crashing
-program).
+We present an algorithm for accurately localizing faults~\cite{tarantula} in
+dynamically-typed Python programs that exhibit non-trivial uncaught runtime
+exceptions. We do not consider syntax errors or references to undefined
+variables. Our algorithm uses machine learning models based on static and
+dynamic features to implicate suspicious expressions. Since studies have
+found that voluminous fault localization output is not useful to
+developers~\cite{orso-parnin,orso-parnin2013}, we focus on producing
+Top-1 and Top-3 rankings.
 
-For each statement and expression in each
-program, we compute features including the following:
+Our algorithm first extracts static and dynamic features from Python
+program (Section~\ref{sec-features}). Next, using a labeled training
+corpus, we learn a machine learning model over those features
+(Section~\ref{sec-model}). Once the model has been learned, we localize
+faults in new Python programs by extracting their features and applying the
+model.
+
+\subsection{Feature Extraction}
+\label{sec-features}
+
+Our key insight is that FIXME-INSIGHT. Are a result, we make use of
+static, dynamic and contextual information. Static features, such as FIXME,
+are effective at capturing FIXME-INFORMATION. By contrast, dynamic
+features, such as FIXME, support the modeling of FIXME. Finally, we also
+include contextual information, such as FIXME, which allows our algorithm
+to accurately handle FIXME. We consider the empirical justification of
+these features in Section~\ref{sec-eval}.
+
+We calculate features for each statement and expression in the program.
 
@@
 of features "Type-Int : 1", "Type-Bool : 0", "Type-Str : 0", etc.) for use with
-standard machine learning tools. The models we train (inc. decision trees/forests,
+standard machine learning tools.
+
+\subsection{Machine Learning and Modeling}
+\label{sec-model}
+
+The models we train (inc. decision trees/forests,
 MLP, GBM) can then rank expressions of a test program in order of how likely they
 are to be the source of the bug, so we can compute Top-k accuracies: how often
-is the true program change located in the model's first k outputs. We report Top-1
-and Top-3...
-
-\section{Approach}
+is the true program change located in the model's first k outputs.
 
 \section{Evaluation}
+\label{sec-eval}
+
+We conducted a large-scale empirical evaluation of our algorithm with the
+aim of addressing a number of research questions: 
+\begin{enumerate}
+
+\item[RQ1]{Can we accurately localize non-trivial faults in Python
+programs?}
+
+\item[RQ2]{Which features are the most important for Python fault
+localization?}
+
+\item[RQ3]{FIXME}
+
+\end{enumerate}
+
+\subsection{Data Set}
+
+Our raw data consist of every Python 3 program that a user executed on
+PythonTutor.com~\cite{Guo2013-vu} (not in ``live'' mode) during 2017, other
+than those with syntax errors or undefined variables.  Each program which
+crashes (throws an uncaught Python exception) is paired with the next
+program (by the same user) that does not crash, under the assumption that
+the latter is the fixed version of the former. We discard pairs where the
+difference between crashing and fixed versions is too high (more than a
+stddev above average), since these are most likely to be violations of that
+assumption (i.e., the program that does not crash is unrelated to the
+crashing program).
+
+In our evaluation, we used FIXME as a training set and FIXME as a held-out
+evaluation set. We employed cross-validation~\cite{kohavi} to help address
+the potential threat of overfitting.
 
 \subsection{Methodology}
+
+For these evaluations, we employ machine learning algorithm FIXME, which
+provided the best balance of FIXME and FIXME (time? accuracy?) on our
+training set.
+
+We report Top-1 and Top-3...
+
+We define FIXME-ACCURACY to be FIXME.
+
+\subsection{RQ 1 --- Fault Localization Accuracy}
+\subsection{RQ 2 --- Feature Predictive Power}
+\subsection{RQ 3 --- FIXME}
 

### 2018-07-30 paper/main.tex
think about plots and research questions, check in some placeholder related work on fault localization

@@
 found that voluminous fault localization output is not useful to
-developers~\cite{orso-parnin,orso-parnin2013}, we focus on producing
+developers~\cite{orso-parnin,orso-parnin2015}, we focus on producing
 Top-1 and Top-3 rankings.
@@
 \item[RQ2]{Which features are the most important for Python fault
-localization?}
+localization overall?}
 
-\item[RQ3]{FIXME}
+\item[RQ3]{Which features are the most important for various categories of
+Python defects?}
+
+\item[RQ4]{FIXME predicates?}
+
+\item[RQ5]{FIXME clustering?}
 
@@
 \subsection{RQ 1 --- Fault Localization Accuracy}
+
+prereq: define accuracy
+
+PLOT Type: Bar Graph (with multiple bars)
+
+PLOT Y Axis: Model Accuracy
+
+PLOT X Axis: Top-1, Top-2, Top3
+
+PLOT Bars: 1 for our algorithm, 1 for a weak baseline (such as the vanilla
+Python interpreter)
+
+Goal: Establish that our algorithm works. This is the most basic question,
+but also the most relevant. If the reader takes away nothing else, they
+should conclude that our algorithm is effective.
+
 \subsection{RQ 2 --- Feature Predictive Power}
-\subsection{RQ 3 --- FIXME}
+
+prereq: define predictive power (leave-one-out, leave-one-in, Relief-F,
+ANOVA, whatever)
+
+PLOT Type: Table
+
+TABLE COLUMNS: Feature Name, Feature Category, Feature Predictive Power
+
+TABLE ROWS: top twelve features
+
+Use as a rough guide: Table 3 on Page 8 of 
+https://web.eecs.umich.edu/~weimerw/p/weimer-icsm2010.pdf
+
+Goal: Establish that we are smart for including all of these features. Note
+which individual features or feature categories were (not) included in
+previous work.
+
+Give a simple narrative about our effectiveness: implicitly, we were smart
+for deciding to include these features, better ingredients make better
+results, we have great results.
+
+\subsection{RQ 3 --- Defect Categories}
+
+prereq: define defect category. This could be either the raw python
+exception name or it could be clusters that we have manually created.
+
+PLOT Type: Bar Graph (three bars per point on the X axis)
+
+PLOT Y Axis: Leave-One-In Accuracy (normalized per X axis point)
+
+PLOT X AXis: Defect Categories
+
+PLOT Bars: Static, Dynamic, Context
+
+Goal: Demonstrate that each category of feature is essential for a certain
+piece of the problem. We need ABC feature to handle DEF class of defects,
+but ABC feature does not work well on GHI class of defects, for those we
+need JKL features. Implicit: we were smart for including all of these
+categories.
+
+\subsection{Qualitative Analysis}
+
+Pick out one example where we did poorly. Walk through why it is ``out of
+scope'', in some sense: we would need features that capture XYZ to handle
+it correctly, but those are either expensive or undecideable. Hint at
+future work.
+
+Pick out a few examples where we did well and the baseline did poorly. Walk
+through how interesting and indicative they are and explain why we did well.
+
+\subsection{Evaluation Summary}
+
+Summarize the evaluation. One sentence per subsection.
 
 \section{Related Work}
+FIXME: This must all be rewritten or dropped, as it is directly copied from
+my previous papers. It serves as a placeholder to indicate length and to
+remind us of some Software Engineering bits we might cite.
+
+There are several fault localization techniques that rely much more heavily
+on dynamic information\footnote{Note: we do not directly compare to these
+state-of-the-art fault localization techniques for several reasons.  First,
+the benchmarks used in this paper do not always have available the test
+suites needed by dynamic techniques and, conversely, those used in the
+previous work do not have the publicly available bug reports required
+by our technique.  Additionally, we return an answer only when confident in
+our result set and also employ a top list of all available results, while
+previous work is typically evaluated using a score metric.  We consider the
+techniques complimentary.}~\cite
+{harrold05,Renieris03,cleve05,wang09}.
+In general,
+these techniques leverage differences between passing and failing program
+executions.  While effective, this type of approach requires program traces
+for not only the specific fault at hand but also from a comprehensive regression
+test suite for comparison.
+A hybrid approach that considers both dynamic
+traces and static natural language similarity was recently proposed by
+Medini \textit{et al.}~\cite{Medini11}.
+In contrast to these approaches, our technique relies purely on static
+information that is readily available in most commercial systems with minimal
+additional developer effort.
+
+Prabhakararao and Ruthruff \emph{et al.} performed two human studies
+to gauge the effectiveness of an interactive fault localization tool
+developed for end users with little to no
+experience~\cite{Prabhakararao03,ruthruff05}.  The goal of their
+studies was to evaluate the use of feedback when locating faults and
+to generally study the process of fault localization, especially by
+users with no expert domain knowledge of the source.  By comparison,
+our human study also examines the fault localization process but for
+the purpose of evaluating software quality metrics.  We are less
+interested in the specific process and more concerned with the
+resulting accuracy and the human intuitions about the code in
+question.  Additionally, our human study is of a much broader scope and
+thus we hope it is more generalizable.
+%Additionally, their studies included 10 and 20
+%participants respectively using a spreadsheet-based visualization tool.
+%Our study of 61 humans looking at actual source code potentially provides
+%more generalizable results and targets specifically the area of software
+%maintenance completed by trained programmers.
+
+Ashok \textit{et al.} propose a similar natural language search technique
+in which users can match an incoming report to previous reports,
+programmers and source code~\cite{Ashok09}.  By comparison, our technique
+is more lightweight and focuses only on searching the code and the
+defect report.
+
+Jones \textit{et al.} developed Tarantula, a technique that performs
+fault localization based on the insight that statements executed often
+during failed test cases likely account for potential fault
+locations~\cite{harrold05}. Their approach is quite effective when
+a rich, indicative test suite is available and can be run as part of
+the fault localization process. It thus requires the fault-inducing
+input but not any natural language defect report. By contrast,
+our approach is lightweight, does not require an indicative test
+suite or fault-inducing input, but does require a natural language
+defect report. Both approaches will yield comparable performance, and
+could even be used in tandem.
+
+Cleve and Zeller localize faults by finding differences between
+correct and failing program execution states, limiting the scope of
+their search to only variables and values of interest to the fault in
+question~\cite{cleve05}. Notably, they focus on those variable and
+values that are relevant to the failure and to those program execution
+points where transitions occur and those variables become causes of
+failure. Their approach is in a strong sense finer-grained than ours:
+while nothing prevents our technique from being applied at the level
+of methods instead of files, their technique can give very precise
+information such as ``the transition to failure happened when $x$
+became 2.'' Our approach is lighter-weight and does not require
+that the program be run, but it does require defect reports.
+
+Renieris and Rice use a ``nearest neighbor'' technique in their
+Whither tool to identify faults based on exposing differences in
+faulty and non-faulty runs that take very similar executions
+paths~\cite{Renieris03}. They assume a large number of correct runs
+(e.g., normal test cases) and one failing run. Their approach uses a
+distance criterion to select the correct run that is closest to the
+failing run and produces a report of ``suspicious'' parts of the
+program. By comparison, we chose to limit the programmatic information
+used by our technique to only that which was reported by users: we 
+do not use test case runs but do need natural language.
+
+Liblit \textit{et al.} use Cooperative Bug Isolation, a statistical
+approach to isolate multiple bugs within a program given a deployed user
+base. By analyzing large amounts of collected execution data from real
+users, they can successfully differentiate between different causes of
+faults in failing software~\cite{liblit05}. Their technique produces
+a ranked list of very specific fault localizations (e.g., ``the fault
+occurs when $i > arrayLen$ on line 57''). In general, their technique
+can produce more precise results than ours, but it requires a set of
+deployed users and works best on those bugs experienced by many users.
+By contrast, we do not require that the program be runnable, much less
+deployed, and use only natural language defect report text.
 

### 2018-07-30 cluster/cluster.py
Clustering on python source strings, based on edit distance

@@ -1,0 +1,40 @@
+import ast
+import scipy.cluster.hierarchy
+from scipy.spatial.distance import pdist
+import zss
+
+from sklearn.cluster import AgglomerativeClustering
+
+def ast_children(node):
+    return list(ast.iter_child_nodes(node))
+
+def ast_label(node):
+    return type(node)
+
+def label_distance(a, b):
+    if a == b:
+        return 0
+    else:
+        return 1
+
+def ast_distance(a, b):
+    return zss.simple_distance(a, b, ast_children, ast_label, label_distance)
+
+# Takes a 1 element list containing the ast to compare
+def metric(a, b):
+    return ast_distance(a[0], b[0])
+
+def cluster(asts, t):
+    distance_matrix = pdist(list(map(lambda x: [x], asts)), metric)
+    linkage = scipy.cluster.hierarchy.linkage(distance_matrix, method='single')
+    return scipy.cluster.hierarchy.fcluster(linkage, t, criterion='distance')
+
+# Takes a list of python source strings and produces an array where element i is
+# the cluster number of the string with index i in the iterable that was passed
+# in.
+#
+# strings: Iterable of python source strings
+# t: Max edit distance for elements in the same cluster
+def cluster_strings(strings, t):
+    asts = map(ast.parse, strings)
+    return cluster(asts, t)

### 2018-07-30 cluster/cluster.py
remove scikit depend

@@ -3,8 +3,6 @@ import scipy.cluster.hierarchy
 from scipy.spatial.distance import pdist
 import zss
 
-from sklearn.cluster import AgglomerativeClustering
-
 def ast_children(node):
     return list(ast.iter_child_nodes(node))
 

### 2018-07-30 cluster/cluster.py
Cluster on pairs data

@@ -22,10 +22,14 @@ def ast_distance(a, b):
 def metric(a, b):
     return ast_distance(a[0], b[0])
 
-def cluster(asts, t):
-    distance_matrix = pdist(list(map(lambda x: [x], asts)), metric)
-    linkage = scipy.cluster.hierarchy.linkage(distance_matrix, method='single')
-    return scipy.cluster.hierarchy.fcluster(linkage, t, criterion='distance')
+def distance_matrix(asts):
+    return pdist(list(map(lambda x: [x], asts)), metric)
+
+def linkage(d_matrix):
+    return scipy.cluster.hierarchy.linkage(d_matrix, method='single')
+
+def cluster(link, t):
+    return scipy.cluster.hierarchy.fcluster(link, t, criterion='distance')
 
 # Takes a list of python source strings and produces an array where element i is
 # the cluster number of the string with index i in the iterable that was passed
@@ -34,5 +38,14 @@ def cluster(asts, t):
 # strings: Iterable of python source strings
 # t: Max edit distance for elements in the same cluster
 def cluster_strings(strings, t):
+    print('Parsing strings...')
     asts = map(ast.parse, strings)
-    return cluster(asts, t)
+
+    print('Computing distance matrix...')
+    dmat = distance_matrix(asts)
+
+    print('Computing linkage...')
+    link = linkage(dmat)
+
+    print('Computing clusters...')
+    return cluster(link, t)

### 2018-07-30 cluster/process.py

@@ -1,0 +1,27 @@
+import sys
+import json
+import glob
+import cluster
+
+decoder = json.JSONDecoder()
+
+paths = glob.glob(sys.argv[1])
+
+strings = []
+
+print('Building list of programs...')
+for path in paths:
+    with open(path) as file:
+        # Get the last element
+        last = next(file, "")
+        try:
+            obj = decoder.decode(last)
+            source = obj['fix']
+            strings.append(source)
+        except Exception as e:
+            print(e)    
+
+print('Clustering...')
+clustering = cluster.cluster_strings(strings, int(sys.argv[2]))
+
+print('Done clustering. Cluster count is %d' % max(clustering))

### 2018-07-30 paper/main.tex
check in our notes from the call

@@
 
-\item[RQ4]{FIXME predicates?}
+\item[RQ4]{Is our algorithm accurate on heterogeneous sets of programs?}
 
@@
 
+Ranjit: ``our algorithms'' --- ``purely syntactic vs. with types'' 
+--- ``with or without slicing'' as extra ``bars''
+
 \subsection{RQ 2 --- Feature Predictive Power}
@@
 
+TODO: what about turning on/off slicing?
+
+TODO: what about turning on/off types? (is that separate from slicing?)
+
+TODO: do ANOVA or ReliefF
+
 \subsection{RQ 3 --- Defect Categories}
@@
 exception name or it could be clusters that we have manually created. 
+Ranjit notes: result could be ``we need everything for everything'', at
+which point this is useless and should be skipped. Per Ben's exam: may find
+that for different categories of errors we should use different ML
+classifiers.
 
@@
 PLOT Y Axis: Leave-One-In Accuracy (normalized per X axis point)
+Ranjit notes: on smaller subsets we could rebuild the classifier each time.
+Wes notes that if we do, it takes us longer. If we re-use the monolithic
+model we can do this very quickly.
 
@@
 categories.
+
+\subsection{RQ 4 --- Diversity of Programs}
+
+In Nate there were 20 different functions. All the programs were doing the
+same 20 functions.
+
+Here, people are doing anything they want. Random stuff going on over here.
+
+We want to show that we get high accuracy *over a heterogenous set of
+programs*. Leon?
+
+prereq: cluster the programs based on similarity. Using agglomerative
+clustering plus python edit distance. How good the performance will be?
+right now it creates a clustering tree. binary tree, based on min distance. 
+can give it a threshhold for distance (elements w/in a group). don't forget
+to rename all variables to X, perhaps normalize (edit distance size).
+
+leon: could also cluster by inconsistency coefficient
+
+PLOT Type: line graph (standard plot)
+
+PLOT X AXIS: clustering distance parameter
+
+PLOT Y AXIS: number of different clusters found
+
+PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''
+
+Goal: show that ``no matter how you decide program differences'', our
+dataset is always more diverse/heterogeneous than is previous work. So our
+success is amazing.
+
+\subsection{RQ 5 --- Wishlist: Partitioning Kinds of Errors}
+
+Can we partition the kinds of errors that people make? What are common
+misconceptions? Do these correspond to paths of decision trees.
+
+If we don't get to the source-sink thing, it could go here.
+
+I don't think we have time, in this paper, to actually implement a
+two-level machine learning.
 

### 2018-07-31 pipeline/3-make-pairs.py
update data pipeline

@@ -22,29 +22,29 @@ os.makedirs(OUT_FOLDER, exist_ok=True)
 for fileName in os.listdir(DATA_FOLDER):
     try:
         fullPath = os.path.join(DATA_FOLDER,fileName)
-        if os.path.isdir(fullPath):
-            continue
-        pairs = []
         with open(fullPath, 'r') as logFile:
-            badProgs = []
-            for line in logFile:
-                dct = json.loads(line)
-                if "PF_exitPipelineReason" in dct:
-                    # if dct["PF_exitPipelineReason"] != 'Does not parse':
-                        # slicer or ANFer crashed or timed out
-                        # TODO Since we can't be sure if this program was good or bad,
-                        # throw out unpaired bad programs before this point
-                        # badProgs = []
-                    continue
-                newProg = dct['user_script']
-                tup = findTuple(dct['PF_slicerOutput'])
-                (result, types, slice, msg, UD_1, _) = tup
-                if msg != None:
-                    msg = msg.split(":")[0]
-                # debugDict[msg] += 1
-                if result == None:
-                    # found a good program!
-                    for (badProg, badSlice, badTypes, badExceptionSpan, badMsg, badUD) in badProgs:
+            lines = [json.loads(line) for line in logFile]
+            lines.sort(key=lambda x: x['serverTimeUTC'])
+        pairs = []
+        badProgs = []
+        for dct in lines:
+            if "PF_exitPipelineReason" in dct:
+                # if dct["PF_exitPipelineReason"] != 'Does not parse':
+                    # slicer or ANFer crashed or timed out
+                    # TODO Since we can't be sure if this program was good or bad,
+                    # throw out unpaired bad programs before this point
+                    # badProgs = []
+                continue
+            newProg = dct['user_script']
+            tup = findTuple(dct['PF_slicerOutput'])
+            (result, types, slice, msg, UD_1, _) = tup
+            if msg != None:
+                msg = msg.split(":")[0]
+            # debugDict[msg] += 1
+            if result == None:
+                # found a good program!
+                if len(badProgs) > 0:
+                    for ((badProg, badSlice, badTypes, badExceptionSpan, badMsg, badUD), isConsecutive) in zip(badProgs, [False]*(len(badProgs)-1)+[True]):
                         varTypes = {}
                         spanTypes = {}
                         for key in badTypes:
@@ -70,14 +70,17 @@ for fileName in os.listdir(DATA_FOLDER):
                                     "varTypes": varTypes,
                                     "spanTypes": spanTypes,
[…]

### 2018-07-31 pipeline/groupAndSort.py

@@ -1,22 +1,0 @@
-from collections import defaultdict
-import os.path, os
-import sys
-import json
-
-OUT_FOLDER = sys.argv[1]
-os.makedirs(OUT_FOLDER, exist_ok=True)
-
-result = defaultdict(list)
-for line in sys.stdin:
-    dct = json.loads(line)
-    if 'PF_exitPipelineReason' in dct:
-        continue
-    result[dct['session_uuid']].append(dct)
-
-for (uuid, dctlist) in result.items():
-    if len(dctlist) < 2:
-        continue
-    dctlist.sort(key=lambda x: x['serverTimeUTC'])
-    with open(os.path.join(OUT_FOLDER,uuid), 'w') as outFile:
-        for dct in dctlist:
-            outFile.write(json.dumps(dct)+"\n")

### 2018-07-31 pipeline/groupBySession.py

@@ -1,0 +1,13 @@
+import os.path, os
+import sys
+import json
+
+OUT_FOLDER = sys.argv[1]
+os.makedirs(OUT_FOLDER, exist_ok=True)
+
+for line in sys.stdin:
+    dct = json.loads(line)
+    if 'PF_exitPipelineReason' in dct:
+        continue
+    with open(os.path.join(OUT_FOLDER,dct['session_uuid']), 'a') as outFile:
+        outFile.write(json.dumps(dct)+"\n")

### 2018-07-31 pipeline/run-all.sh

@@ -48,7 +48,7 @@ foo "Docker build" "docker build -t python-pipeline ."
 
 foo "Run slicer" "gunzip -c $outDir/py3-web_exec.jsonl.gz | docker run -i python-pipeline /bin/bash -c 'python3 slicePass.py $rerun' | gzip > $outDir/sliced.jsonl.gz"
 
-foo "Group by session" "gunzip -c $outDir/sliced.jsonl.gz | python3 groupAndSort.py $outDir/sessions"
+foo "Group by session" "gunzip -c $outDir/sliced.jsonl.gz | python3 groupBySession.py $outDir/sessions"
 
 foo "Make pairs" "python3 3-make-pairs.py $outDir/sessions $outDir/pairs"
 

### 2018-08-01 bin/Generate.hs
data pipeline

@@ -65,6 +65,7 @@ main = do
   let features = if size then fsNormalCtx ++ [fSizeCtx] else fsNormalCtx
   let features' = if context then (requestTypeMap features fTypeCtx) else (requestTypeMap (noCtx <$> features) (noCtx . fTypeCtx))
   let features'' a = if oneHot then concatMap toOneHot (features' a) else (features' a)
+
   let useSlice = if slice then All else JustSlice
 
   let nm = "blah" ++ (if oneHot then "+oneHot" else "")
@@ -72,7 +73,7 @@ main = do
               ++ (if slice then "+slice" else "")
               ++ (if size then "+size" else "")
 
-  mkBadFeaturesWithSlice useSlice out nm features' jsons
+  mkBadFeaturesWithSlice useSlice out nm features'' jsons
 
 requestTypeMap :: [Feature a] -> (TypeMap -> Feature a) -> TypeMap -> [Feature a]
 requestTypeMap fs tf tm = fs ++ [tf tm]
@@ -100,7 +101,7 @@ mkBadFeaturesWithSlice withSlice out nm fs jsons = do
         -- putStrLn fix
         return ()
       | otherwise -> do
-        let fn = printf "%04d" (i :: Int)
+        let fn = printf "%05d" (i :: Int)
         let path = out </> nm </> fn <.> "csv"
         createDirectoryIfMissing True (takeDirectory path)
         LBSC.writeFile path $ encodeByName header features

### 2018-08-01 pipeline/subsets.py

@@ -1,0 +1,20 @@
+import os, sys, json
+
+dataDir = sys.argv[1]
+outDir = os.path.join(dataDir, 'categories')
+os.mkdir(outDir)
+finalDir = os.path.join(outDir, 'final')
+os.mkdir(finalDir)
+consecutiveDir = os.path.join(outDir, 'consecutive')
+os.mkdir(consecutiveDir)
+
+pairsFile = os.path.join(dataDir, 'goodPairs.jsonl')
+with open(pairsFile, 'r') as pairsList:
+    for line in pairsList:
+        pair = json.loads(line)
+        csvFilename = "%04d" % pair['index']+'.csv'
+        src = os.path.join(dataDir, 'foo', 'blah+context+slice+size', csvFilename)
+        if pair['isFinal']:
+            os.symlink(src, os.path.join(finalDir, csvFilename))
+        if pair['isConsecutive']:
+            os.symlink(src, os.path.join(consecutiveDir, csvFilename))

### 2018-08-01 pipeline/subsets.py
data pipeline

@@ -1,20 +1,18 @@
 import os, sys, json
 
 dataDir = sys.argv[1]
-outDir = os.path.join(dataDir, 'categories')
-os.mkdir(outDir)
-finalDir = os.path.join(outDir, 'final')
-os.mkdir(finalDir)
-consecutiveDir = os.path.join(outDir, 'consecutive')
-os.mkdir(consecutiveDir)
 
+outDir = os.path.join(dataDir, 'subsets')
+os.mkdir(outDir)
+
+finalsFilename = os.path.join(outDir, 'finals.txt')
+consecutivesFilename = os.path.join(outDir, 'consecutives.txt')
 pairsFile = os.path.join(dataDir, 'goodPairs.jsonl')
-with open(pairsFile, 'r') as pairsList:
+with open(pairsFile, 'r') as pairsList, open(finalsFilename, 'w') as finalsFile, open(consecutivesFilename, 'w') as consecutivesFile:
     for line in pairsList:
         pair = json.loads(line)
         csvFilename = "%04d" % pair['index']+'.csv'
-        src = os.path.join(dataDir, 'foo', 'blah+context+slice+size', csvFilename)
         if pair['isFinal']:
-            os.symlink(src, os.path.join(finalDir, csvFilename))
+            finalsFile.write(csvFilename+'\n')
         if pair['isConsecutive']:
-            os.symlink(src, os.path.join(consecutiveDir, csvFilename))
+            consecutivesFile.write(csvFilename+'\n')

### 2018-08-06 paper/main.tex
motivating example

@@
 \section{Motivating Example}
+
+\begin{figure}
+\begin{verbatim}
+year = int(time.strftime("%Y"))
+age = input("Enter your age")
+print("You will be twice as old in:")
+print(year + age)
+\end{verbatim}
+\caption{A buggy program}
+\end{figure}
+
+Consider the program in Fig. 1, adapted from our dataset. The program
+attempts to use the current year and the user's age to compute their
+birth year and the year they can vote. However, while line 1 correctly
+turns the current year from a string into an int, this `int' call is
+missing from line 2 so `age' remains a string, and thus the addition
+on line 4 causes the program to crash.
+
+Notice that while a correct fix (and the one that the programmer
+actually used) is to add an `int' call to line 2, another possible ``fix''
+would be to \textit{remove} the `int' call from line 1 - both year and
+age would then be strings, and the addition on line 4 would be interpreted
+as string concatenation. This shows the necessity of a heuristic, data-driven
+approach, since without test cases there is no way to tell for sure which
+fix is better.
+
+We can also start to see some features which will be
+important for finding bugs: we will need the \textit{type} of age, as
+well as some \textit{contextual} information about how it's used in the program.
+
+% TODO: would also be nice if the example showed that the place Python crashes
+% may not be a place that should be fixed?
 

### 2018-08-06 src/NanoML/Classify.hs

@@ -343,6 +343,7 @@ data ESKind
   | DotK (Ident ())
   | SliceK
   | SubscriptK
+  | DictionaryK Int
   | NoneK
   | EllipsisK
   | SetK
@@ -417,7 +418,7 @@ exprKind = \case
   -- Generator
   -- ListComp
   List {} -> ListK
-  -- Dictionary
+  Dictionary pairs _ -> DictionaryK (len pairs)
   -- DictComp
   Set {} -> SetK
   -- SetComp

### 2018-08-06 src/NanoML/Classify.hs
data pipeline

@@ -349,6 +349,9 @@ data ESKind
   | SetK
   | StarredK
   | StringConversionK
+  | ListCompK (Comprehension ())
+  | DictCompK (Comprehension ())
+  | SetCompK (Comprehension ())
   deriving (Eq, Show)
 
 --TODO a Suite should be alowed to change size without effectively changing Kind?
@@ -416,16 +419,39 @@ exprKind = \case
   Tuple {} -> TupleK
   -- Yield
   -- Generator
-  -- ListComp
+  ListComp c _ -> ListCompK $ canonicalComprehension c
   List {} -> ListK
-  Dictionary pairs _ -> DictionaryK (len pairs)
-  -- DictComp
+  Dictionary pairs _ -> DictionaryK (length pairs)
+  DictComp c _ -> DictCompK $ canonicalComprehension c
   Set {} -> SetK
-  -- SetComp
+  SetComp c _ -> SetCompK $ canonicalComprehension c
   Starred {} -> StarredK
   Paren {} -> ParenK
   StringConversion {} -> StringConversionK
   e -> error $ "unhandled case of exprKind: " ++ (show e)
+
+defaultExpr = None ()
+
+canonicalComprehension :: Comprehension () -> Comprehension ()
+canonicalComprehension (Comprehension e f _) = Comprehension e' f' ()
+  where
+    e' = case e of
+      ComprehensionExpr _ -> ComprehensionExpr (defaultExpr)
+      ComprehensionDict _ -> ComprehensionDict (DictMappingPair (defaultExpr) (defaultExpr))
+    f' = canonicalCompFor f
+
+canonicalCompFor :: CompFor () -> CompFor ()
+canonicalCompFor (CompFor es _ it _) = CompFor es' defaultExpr it' ()
+  where
+    es' = map (const defaultExpr) es
+    it' = canonicalCompIter <$> it
+
+canonicalCompIter :: CompIter () -> CompIter ()
+canonicalCompIter (IterFor cf _) = IterFor (canonicalCompFor cf) ()
+canonicalCompIter (IterIf ci _) = IterIf (canonicalCompIf ci) ()
+
+canonicalCompIf :: CompIf () -> CompIf ()
+canonicalCompIf (CompIf _ ci _) = CompIf defaultExpr (canonicalCompIter <$> ci) ()
 
 subESes' :: ES -> [ES]
 subESes' = \case

### 2018-08-06 src/NanoML/Classify.hs
data pipeline

@@ -18,10 +18,14 @@ instance Span ES where
 
 data Literal
   = LI Integer
+  | LLI Integer
+  | LIm Double
   | LD Double
   | LB Bool
   | LC Char
   | LS [String]
+  | LBS [String]
+  | LUS [String]
   deriving (Show, Eq)
 
 diff :: ExprSpan -> ExprSpan -> Set SrcSpan
@@ -399,15 +403,15 @@ exprKind :: Expr () -> ESKind
 exprKind = \case
   Var v _ -> VarK v
   Int i _ _ -> LitK (LI i)
-  -- LongInt i _ _ -> LitK (LI i)
+  LongInt i _ _ -> LitK (LLI i)
   Float f _ _ -> LitK (LD f)
-  -- Imaginary {} -> []
+  Imaginary v _ _ -> LitK (LIm v)
   Bool b _ -> LitK (LB b)
   None _ -> NoneK
   Ellipsis _ -> EllipsisK
-  -- ByteStrings {} -> []
+  ByteStrings ss _ -> LitK (LBS ss)
   Strings ss _ -> LitK (LS ss)
-  -- UnicodeStrings {} -> []
+  UnicodeStrings ss _ -> LitK (LUS ss)
   Call {} -> AppK --TODO different kinds of arguments
   Subscript x y _ -> SubscriptK
   SlicedExpr {} -> SliceK --TODO different kinds of Slices

### 2018-08-06 src/NanoML/Classify.hs
data pipeline

@@ -356,6 +356,7 @@ data ESKind
   | ListCompK (Comprehension ())
   | DictCompK (Comprehension ())
   | SetCompK (Comprehension ())
+  | YieldK (Maybe (YieldArg ()))
   deriving (Eq, Show)
 
 --TODO a Suite should be alowed to change size without effectively changing Kind?
@@ -421,14 +422,14 @@ exprKind = \case
   Dot _ a _ -> DotK a
   -- Lambda _ e _ -> [e]
   Tuple {} -> TupleK
-  -- Yield
+  Yield a _ -> YieldK (defaultYieldArg <$> a)
   -- Generator
-  ListComp c _ -> ListCompK $ canonicalComprehension c
+  ListComp c _ -> ListCompK $ defaultComprehension c
   List {} -> ListK
   Dictionary pairs _ -> DictionaryK (length pairs)
-  DictComp c _ -> DictCompK $ canonicalComprehension c
+  DictComp c _ -> DictCompK $ defaultComprehension c
   Set {} -> SetK
-  SetComp c _ -> SetCompK $ canonicalComprehension c
+  SetComp c _ -> SetCompK $ defaultComprehension c
   Starred {} -> StarredK
   Paren {} -> ParenK
   StringConversion {} -> StringConversionK
@@ -436,26 +437,30 @@ exprKind = \case
 
 defaultExpr = None ()
 
-canonicalComprehension :: Comprehension () -> Comprehension ()
-canonicalComprehension (Comprehension e f _) = Comprehension e' f' ()
+defaultComprehension :: Comprehension () -> Comprehension ()
+defaultComprehension (Comprehension e f _) = Comprehension e' f' ()
   where
     e' = case e of
       ComprehensionExpr _ -> ComprehensionExpr (defaultExpr)
       ComprehensionDict _ -> ComprehensionDict (DictMappingPair (defaultExpr) (defaultExpr))
-    f' = canonicalCompFor f
+    f' = defaultCompFor f
 
-canonicalCompFor :: CompFor () -> CompFor ()
-canonicalCompFor (CompFor es _ it _) = CompFor es' defaultExpr it' ()
+defaultCompFor :: CompFor () -> CompFor ()
+defaultCompFor (CompFor es _ it _) = CompFor es' defaultExpr it' ()
   where
     es' = map (const defaultExpr) es
[…]

### 2018-08-06 src/NanoML/Classify.hs
data pipeline

@@ -521,7 +521,9 @@ subExprs = \case
   Dot e _ _ -> [e]
   Lambda _ e _ -> [e]
   Tuple es _ -> es
-  -- Yield
+  Yield Nothing _ -> []
+  Yield (Just (YieldFrom e _)) _ -> [e]
+  Yield (Just (YieldExpr e)) _ -> [e]
   Generator c _ -> comprehensionSubExprs c
   ListComp c _ -> comprehensionSubExprs c
   List es _ -> es
@@ -532,7 +534,6 @@ subExprs = \case
   Starred e _ -> [e]
   Paren e _ -> [e]
   StringConversion e _ -> [e]
-  e -> error $ "unhandled case of subExprs: " ++ (show e)
 
 comprehensionSubExprs (Comprehension ce cf _) = ceSubExprs ce ++ cfSubExprs cf
   where

### 2018-08-07 src/NanoML/Classify.hs
data pipeline

@@ -5,7 +5,7 @@ import Control.Monad
 import Data.Monoid
 import qualified Data.Set as Set
 import Data.Set (Set)
-import Data.Maybe (catMaybes, isJust)
+import Data.Maybe (catMaybes, isJust, isNothing)
 
 import Language.Python.Common hiding ((<>))
 
@@ -30,14 +30,6 @@ data Literal
 
 diff :: ExprSpan -> ExprSpan -> Set SrcSpan
 diff e1 e2 = case (e1, e2) of
-  (Lambda {}, _)
-    -> error "Lambda unhandled in diff"
-  (_, Lambda {})
-    -> error "Lambda unhandled in diff"
-  (Yield {}, _)
-    -> error "Yield unhandled in diff"
-  (_, Yield {})
-    -> error "Yield unhandled in diff"
   (Var x _, Var y _)
     | void x == void y
       -> mempty
@@ -90,11 +82,18 @@ diff e1 e2 = case (e1, e2) of
   (Dot x1 a1 _, Dot x2 a2 _)
     | void a1 == void a2
       -> diff x1 x2
-  -- Lambda _ e _
+  (Lambda a1 e1 _, Lambda a2 e2 _)
+    | (length a1 == length a2) && sameParameterTypes a1 a2
+      -> merge $ (diff e1 e2) : (zipWith diffParameter a1 a2)
   (Tuple es1 _, Tuple es2 _)
     | length es1 == length es2
       -> merge $ zipWith diff es1 es2
-  -- Yield
+  (Yield Nothing _, Yield Nothing _)
+      -> mempty
+  (Yield (Just (YieldFrom e1 _)) _, Yield (Just (YieldFrom e2 _)) _)
+      -> diff e1 e2
+  (Yield (Just (YieldExpr e1)) _, Yield (Just (YieldExpr e2)) _)
+      -> diff e1 e2
   (Generator c1 _, Generator c2 _)
     | Just ss <- diffComprehension c1 c2
       -> ss
@@ -151,10 +150,11 @@ diffDictPair (DictMappingPair x1 y1) (DictMappingPair x2 y2) = mconcat [diff x1 
 sameArgTypes :: [Argument a] -> [Argument a] -> Bool
 sameArgTypes as1 as2 = all sameArgType (zip as1 as2)
   where
-    sameArgType (ArgExpr _ _, ArgExpr _ _) = True
-    sameArgType (ArgVarArgsPos _ _, ArgVarArgsPos _ _) = True
[…]

### 2018-08-08 pipeline/4-run-on-single-lines.py
data pipeline

@@ -15,7 +15,7 @@ with open(OUT_FOLDER+"/failPairs.jsonl", 'w') as failFile, open(OUT_FOLDER+"/goo
                 dct = json.loads(line.strip())
                 with open("temp.json", 'w') as tempFile:
                     tempFile.write(line)
-                cmd = "--source temp.json --out %s" % (OUT_FOLDER+"/temp")
+                cmd = "--source temp.json --context --slice --size --out %s" % (OUT_FOLDER+"/vectors")
                 p = run([GENERATE_FEATURES_BIN]+cmd.split(),
                         stdout=PIPE, stderr=PIPE, encoding="utf-8", errors="strict")
                 # dct['PF_output'] = (p.stdout, p.stderr) # for debugging

### 2018-08-08 pipeline/run-all.sh

@@ -52,9 +52,6 @@ foo "Group by session" "gunzip -c $outDir/sliced.jsonl.gz | python3 groupBySessi
 
 foo "Make pairs" "python3 3-make-pairs.py $outDir/sessions $outDir/pairs"
 
-foo "Filter pairs" "python3 4-run-on-single-lines.py $outDir/pairs $outDir"
+foo "Generate features" "python3 4-run-on-single-lines.py $outDir/pairs $outDir"
 
-# foo "Generate features (oneHot)" "stack exec -- generate-features --source $outDir/goodPairs.jsonl --oneHot --context --slice --size --out $outDir/foo"
-foo "Generate features (categorical)" "stack exec -- generate-features --source $outDir/goodPairs.jsonl --context --slice --size --out $outDir/foo"
-
-foo "Remove outliers" "python3 removeOutliers.py $outDir/foo/blah+context+slice+size $outDir/outliersRemoved"
+foo "Remove outliers" "python3 removeOutliers.py $outDir/vectors/blah+context+slice+size $outDir/outliersRemoved"

### 2018-08-13 cluster/cluster.py
support for the ocaml dataset

@@ -1,35 +1,17 @@
-import ast
 import scipy.cluster.hierarchy
+import numpy as np
+import os
 from scipy.spatial.distance import pdist
-import zss
+import stringify
 
-def ast_children(node):
-    return list(ast.iter_child_nodes(node))
-
-def ast_label(node):
-    return type(node)
-
-def label_distance(a, b):
-    if a == b:
-        return 0
-    else:
-        return 1
-
-def ast_distance(a, b):
-    return zss.simple_distance(a, b, ast_children, ast_label, label_distance)
-
-# Takes a 1 element list containing the ast to compare
-def metric(a, b):
-    return ast_distance(a[0], b[0])
-
-def distance_matrix(asts):
-    return pdist(list(map(lambda x: [x], asts)), metric)
+def distance_matrix(strings):
+    return pdist(list(map(lambda x: [x], strings)), stringify.metric)
 
 def linkage(d_matrix):
     return scipy.cluster.hierarchy.linkage(d_matrix, method='single')
 
 def cluster(link, t):
-    return scipy.cluster.hierarchy.fcluster(link, t, criterion='distance')
+    return scipy.cluster.hierarchy.fcluster(link, t)
 
 # Takes a list of python source strings and produces an array where element i is
 # the cluster number of the string with index i in the iterable that was passed
@@ -37,15 +19,34 @@ def cluster(link, t):
 #
 # strings: Iterable of python source strings
 # t: Max edit distance for elements in the same cluster
-def cluster_strings(strings, t):
-    print('Parsing strings...')
-    asts = map(ast.parse, strings)
-
+def cluster_strings(strings, data_name):
+    # Used to store intermediate things in case a bug causes crashes
+    try:
+        os.mkdir(data_name + '_progress')
+    except:
+        # Assume any error means the directory exists
+        pass
+    
     print('Computing distance matrix...')
-    dmat = distance_matrix(asts)
+    dmat = distance_matrix(strings)
+    dmat.tofile(data_name + '_progress/dmat')
 
     print('Computing linkage...')
     link = linkage(dmat)
+    link.tofile(data_name + '_progress/link')
 
     print('Computing clusters...')
-    return cluster(link, t)
+    cluster_counts = []
+    for i in range(200):
+        t = i / 100
+        cluster_counts.append(cluster(link, t).max())
+    np.array(cluster_counts).tofile(data_name + '_cluster_counts', sep='\n')
+
[…]

### 2018-08-13 cluster/process.py

@@ -1,27 +1,50 @@
+from parse import python_sources_from_glob
 import sys
-import json
-import glob
 import cluster
+import random
+import stringify
+import ast
+import json
+import numpy as np
 
-decoder = json.JSONDecoder()
+def read_python():
+    print('Building list of programs...')
+    strings = python_sources_from_glob(sys.argv[1])
 
-paths = glob.glob(sys.argv[1])
+    print('Transforming...')
+    asts = map(ast.parse, strings)
+    return list(map(stringify.transform_python, asts))
 
-strings = []
+def read_ocaml():
+    decoder = json.JSONDecoder()
+    
+    strings = []
+    print('Building list of transformed programs...')
+    with open('ocaml-json') as f:
+        for line in f:
+            obj = decoder.decode(line)
+            strings.append(stringify.transform_ocaml(obj))
 
-print('Building list of programs...')
-for path in paths:
-    with open(path) as file:
-        # Get the last element
-        last = next(file, "")
-        try:
-            obj = decoder.decode(last)
-            source = obj['fix']
-            strings.append(source)
-        except Exception as e:
-            print(e)    
+    return strings
+
+data_name = ''
+
+if len(sys.argv) == 1:
+    data_name = 'ocaml'
+    transformeds = read_ocaml()
+elif len(sys.argv) == 2:
+    data_name = 'python'
+    transformeds = read_python()
+else:
+    print('0 args for ocaml, 1 arg for python')
+    assert(false)
+
+print('Sampling programs...')
+indices = random.sample(list(range(len(transformeds))), 5000)
+np.array(indices).tofile(data_name + '_indices', sep='\n')
+selection = [transformeds[i] for i in indices]
 
 print('Clustering...')
-clustering = cluster.cluster_strings(strings, int(sys.argv[2]))
+cluster.cluster_strings(selection, data_name)
 
-print('Done clustering. Cluster count is %d' % max(clustering))
+print('Done clustering. Cluster counts output to %s_cluster_counts.' % data_name)

### 2018-08-13 cluster/parse.py
add important files that were missing from commit

@@ -1,0 +1,37 @@
+import json
+import glob
+
+decoder = json.JSONDecoder()
+
+def read_from_glob(fun, g):
+    paths = glob.glob(g)
+    results = []
+    for path in paths:
+        with open(path) as f:
+            try:
+                results.append(fun(f))
+            except Exception as e:
+                print(e)
+    return results
+
+def read_python(f):
+    last = next(f, "")
+    obj = decoder.decode(last)
+    return obj['fix']
+
+def python_sources_from_glob(g):
+    return read_from_glob(read_python, g)
+
+def python_sources_from_glob(g):
+    paths = glob.glob(g)
+    sources = []
+    for path in paths:
+        with open(path) as f:
+            last = next(f, "")
+            try:
+                obj = decoder.decode(last)
+                source = obj['fix']
+                sources.append(source)
+            except Exception as e:
+                print(e)
+    return sources

### 2018-08-13 cluster/string_id.py

@@ -1,0 +1,21 @@
+class StringIdMaker(object):
+    def __init__(self):
+        self.id_map = {}
+        self.next_id = 0
+    
+    """
+    Input: string
+    Output: an int id
+
+    Don't save this out anywhere, it can change depending on the order of calls to it
+    They are guaranteed within one session to have a 1-to-1 correspondence with input
+    strings.
+    """
+    def make_id(self, str):
+        if str not in self.id_map:
+            self.id_map[str] = self.next_id
+            self.next_id += 1
+
+        return self.id_map[str]
+
+

### 2018-08-13 cluster/stringify.py

@@ -1,0 +1,70 @@
+import ast
+import json
+import Levenshtein as lev
+from string_id import StringIdMaker
+
+ider = StringIdMaker()
+
+class Node():
+    def __init__(self, a):
+        self.children = list(map(Node, ast.iter_child_nodes(a)))
+        self.name = ider.make_id(type(a))
+        
+def node_transform(a):
+    return Node(a)
+
+def build_node_strings(out_list, node):
+    # Mark the beginning of this node by an id for that node type
+    out_list.append(chr(node.name))
+
+    # Add the contents of this node
+    for child in node.children:
+        build_node_strings(out_list, child)
+
+    # Token marking end of node
+    out_list.append(chr(ider.make_id('end')))
+
+# Takes 1 element arrays containing the strings
+def metric(a, b):
+    return lev.distance(a[0],b[0])
+
+def transform_python(a):
+    node = node_transform(a)
+    string_list = []
+    build_node_strings(string_list, node)
+    return ''.join(string_list)
+
+ocaml_node_list = ['int', 'double', 'bool', 'char', 'string',
+                   'let', 'letrec', 'eval', 'type', 'exception',
+                   'var', 'fun', 'app', 'bop', 'uop', 'lit', 'ite', 'seq', 'case',
+                   'tuple', 'con', 'record', 'getField', 'setField', 'array',
+                   'list', 'try',
+                   'interval', 'const', 'wild', 'or', 'as', 'type']
+
+ocaml_node_set = set(ocaml_node_list)
+
+def build_ocaml_strings(out_list, j):
+    if isinstance(j, list):
+        for element in j:
+            build_ocaml_strings(out_list, element)
+    elif isinstance(j, dict):
+        # Sort them so that children in dicts are always output in the same order
+        keys = sorted(j.keys())
+
+        for k in keys:
+            if k in ocaml_node_set:
+                # This is part of the ast where multiple node types would be
+                # valid. Add this to the output to preserve this info
+                out_list.append(chr(ider.make_id(k)))
+                build_ocaml_strings(out_list, j[k])
+
+                # End of node marker
+                out_list.append(chr(ider.make_id('end')))
+            else:
+                build_ocaml_strings(out_list, j[k])
+                
+
+def transform_ocaml(j):
+    string_list = []
+    build_ocaml_strings(string_list, j)
+    return ''.join(string_list)

### 2018-08-15 cluster/cluster.py
clustered data

@@ -1,17 +1,11 @@
-import scipy.cluster.hierarchy
 import numpy as np
 import os
 from scipy.spatial.distance import pdist
+from scipy.cluster.hierarchy import linkage, fcluster
 import stringify
 
 def distance_matrix(strings):
     return pdist(list(map(lambda x: [x], strings)), stringify.metric)
-
-def linkage(d_matrix):
-    return scipy.cluster.hierarchy.linkage(d_matrix, method='single')
-
-def cluster(link, t):
-    return scipy.cluster.hierarchy.fcluster(link, t)
 
 # Takes a list of python source strings and produces an array where element i is
 # the cluster number of the string with index i in the iterable that was passed
@@ -31,22 +25,15 @@ def cluster_strings(strings, data_name):
     dmat = distance_matrix(strings)
     dmat.tofile(data_name + '_progress/dmat')
 
-    print('Computing linkage...')
-    link = linkage(dmat)
-    link.tofile(data_name + '_progress/link')
+    for method in ['single', 'average', 'complete']:
+        print('Computing linkage (%s)...' % method)
+        link = linkage(dmat, method=method)
+        link.tofile(data_name + '_progress/link_' + method)
 
-    print('Computing clusters...')
-    cluster_counts = []
-    for i in range(200):
-        t = i / 100
-        cluster_counts.append(cluster(link, t).max())
-    np.array(cluster_counts).tofile(data_name + '_cluster_counts', sep='\n')
-
-def cluster_only():
-    link = np.fromfile(data_name + '_progress/link').reshape(-1, 4)
-    cluster_counts = []
-    for i in range(200):
-        t = i / 100
-        cluster_counts.append(cluster(link, t).max())
-    np.array(cluster_counts).tofile(data_name + '_cluster_counts', sep='\n')
+        print('Computing clusters (%s)...' % method)
+        cluster_counts = []
+        for i in range(201):
+            t = i / 100
+            cluster_counts.append(fcluster(link, t).max())
+            np.array(cluster_counts).tofile(data_name + '_cluster_counts_' + method, sep='\n')
     

### 2018-08-15 cluster/process.py

@@ -47,4 +47,4 @@ selection = [transformeds[i] for i in indices]
 print('Clustering...')
 cluster.cluster_strings(selection, data_name)
 
-print('Done clustering. Cluster counts output to %s_cluster_counts.' % data_name)
+print('Done clustering. Cluster counts output to %s_cluster_counts_*.' % data_name)

### 2018-08-15 paper/main.tex
graphs

@@
 \usepackage{xcolor}
+\usepackage{pgfplots}
 \def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
@@
 
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[
+    ybar,
+    symbolic x coords={Top-1,Top-3},
+    xtick=data,
+    ymin=0,
+    ylabel=Model Accuracy,
+    enlarge x limits=0.5,
+    legend style={at={(0.5,-0.15)},anchor=north}
+]
+    \addplot table[x=scoreName, y=withEverything, col sep=comma]{fault-localization.csv};
+    \addplot table[x=scoreName, y=baseline, col sep=comma]{fault-localization.csv};
+\legend{Our best model,Python interpreter}
+\end{axis}
+\end{tikzpicture}
+\caption{Based on a decision tree on the largest
+dataset (272534 pairs). TODO: also train for Top-2}
+\label{fig:full-dataset-acc}
+\end{figure}
+
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[
+    ybar,
+    symbolic x coords={Top-1,Top-3},
+    xtick=data,
+    ymin=0,
+    ylabel=Model Accuracy,
+    enlarge x limits=0.5,
+    legend style={at={(0.5,-0.15)},anchor=north}
+]
+    \addplot table[x=scoreName, y=all, col sep=comma]{removing-features.csv};
+    \addplot table[x=scoreName, y=noTypes, col sep=comma]{removing-features.csv};
+    \addplot table[x=scoreName, y=noContext, col sep=comma]{removing-features.csv};
+    \addplot table[x=scoreName, y=noSlice, col sep=comma]{removing-features.csv};
+\legend{All Features,No Types, No Context, No Slicing}
+\end{axis}
+\end{tikzpicture}
+\caption{Accuracy when features are removed, based on decision trees on a random subset of 20000 pairs.
+TODO: train better model and/or
+train on greater fraction of the data? TODO: Top-2. NOTE: removing slicing makes basically no difference...}
+\label{fig:removing-features}
+\end{figure}
+
 prereq: define accuracy
-
-PLOT Type: Bar Graph (with multiple bars)
-
-PLOT Y Axis: Model Accuracy
-
-PLOT X Axis: Top-1, Top-2, Top3
-
-PLOT Bars: 1 for our algorithm, 1 for a weak baseline (such as the vanilla
-Python interpreter)
 
@@
 
-Ranjit: ``our algorithms'' --- ``purely syntactic vs. with types''
---- ``with or without slicing'' as extra ``bars''
+PLOT: See Figs. \ref{fig:full-dataset-acc} and \ref{fig:removing-features}
 
 \subsection{RQ 2 --- Feature Predictive Power}
+
+\begin{table}[]
+\begin{tabular}{l|l|l}
+Feature Name & Feature Category & Feature Predictive Power \\ \hline
+F-Type-C1=unknown & Contextual (Type) & 0.112 \\
+F-Type-P=TUPLE & Contextual (Type) & 0.108 \\
+F-Stmt-Constr-C2=Return & Contextual (Syntactic) & 0.068 \\
+F-Expr-Constr=Paren & Syntactic & 0.059 \\
+F-InSlice=0.0 & Slicing & 0.044 \\
+F-Type-P=str & Contextual (Type) & 0.036 \\
+F-Op-Contr-C3=BinaryAnd & Contextual (Syntactic) & 0.026 \\
+F-Type-C2= & Contextual (Type) & 0.025 \\
+F-Type-C1=str & Contextual (Type) & 0.023 \\
+F-Type-P=float & Contextual (Type) & 0.019 \\
+\end{tabular}
+\caption{Based on the Top-1 Decision Tree on the full dataset. Feature Predictive Power here is the Gini Importance. TODO: add
+better-known measure like ANOVA or ReliefF. TODO: add a couple more features?}
+\label{tab:feature-predictive-power}
+\end{table}
 
@@
 
-PLOT Type: Table
-
-TABLE COLUMNS: Feature Name, Feature Category, Feature Predictive Power
-
-TABLE ROWS: top twelve features
+PLOT: See Table \ref{tab:feature-predictive-power}
 
@@
 TODO: what about turning on/off types? (is that separate from slicing?)
-
-TODO: do ANOVA or ReliefF
 

### 2018-08-17 cluster/parse.py
baseline data

@@ -3,24 +3,14 @@ import glob
 
 decoder = json.JSONDecoder()
 
-def read_from_glob(fun, g):
+def read_files_from_glob(g):
     paths = glob.glob(g)
-    results = []
+    sources = []
     for path in paths:
         with open(path) as f:
-            try:
-                results.append(fun(f))
-            except Exception as e:
-                print(e)
-    return results
+            sources.append(f.read())
 
-def read_python(f):
-    last = next(f, "")
-    obj = decoder.decode(last)
-    return obj['fix']
-
-def python_sources_from_glob(g):
-    return read_from_glob(read_python, g)
+    return sources
 
 def python_sources_from_glob(g):
     paths = glob.glob(g)

### 2018-08-17 cluster/process.py

@@ -1,4 +1,4 @@
-from parse import python_sources_from_glob
+from parse import python_sources_from_glob, read_files_from_glob
 import sys
 import cluster
 import random
@@ -15,6 +15,19 @@ def read_python():
     asts = map(ast.parse, strings)
     return list(map(stringify.transform_python, asts))
 
+# Note: could be faster if one file is read / transformed
+def read_onecluster():
+    return random.sample(read_python(), 1) * 5000
+
+# Note: could be faster if only twenty files are read / transformed
+def read_twenty():
+    print('Building list of programs...')
+    sources = read_files_from_glob(sys.argv[1])
+
+    print('Transforming...')
+    asts = map(ast.parse, sources)
+    return list(map(stringify.transform_python, asts))
+
 def read_ocaml():
     decoder = json.JSONDecoder()
     
@@ -27,7 +40,7 @@ def read_ocaml():
 
     return strings
 
-data_name = ''
+data_count = 5000
 
 if len(sys.argv) == 1:
     data_name = 'ocaml'
@@ -35,12 +48,24 @@ if len(sys.argv) == 1:
 elif len(sys.argv) == 2:
     data_name = 'python'
     transformeds = read_python()
+elif len(sys.argv) == 3:
+    # Sort of a hack baselines here
+    if sys.argv[2] == 'onecluster':
+        data_name = 'python_onecluster'
+        transformeds = read_onecluster()
+    elif sys.argv[2] == 'twenty':
+        data_name = 'python_twenty'
+        transformeds = read_twenty()
+        data_count = len(transformeds)
+    else:
+        print('Baselines should be one of onecluster, twenty')
+        raise 'die'
 else:
     print('0 args for ocaml, 1 arg for python')
-    assert(false)
+    raise 'die'
 
 print('Sampling programs...')
-indices = random.sample(list(range(len(transformeds))), 5000)
+indices = random.sample(list(range(len(transformeds))), data_count)
 np.array(indices).tofile(data_name + '_indices', sep='\n')
 selection = [transformeds[i] for i in indices]
 

### 2018-08-17 cluster/twenty/avl.py

@@ -1,0 +1,317 @@
+# Copied from https://rosettacode.org/wiki/AVL_tree#Python 
+"""
+ 
+Python AVL tree example based on 
+ 
+https://ocw.mit.edu/courses/electrical-engineering-and-computer-science/6-006-introduction-to-algorithms-fall-2011/lecture-videos/lec06_code.zip
+ 
+Simplified for Rosetta Code example.
+ 
+See also:
+ 
+https://ocw.mit.edu/courses/electrical-engineering-and-computer-science/6-006-introduction-to-algorithms-fall-2011/lecture-videos/MIT6_006F11_lec06_orig.pdf
+ 
+https://ocw.mit.edu/courses/electrical-engineering-and-computer-science/6-006-introduction-to-algorithms-fall-2011/lecture-videos/lecture-6-avl-trees-avl-sort/
+ 
+"""
+ 
+class AVLNode(object):
+    """A node in the AVL tree."""
+ 
+    def __init__(self, parent, k):
+        """Creates a node.
+ 
+        Args:
+            parent: The node's parent.
+            k: key of the node.
+        """
+        self.key = k
+        self.parent = parent
+        self.left = None
+        self.right = None
+ 
+    def _str(self):
+        """Internal method for ASCII art."""
+        label = str(self.key)
+        if self.left is None:
+            left_lines, left_pos, left_width = [], 0, 0
+        else:
+            left_lines, left_pos, left_width = self.left._str()
+        if self.right is None:
+            right_lines, right_pos, right_width = [], 0, 0
+        else:
+            right_lines, right_pos, right_width = self.right._str()
+        middle = max(right_pos + left_width - left_pos + 1, len(label), 2)
+        pos = left_pos + middle // 2
+        width = left_pos + middle + right_width - right_pos
+        while len(left_lines) < len(right_lines):
+            left_lines.append(' ' * left_width)
+        while len(right_lines) < len(left_lines):
+            right_lines.append(' ' * right_width)
+        if (middle - len(label)) % 2 == 1 and self.parent is not None and \
+           self is self.parent.left and len(label) < middle:
+            label += '.'
+        label = label.center(middle, '.')
+        if label[0] == '.': label = ' ' + label[1:]
+        if label[-1] == '.': label = label[:-1] + ' '
+        lines = [' ' * left_pos + label + ' ' * (right_width - right_pos),
+                 ' ' * left_pos + '/' + ' ' * (middle-2) +
[…]

### 2018-08-17 cluster/twenty/csv.py

@@ -1,0 +1,12 @@
+# Copied from https://rosettacode.org/wiki/CSV_data_manipulation#Python
+import fileinput
+ 
+changerow, changecolumn, changevalue = 2, 4, '"Spam"'
+ 
+with fileinput.input('csv_data_manipulation.csv', inplace=True) as f:
+    for line in f:
+        if fileinput.filelineno() == changerow:
+            fields = line.rstrip().split(',')
+            fields[changecolumn-1] = changevalue
+            line = ','.join(fields) + '\n'
+        print(line, end='')

### 2018-08-17 cluster/twenty/currency.py

@@ -1,0 +1,27 @@
+# Copied from https://rosettacode.org/wiki/Currency#Python
+from decimal import Decimal as D
+from collections import namedtuple
+ 
+Item = namedtuple('Item', 'price, quant')
+ 
+items = dict( hamburger=Item(D('5.50'), D('4000000000000000')),
+              milkshake=Item(D('2.86'), D('2')) )
+tax_rate = D('0.0765')
+ 
+fmt = "%-10s %8s %18s %22s"
+print(fmt % tuple('Item Price Quantity Extension'.upper().split()))
+ 
+total_before_tax = 0
+for item, (price, quant) in sorted(items.items()):
+    ext = price * quant
+    print(fmt % (item, price, quant, ext))
+    total_before_tax += ext
+print(fmt % ('', '', '', '--------------------'))
+print(fmt % ('', '', 'subtotal', total_before_tax))
+ 
+tax = (tax_rate * total_before_tax).quantize(D('0.00'))
+print(fmt % ('', '', 'Tax', tax))
+ 
+total = total_before_tax + tax
+print(fmt % ('', '', '', '--------------------'))
+print(fmt % ('', '', 'Total', total))

### 2018-08-17 cluster/twenty/fft.py

@@ -1,0 +1,14 @@
+# Copied from https://rosettacode.org/wiki/Fast_Fourier_transform#Python
+from cmath import exp, pi
+ 
+def fft(x):
+    N = len(x)
+    if N <= 1: return x
+    even = fft(x[0::2])
+    odd =  fft(x[1::2])
+    T= [exp(-2j*pi*k/N)*odd[k] for k in range(N//2)]
+    return [even[k] + T[k] for k in range(N//2)] + \
+           [even[k] - T[k] for k in range(N//2)]
+ 
+print( ' '.join("%5.3f" % abs(f) 
+                for f in fft([1.0, 1.0, 1.0, 1.0, 0.0, 0.0, 0.0, 0.0])) )

### 2018-08-17 cluster/twenty/fizzbuzz.py

@@ -1,0 +1,11 @@
+# Copied from https://rosettacode.org/wiki/General_FizzBuzz#Python
+def genfizzbuzz(factorwords, numbers):
+    factorwords.sort(key=lambda p: p[0])
+    lines = []
+    for num in numbers:
+        words = ''.join(wrd for fact, wrd in factorwords if (num % fact) == 0)
+        lines.append(words if words else str(num))
+    return '\n'.join(lines)
+ 
+if __name__ == '__main__':
+    print(genfizzbuzz([(5, 'Buzz'), (3, 'Fizz'), (7, 'Baxx')], range(1, 21)))

### 2018-08-17 cluster/twenty/gcd.py

@@ -1,0 +1,24 @@
+# Copied from https://rosettacode.org/wiki/Greatest_common_divisor#Python
+def gcd_bin(u, v):
+    u, v = abs(u), abs(v) # u >= 0, v >= 0
+    if u < v:
+        u, v = v, u # u >= v >= 0
+    if v == 0:
+        return u
+ 
+    # u >= v > 0
+    k = 1
+    while u & 1 == 0 and v & 1 == 0: # u, v - even
+        u >>= 1; v >>= 1
+        k <<= 1
+ 
+    t = -v if u & 1 else u
+    while t:
+        while t & 1 == 0:
+            t >>= 1
+        if t > 0:
+            u = t
+        else:
+            v = -t
+        t = u - v
+    return u * k

### 2018-08-17 cluster/twenty/hashjoin.py

@@ -1,0 +1,24 @@
+# Copied from https://rosettacode.org/wiki/Hash_join#Python
+from collections import defaultdict
+ 
+def hashJoin(table1, index1, table2, index2):
+    h = defaultdict(list)
+    # hash phase
+    for s in table1:
+        h[s[index1]].append(s)
+    # join phase
+    return [(s, r) for r in table2 for s in h[r[index2]]]
+ 
+table1 = [(27, "Jonah"),
+          (18, "Alan"),
+          (28, "Glory"),
+          (18, "Popeye"),
+          (28, "Alan")]
+table2 = [("Jonah", "Whales"),
+          ("Jonah", "Spiders"),
+          ("Alan", "Ghosts"),
+          ("Alan", "Zombies"),
+          ("Glory", "Buffy")]
+ 
+for row in hashJoin(table1, 1, table2, 0):
+    print(row)

### 2018-08-17 cluster/twenty/knightstour.py

@@ -1,0 +1,66 @@
+# Copied from https://rosettacode.org/wiki/Knight%27s_tour#Python
+import copy
+ 
+boardsize=6
+_kmoves = ((2,1), (1,2), (-1,2), (-2,1), (-2,-1), (-1,-2), (1,-2), (2,-1)) 
+ 
+ 
+def chess2index(chess, boardsize=boardsize):
+    'Convert Algebraic chess notation to internal index format'
+    chess = chess.strip().lower()
+    x = ord(chess[0]) - ord('a')
+    y = boardsize - int(chess[1:])
+    return (x, y)
+ 
+def boardstring(board, boardsize=boardsize):
+    r = range(boardsize)
+    lines = ''
+    for y in r:
+        lines += '\n' + ','.join('%2i' % board[(x,y)] if board[(x,y)] else '  '
+                                 for x in r)
+    return lines
+ 
+def knightmoves(board, P, boardsize=boardsize):
+    Px, Py = P
+    kmoves = set((Px+x, Py+y) for x,y in _kmoves)
+    kmoves = set( (x,y)
+                  for x,y in kmoves
+                  if 0 <= x < boardsize
+                     and 0 <= y < boardsize
+                     and not board[(x,y)] )
+    return kmoves
+ 
+def accessibility(board, P, boardsize=boardsize):
+    access = []
+    brd = copy.deepcopy(board)
+    for pos in knightmoves(board, P, boardsize=boardsize):
+        brd[pos] = -1
+        access.append( (len(knightmoves(brd, pos, boardsize=boardsize)), pos) )
+        brd[pos] = 0
+    return access
+ 
+def knights_tour(start, boardsize=boardsize, _debug=False):
+    board = {(x,y):0 for x in range(boardsize) for y in range(boardsize)}
+    move = 1
+    P = chess2index(start, boardsize)
+    board[P] = move
+    move += 1
+    if _debug:
+        print(boardstring(board, boardsize=boardsize))
+    while move <= len(board):
+        P = min(accessibility(board, P, boardsize))[1]
+        board[P] = move
+        move += 1
+        if _debug:
+            print(boardstring(board, boardsize=boardsize))
+            input('\n%2i next: ' % move)
+    return board
+ 
+if __name__ == '__main__':
+    while 1:
+        boardsize = int(input('\nboardsize: '))
+        if boardsize < 5:
+            continue
+        start = input('Start position: ')
+        board = knights_tour(start, boardsize)
+        print(boardstring(board, boardsize=boardsize))

### 2018-08-17 cluster/twenty/levenshtein.py

@@ -1,0 +1,19 @@
+# Copied from https://rosettacode.org/wiki/Levenshtein_distance#Python
+def minimumEditDistance(s1,s2):
+    if len(s1) > len(s2):
+        s1,s2 = s2,s1
+    distances = range(len(s1) + 1)
+    for index2,char2 in enumerate(s2):
+        newDistances = [index2+1]
+        for index1,char1 in enumerate(s1):
+            if char1 == char2:
+                newDistances.append(distances[index1])
+            else:
+                newDistances.append(1 + min((distances[index1],
+                                             distances[index1+1],
+                                             newDistances[-1])))
+        distances = newDistances
+    return distances[-1]
+ 
+print(minimumEditDistance("kitten","sitting"))
+print(minimumEditDistance("rosettacode","raisethysword"))

### 2018-08-17 cluster/twenty/matrixchain.py

@@ -1,0 +1,30 @@
+# Copied from https://rosettacode.org/wiki/Matrix_chain_multiplication#Python
+def optim4(a):
+    global u
+    n = len(a) - 1
+    u = [None] * n
+    u[0] = [[None, 0]] * n
+    for j in range(1, n):
+        v = [None] * (n - j)
+        for i in range(n - j):
+            m = None
+            for k in range(j):
+                s1, c1 = u[k][i]
+                s2, c2 = u[j - k - 1][i + k + 1]
+                c = c1 + c2 + a[i] * a[i + k + 1] * a[i + j + 1]
+                if m is None or c < m:
+                    s = k
+                    m = c
+            v[i] = [s, m]
+        u[j] = v
+    def aux(i, j):
+        s, c = u[j][i]
+        if s is None:
+            return i
+        else:
+            return [aux(i, s), aux(i + s + 1, j - s - 1)]
+    return u[n - 1][0][1], aux(0, n - 1)
+ 
+ 
+print(optim4([1, 5, 25, 30, 100, 70, 2, 1, 100, 250, 1, 1000, 2]))
+print(optim4([1000, 1, 500, 12, 1, 700, 2500, 3, 2, 5, 14, 10]))

### 2018-08-17 cluster/twenty/minesweeper.py

@@ -1,0 +1,137 @@
+# Copied from https://rosettacode.org/wiki/Minesweeper_game#Python 
+'''
+Minesweeper game.
+ 
+    There is an n by m grid that has a random number of between 20% to 60%
+    of randomly hidden mines that need to be found. 
+ 
+    Positions in the grid are modified by entering their coordinates
+    where the first coordinate is horizontal in the grid and the second
+    vertical. The top left of the grid is position 1,1; the bottom right is
+    at n,m.
+ 
+    * The total number of mines to be found is shown at the beginning of the
+    game.
+    * Each mine occupies a single grid point, and its position is initially
+    unknown to the player
+    * The grid is shown as a rectangle of characters between moves.
+    * You are initially shown all grids as obscured, by a single dot '.'
+    * You may mark what you think is the position of a mine which will show
+    as a '?'
+    * You can mark what you think is free space by entering its coordinates.
+    :*  If the point is free space then it is cleared, as are any adjacent
+    points that are also free space- this is repeated recursively for
+    subsequent adjacent free points unless that point is marked as a mine
+    or is a mine.
+    ::*   Points marked as a mine show as a '?'.
+    ::*   Other free points show as an integer count of the number of adjacent
+    true mines in its immediate neighbourhood, or as a single space ' ' if the
+    free point is not adjacent to any true mines.
+    * Of course you loose if you try to clear space that starts on a mine.
+    * You win when you have correctly identified all mines.
+ 
+ 
+    When prompted you may:
+        Toggle where you think a mine is at position x, y:
+          m <x> <y>
+        Clear the grid starting at position x, y (and print the result):
+          c <x> <y>
+        Print the grid so far:
+          p
+        Resign
+          r
+    Resigning will first show the grid with an 'N' for unfound true mines, a
+    'Y' for found true mines and a '?' for where you marked clear space as a
+    mine 
+ 
+'''
+ 
+ 
+gridsize  = (6, 4)
+minerange = (0.2, 0.6)
+ 
+ 
+try:
+    raw_input
+except:
+    raw_input = input
+ 
+import random
+from itertools import product
+from pprint import pprint as pp
+ 
+ 
+def gridandmines(gridsize=gridsize, minerange=minerange):
+    xgrid, ygrid = gridsize
+    minmines, maxmines = minerange
[…]

### 2018-08-17 cluster/twenty/montyhall.py

@@ -1,0 +1,34 @@
+# Copied from https://rosettacode.org/wiki/Monty_Hall_problem#Python
+import random
+ #1 represents a car
+ #0 represent a goat
+ 
+stay = 0  #amount won if stay in the same position
+switch = 0 # amount won if you switch 
+ 
+for i in range(1000):
+    lst = [1,0,0]           # one car and two goats
+    random.shuffle(lst)     # shuffles the list randomly
+ 
+    ran = random.randrange(3) # gets a random number for the random guess
+ 
+    user = lst[ran] #storing the random guess 
+ 
+    del(lst[ran]) # deleting the random guess
+ 
+    huh = 0
+    for i in lst: # getting a value 0 and deleting it
+        if i ==0:
+            del(lst[huh]) # deletes a goat when it finds it
+            break
+        huh+=1
+ 
+    if user ==1: # if the original choice is 1 then stay adds 1
+        stay+=1
+ 
+    if lst[0] == 1: # if the switched value is 1 then switch adds 1
+        switch+=1
+ 
+print("Stay =",stay)
+print("Switch = ",switch)
+#Done by Sam Witton 09/04/2014

### 2018-08-17 cluster/twenty/radixsort.py

@@ -1,0 +1,47 @@
+# Copied from https://rosettacode.org/wiki/Sorting_algorithms/Radix_sort#Python
+from math import log
+ 
+def getDigit(num, base, digit_num):
+    # pulls the selected digit
+    return (num // base ** digit_num) % base  
+ 
+def makeBlanks(size):
+    # create a list of empty lists to hold the split by digit
+    return [ [] for i in range(size) ]  
+ 
+def split(a_list, base, digit_num):
+    buckets = makeBlanks(base)
+    for num in a_list:
+        # append the number to the list selected by the digit
+        buckets[getDigit(num, base, digit_num)].append(num)  
+    return buckets
+ 
+# concatenate the lists back in order for the next step
+def merge(a_list):
+    new_list = []
+    for sublist in a_list:
+       new_list.extend(sublist)
+    return new_list
+ 
+def maxAbs(a_list):
+    # largest abs value element of a list
+    return max(abs(num) for num in a_list)
+ 
+def split_by_sign(a_list):
+    # splits values by sign - negative values go to the first bucket,
+    # non-negative ones into the second
+    buckets = [[], []]
+    for num in a_list:
+        if num < 0:
+            buckets[0].append(num)
+        else:
+            buckets[1].append(num)
+    return buckets
+ 
+def radixSort(a_list, base):
+    # there are as many passes as there are digits in the longest number
+    passes = int(round(log(maxAbs(a_list), base)) + 1) 
+    new_list = list(a_list)
+    for digit_num in range(passes):
+        new_list = merge(split(new_list, base, digit_num))
+    return merge(split_by_sign(new_list))

### 2018-08-17 cluster/twenty/rot13.py

@@ -1,0 +1,21 @@
+#!/usr/bin/env python
+# Copied from https://rosettacode.org/wiki/Rot-13#Python
+import string
+ 
+TRANSLATION_TABLE = str.maketrans(
+    string.ascii_uppercase + string.ascii_lowercase,
+    string.ascii_uppercase[13:] + string.ascii_uppercase[:13] +
+    string.ascii_lowercase[13:] + string.ascii_lowercase[:13]
+)
+ 
+ 
+def rot13(s):
+    """Return the rot-13 encoding of s."""
+    return s.translate(TRANSLATION_TABLE)
+ 
+ 
+if __name__ == "__main__":
+    """rot-13 encode the input files, or stdin if no files are provided."""
+    import fileinput
+    for line in fileinput.input():
+        print(rot13(line), end="")

### 2018-08-17 cluster/twenty/supersequence.py

@@ -1,0 +1,22 @@
+# Copied from https://rosettacode.org/wiki/Shortest_common_supersequence#Python
+# Use the Longest Common Subsequence algorithm
+ 
+def shortest_common_supersequence(a, b):
+    lcs = longest_common_subsequence(a, b)
+    scs = ""
+    # Consume lcs
+    while len(lcs) > 0:
+        if a[0]==lcs[0] and b[0]==lcs[0]:
+        # Part of the LCS, so consume from all strings
+            scs += lcs[0]
+            lcs = lcs[1:]
+            a = a[1:]
+            b = b[1:]
+        elif a[0]==lcs[0]:
+            scs += b[0]
+            b = b[1:]
+        else:
+            scs += a[0]
+            a = a[1:]
+    # append remaining characters
+    return scs + a + b

### 2018-08-17 cluster/twenty/toposort.py

@@ -1,0 +1,38 @@
+# Copied from https://rosettacode.org/wiki/Topological_sort#Python
+
+try:
+    from functools import reduce
+except:
+    pass
+ 
+data = {
+    'des_system_lib':   set('std synopsys std_cell_lib des_system_lib dw02 dw01 ramlib ieee'.split()),
+    'dw01':             set('ieee dw01 dware gtech'.split()),
+    'dw02':             set('ieee dw02 dware'.split()),
+    'dw03':             set('std synopsys dware dw03 dw02 dw01 ieee gtech'.split()),
+    'dw04':             set('dw04 ieee dw01 dware gtech'.split()),
+    'dw05':             set('dw05 ieee dware'.split()),
+    'dw06':             set('dw06 ieee dware'.split()),
+    'dw07':             set('ieee dware'.split()),
+    'dware':            set('ieee dware'.split()),
+    'gtech':            set('ieee gtech'.split()),
+    'ramlib':           set('std ieee'.split()),
+    'std_cell_lib':     set('ieee std_cell_lib'.split()),
+    'synopsys':         set(),
+    }
+ 
+def toposort2(data):
+    for k, v in data.items():
+        v.discard(k) # Ignore self dependencies
+    extra_items_in_deps = reduce(set.union, data.values()) - set(data.keys())
+    data.update({item:set() for item in extra_items_in_deps})
+    while True:
+        ordered = set(item for item,dep in data.items() if not dep)
+        if not ordered:
+            break
+        yield ' '.join(sorted(ordered))
+        data = {item: (dep - ordered) for item,dep in data.items()
+                if item not in ordered}
+    assert not data, "A cyclic dependency exists amongst %r" % data
+ 
+print ('\n'.join( toposort2(data) ))

### 2018-08-17 cluster/twenty/trabbpardoknuth.py

@@ -1,0 +1,18 @@
+# Copied from https://rosettacode.org/wiki/Trabb_Pardo%E2%80%93Knuth_algorithm#Python
+def f(x):
+    return abs(x) ** 0.5 + 5 * x**3
+ 
+def ask():
+    return [float(y)
+            for y in input('\n11 numbers: ').strip().split()[:11]]
+ 
+if __name__ == '__main__':
+    s = ask()
+    s.reverse()
+    for x in s:
+        result = f(x)
+        if result > 400:
+            print(' %s:%s' % (x, "TOO LARGE!"), end='')
+        else:
+            print(' %s:%s' % (x, result), end='')
+    print('')

### 2018-08-17 cluster/twenty/turing.py

@@ -1,0 +1,88 @@
+# Copied from https://rosettacode.org/wiki/Universal_Turing_machine#Python
+from __future__ import print_function
+ 
+def run_utm(
+        state = None,
+        blank = None,
+        rules = [],
+        tape = [],
+        halt = None,
+        pos = 0):
+    st = state
+    if not tape: tape = [blank]
+    if pos < 0: pos += len(tape)
+    if pos >= len(tape) or pos < 0: raise Error( "bad init position")
+    rules = dict(((s0, v0), (v1, dr, s1)) for (s0, v0, v1, dr, s1) in rules)
+ 
+    while True:
+        print(st, '\t', end=" ")
+        for i, v in enumerate(tape):
+            if i == pos: print("[%s]" % (v,), end=" ")
+            else: print(v, end=" ")
+        print()
+ 
+        if st == halt: break
+        if (st, tape[pos]) not in rules: break
+ 
+        (v1, dr, s1) = rules[(st, tape[pos])]
+        tape[pos] = v1
+        if dr == 'left':
+            if pos > 0: pos -= 1
+            else: tape.insert(0, blank)
+        if dr == 'right':
+            pos += 1
+            if pos >= len(tape): tape.append(blank) 
+        st = s1
+ 
+ 
+# EXAMPLES
+ 
+print("incr machine\n")
+run_utm(
+    halt = 'qf',
+	state = 'q0',
+	tape = list("111"),
+	blank = 'B',
+	rules = map(tuple, 
+               ["q0 1 1 right q0".split(),
+		        "q0 B 1 stay  qf".split()]
+        )
+    )
+ 
+print("\nbusy beaver\n")
+run_utm(
+    halt = 'halt',
+	state = 'a',
+	blank = '0',
+	rules = map(tuple,
+        ["a 0 1 right b".split(),
+         "a 1 1 left  c".split(),
+         "b 0 1 left  a".split(),
+         "b 1 1 right b".split(),
+         "c 0 1 left  b".split(),
+         "c 1 1 stay  halt".split()]
+        )
+    )
+ 
+print("\nsorting test\n")
+run_utm(halt = 'STOP',
+	state = 'A',
+	blank = '0',
+	tape = "2 2 2 1 2 2 1 2 1 2 1 2 1 2".split(),
+	rules = map(tuple,
+       ["A 1 1 right A".split(),
[…]

### 2018-08-17 cluster/twenty/unbias.py

@@ -1,0 +1,29 @@
+# Copied from https://rosettacode.org/wiki/Unbias_a_random_generator#Python
+from __future__ import print_function
+import random
+ 
+def randN(N):
+    " 1,0 random generator factory with 1 appearing 1/N'th of the time"
+    return lambda: random.randrange(N) == 0
+ 
+def unbiased(biased):
+    'uses a biased() generator of 1 or 0, to create an unbiased one'
+    this, that = biased(), biased()
+    while this == that: # Loop until 10 or 01
+        this, that = biased(), biased()
+    return this         # return the first
+ 
+if __name__ == '__main__':
+    from collections import namedtuple
+ 
+    Stats = namedtuple('Stats', 'count1 count0 percent')
+ 
+    for N in range(3, 7):
+        biased = randN(N)
+        v = [biased() for x in range(1000000)]
+        v1, v0 = v.count(1), v.count(0)
+        print ( "Biased(%i)  = %r" % (N, Stats(v1, v0, 100. * v1/(v1 + v0))) )
+ 
+        v = [unbiased(biased) for x in range(1000000)]
+        v1, v0 = v.count(1), v.count(0)
+        print ( "  Unbiased = %r" % (Stats(v1, v0, 100. * v1/(v1 + v0)), ) )

### 2018-08-17 cluster/twenty/yinyang.py

@@ -1,0 +1,22 @@
+# Copied from https://rosettacode.org/wiki/Yin_and_yang#Python
+import math
+def yinyang(n=3):
+	radii   = [i * n for i in (1, 3, 6)]
+	ranges  = [list(range(-r, r+1)) for r in radii]
+	squares = [[ (x,y) for x in rnge for y in rnge]
+		   for rnge in ranges]
+	circles = [[ (x,y) for x,y in sqrpoints
+		     if math.hypot(x,y) <= radius ]
+		   for sqrpoints, radius in zip(squares, radii)]
+	m = {(x,y):' ' for x,y in squares[-1]}
+	for x,y in circles[-1]:
+		m[x,y] = '*'
+	for x,y in circles[-1]:
+		if x>0: m[(x,y)] = '·'
+	for x,y in circles[-2]:
+		m[(x,y+3*n)] = '*'
+		m[(x,y-3*n)] = '·'
+	for x,y in circles[-3]:
+		m[(x,y+3*n)] = '·'
+		m[(x,y-3*n)] = '*'
+	return '\n'.join(''.join(m[(x,y)] for x in reversed(ranges[-1])) for y in ranges[-1])

### 2018-08-17 cluster/make_csv.py
cluster counts as one big table

@@ -1,0 +1,24 @@
+import sys
+
+paths = sys.argv[1:]
+cluster_counts = {}
+
+for path in paths:
+    with open(path) as f:
+        cluster_counts[path] = []
+        for x in f:
+            cluster_counts[path].append(x.strip())
+
+with open('cluster_counts.csv', 'wt') as out_file:
+    out_file.write('threshold')
+
+    for path in paths:
+        out_file.write(',')
+        out_file.write(path)
+    
+    for i in range(200):
+        out_file.write(str(i / 100.0))
+        for path in paths:
+            out_file.write(',')
+            out_file.write(cluster_counts[path][i])
+        out_file.write('\n')

### 2018-08-18 paper/main.tex
paper updates

@@
 is the true program change located in the model's first k outputs.
+
+\section{Approach, in more detail}
+
+We follow the overall strategy of \cite{learning-to-blame} and represent each
+buggy program by a Bag of Abstracted Terms. In our setting, a `term' means either
+a statment or an expression. For each term in the program we compute the following features:
+
+\subsection{Dynamic features}
+
+We first convert each buggy program to an equivalent program that is closer
+to ANF, then run it through the PythonTutor backend. This backend is based on the
+Python debigging library BDB; it returns to us a trace of the program along with
+the state of the heap at each execution step. We use this information to compute
+the following features for each term:
+
+\emph{Type} The possible values for the Type feature include all the basic Python
+types (such as int and tuple) as well as a few special values:
+\begin{itemize}
+    \item IsStatement - Given to all statements, since our terms include both
+    statments and expressions but only expressions actually have types.
+    \item Unknown - Given to expressions that are never evaluated in the dynamic
+    trace.
+    \item Multiple - Given to expressions which are evaluated multiple times in
+    a trace and do not always have the same type.
+\end{itemize}
+
+\emph{Slice} (TODO - Leon can you expand this? What did you base our slicing on?)
+The goal of this feature is to help eliminate terms that cannot be the source of
+the crash. These include lines that are never run, lines that run successfully
+and don't affect any variables that are used later, and so on. More specifically, we
+build up a dependency graph of all lines in the program trace. Edges in the graph
+represent either def-use dependence or control-flow dependence. Then ideally,
+any node in the graph that is not a transitive dependency of the line that crashes
+can't be responsible for the crash. However dynamic slicing is not perfect (TODO
+include example?), so unlike in prior work, we do not treat membership in the slice
+as a hard constraint.
+
+\emph{Crash location} As discussed later in the evaluation, the precise term that
+causes Python to crash is in fact the one the user decides to change about 30\% of
+the time, so we use it as a feature (and expect it to be a high-signal one, too.)
+
+\emph{Exception type} The type of error thrown by the program is useful context for
+finding the error. For example, if asked how likely a random `0` is of being a bug, you'd give
+a higher answer if you knew that the program eventually crashes with a division by
+zero exception. Thus we attach the exception type as a feature to \emph{every}
+vector derived from that program.
+
+\subsection{Contextual features}
+
+The approach of representing a program by a Bag of Abstracted Terms, if implemented
+naively, runs into the problem that the terms lose their context in the program
+as a whole. For example, the features for the term `0` include its type (int) and
+syntactic description (int literal), but there is no way that's enough to decide
+if it's innocuous, like the 0 in `x = 0`, or clearly the source of a bug, like the
+0 in `x / 0`. To restore some of this context, we introduce contextual features.
+This involves duplicating all our other features four times - once for the parent
+and once each for a term's first three children in the AST. So since `x / 0` has
+the feature SyntacticForm = DivisionOp, the `0` of `x / 0` has the feature
+ParentSyntacticForm = DivisionOp. Similarly, since `0` has SyntacticForm = IntLiteral,
+`x / 0` has Child1SyntacticForm = IntLiteral. In order to make all feature vectors
+the same length, nodes that have no parent or fewer than three children have the
+relevant features filled out with a special NotApplicable value.
+
+\subsection{Static features}
+
+\emph{Syntactic form} The first feature we consider is what kind of node the term
+is in the AST. For statements this has values like Return or Import; for expressions
+Variable or Application.
+
+\emph{Size} We also count the number of descendent nodes in the subtree rooted at that node.
+
+\subsection{Machine learning}
+
+\emph{Pairing programs} For each PythonTutor user, we look at the sequence of
+programs they attempted to run. Each program that crashes is paired with the next
+program by that user which does not crash. Many programs do not make it to the
+end of our data pipeline; here are the most common reasons:
+\begin{itemize}
+    \item The user submitted only programs that crash or only programs that do
+    not crash, so we can't create any pairs for that user.
+    \item The program is out of scope for PythonTutor. For
+    example, the program imports forbidden libraries (like os), or the program
+    runs in over 1000 execution steps.
+    \item The proportion of the program that has changed between the buggy and
+    fixed versions is more than a standard deviation above the mean.
+\end{itemize}
+We end up with \textasciitilde270,000 usable pairs of programs.
+
+\emph{Training methodology} At this point we have feature vectors describing every term of every program in
+our dataset. The vectors from a random 30\% of the program pairs are set aside for testing and
+the other 70\% are used for training. Within the training data, we use 3-fold
+cross-validation to select the best parameters for our models, before training
+using those parameters on the full training set and then scoring the models
+using the testing set. Each trained model takes in any vector representing a term
+in a buggy program, and returns a confidence score representing how likely it is that
+that term was one of the terms changed between the fixed and buggy versions.
+Thus we can treat the model as providing a \emph{ranking} over all terms by
+confidence, where the top result is the one that the model deems most likely
+to be changed. For
+a given k, we score the model based on Top-k accuracy: for what proportion of the
+programs is a correct answer (i.e. a term that is actually changed) present in the
+top k results. We retrain the models for each value of k, since a model optimized
+to produce the single best result might be different from one optimized to get a
+good result into the top three.
+
+In each case the most effective model we trained was a decision tree. (The other
+models we tried were Logistic Regression and MLP. Yaoyuan - anything to add to
+this section?)
 
@@
 \bibliographystyle{abbrv}
-\bibliography{nanomaly,sw}
+\bibliography{nanomaly,sw,temp}
 

### 2018-08-18 paper/main.tex
Add diversity graphs

@@
 \author{\IEEEauthorblockN{Benjamin Cosman}
-\IEEEauthorblockA{\textit{UC San Digeo}\\
+\IEEEauthorblockA{\textit{UC San Diego}\\
 blcosman@eng.ecsd.edu}
@@
 \IEEEauthorblockN{Leon Medvinsky}
-\IEEEauthorblockA{\textit{UC San Digeo}\\
+\IEEEauthorblockA{\textit{UC San Diego}\\
 lmedvinsky@eng.ecsd.edu}
@@
 \IEEEauthorblockN{Ranjit Jhala}
-\IEEEauthorblockA{\textit{UC San Digeo}\\
+\IEEEauthorblockA{\textit{UC San Diego}\\
 jhala@cs.ecsd.edu}
@@
 
+\begin{figure}
+\foreach \method in {single, complete, average}
+{
+\begin{tikzpicture}
+  \begin{axis}[
+      xlabel=Threshold,
+      ylabel=\# of clusters,
+      ytick distance=2,
+      title=Linkage: \method
+    ]
+    \addplot table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+\end{axis}
+\end{tikzpicture}
+}
+\caption{The number of clusters when clustering a baseline of twenty dissimilar programs from
+  Rosetta Code.}
+\label{fig:diversity-baseline}
+\end{figure}
+
+\begin{figure}
+\foreach \method in {single, complete, average}
+{
+\begin{tikzpicture}
+  \begin{axis}[
+      xlabel=Threshold,
+      ylabel=\# of clusters,
+      ytick distance=1000,
+      title=Linkage: \method
+    ]
+    \foreach \language in {ocaml, python}
+    {
+      \addplot table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+    }
+\legend{OCaml, Python}
+\end{axis}
+\end{tikzpicture}
+}
+\caption{Number of clusters when data is clustered by single, complete, and average linkage.
+  A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
+  into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
+  of a link is its distance, z-scored against other nearby links.}
+\label{fig:diversity}
+\end{figure}
+
 \subsection{RQ 5 --- Wishlist: Partitioning Kinds of Errors}

### 2018-08-18 cluster/make_csv.py

@@ -15,6 +15,7 @@ with open('cluster_counts.csv', 'wt') as out_file:
     for path in paths:
         out_file.write(',')
         out_file.write(path)
+    out_file.write('\n')
     
     for i in range(200):
         out_file.write(str(i / 100.0))

### 2018-08-18 paper/main.tex
clean up diversity graphs

@@
     ]
-    \addplot table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+    \addplot+[mark=none] table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
 \end{axis}
@@
 \caption{The number of clusters when clustering a baseline of twenty dissimilar programs from
-  Rosetta Code.}
+  Rosetta Code. We give no plot for the baseline of 5000 identical programs, as they were
+  consistently placed in just 1 cluster.}
 \label{fig:diversity-baseline}
@@
     {
-      \addplot table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
     }

### 2018-08-19 paper/main.tex
update feature importances table

@@
 Feature Name & Feature Category & Feature Predictive Power \\ \hline
-F-Type-C1=unknown & Contextual (Type) & 0.112 \\
-F-Type-P=TUPLE & Contextual (Type) & 0.108 \\
-F-Stmt-Constr-C2=Return & Contextual (Syntactic) & 0.068 \\
-F-Expr-Constr=Paren & Syntactic & 0.059 \\
-F-InSlice=0.0 & Slicing & 0.044 \\
-F-Type-P=str & Contextual (Type) & 0.036 \\
-F-Op-Contr-C3=BinaryAnd & Contextual (Syntactic) & 0.026 \\
-F-Type-C2= & Contextual (Type) & 0.025 \\
-F-Type-C1=str & Contextual (Type) & 0.023 \\
-F-Type-P=float & Contextual (Type) & 0.019 \\
+F-Size-P & Contextual (Syntactic) & 0.112624 \\
+F-Stmt-Constr-C3= & Contextual (Syntactic) & 0.061552 \\
+F-Expr-Constr-P=List & Contextual (Syntactic) & 0.054632 \\
+F-PythonBlame & Dynamic (Error location) & 0.039189 \\
+F-Type=unknown & Dynamic (Type) & 0.037390 \\
+F-Type-P=unknown & Contextual (Type) & 0.037330 \\
+F-PythonMsg=IndexError & Dynamic (Error message) & 0.032749 \\
+F-Size & Syntactic & 0.028203 \\
+F-Expr-Constr-P=Dictionary & Contextual (Syntactic) & 0.026663 \\
+F-Size-C1 & Contextual (Syntactic) & 0.023662 \\
+F-Expr-Constr=Var & Syntactic & 0.020390 \\
+F-Type-C1=unknown & Contextual (Type) & 0.019815 \\
 \end{tabular}
-\caption{Based on the Top-1 Decision Tree on the full dataset. Feature Predictive Power here is the Gini Importance. TODO: add
-better-known measure like ANOVA or ReliefF. TODO: add a couple more features?}
+\caption{Based on the Top-3 Decision Tree on the full dataset. Feature Predictive Power here is the Gini Importance. TODO: add
+better-known measure like ANOVA or ReliefF.}
 \label{tab:feature-predictive-power}
@@
       ylabel=\# of clusters,
-      ytick distance=2,
+    %   ytick distance=2,
       title=Linkage: \method
@@
       ylabel=\# of clusters,
-      ytick distance=1000,
+    %   ytick distance=1000,
       title=Linkage: \method

### 2018-08-19 paper/main.tex
paper updates

@@
 
+\emph{Models} The following descriptions of Logistic Regression, Decision Trees,
+and MLP will have to be redone as they are copied from Learning to Blame.
+
+\emph{Logistic Regression}
+The simplest classifier we investigate is logistic regression:
+a linear model where the goal is to learn a set of weights $W$
+that describe the following model for predicting a label
+$b$ from a feature vector $v$:
+%
+\[ \Pr(b = 1 | v) = \frac{1}{1 + e^{-W^{\top} v}} \]
+%
+The weights $W$ are learnt from training data, and the value of
+$\Pr(b | v)$ naturally leads to a confidence score $\mathcal{C}$.
+%
+Logistic regression is a widely used classification algorithm, preferred
+for its simplicity, ease of generalization, and interpretability.
+%
+Its main limitation is that the prediction rule is constrained to be a
+linear combination of the features, and hence relatively simple.
+%
+While this can be somewhat mitigated by adding higher order (quadratic
+or cubic) features, this often requires substantial domain knowledge.
+
+\emph{Decision Trees}
+Decision tree algorithms learn a tree of binary predicates over the
+features, recursively partitioning the input space until a final
+classification can be made.
+%
+Each node in the tree contains a single predicate of the form
+$v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
+whether a given input should proceed down the left or right subtree.
+%
+Each leaf is labeled with a prediction and the fraction of
+correctly-labeled training samples that would reach it; the latter
+quantity can be interpreted as the decision tree's confidence in its
+prediction.
+%
+This leads to a prediction rule that can be quite expressive depending
+on the data used to build it.
+
+Training a decision tree entails finding both a set of good partitioning
+predicates and a good ordering of the predicates based on data.
+%
+This is usually done in a top-down greedy manner, and there are several
+standard training algorithms such as C4.5 \cite{Quinlan1993-de} and
+CART \cite{Breiman1984-qy}.
+
+Another advantage of decision trees is their ease of interpretation ---
+the decision rule is a white-box model that can be readily described to
+a human, especially when the tree is small.
+%
+However, the main limitation is that these trees often do not generalize
+well, though this can be somewhat mitigated by \emph{pruning} the tree.
+
+\emph{Neural Networks}
+%
+The last (and most complex) model we use is a type of neural network
+called a \emph{multi-layer perceptron} (see \cite{Nielsen2015-pu} for
+an introduction to neural networks).
+%
+A multi-layer perceptron can be represented as a directed acyclic
+graph whose nodes are arranged in layers that are fully connected by
+weighted edges.
+%
+The first layer corresponds to the input features, and the final to the
+output.
+%
+%The output of a node $v$ in an internal layer is given by:
+The output of an internal node $v$ is
+%
+\[ h_v = g(\sum_{j \in N(v)} W_{jv} h_j ) \]
+%
+where $N(v)$ is the set of nodes in the previous layer that are adjacent
+to $v$, $W_{jv}$ is the weight of the $(j, v)$ edge and $h_j$ is the
+output of node $j$ in the previous layer.
+%
+Finally $g$ is a non-linear function, called the activation function,
+which in recent work is commonly chosen to be the \emph{rectified linear
+  unit} (ReLU), defined as $g(x) = \mathsf{max}(0,x)$
+\cite{Nair2010-xg}.
+%
+The number of layers, the number of neurons per layer, and the
+connections between layers constitute the \emph{architecture} of a
+neural network.
+%
+In this work, we use relatively simple neural networks which have an
+input layer, a single hidden layer and an output layer.
+
+A major advantage of neural networks is their ability to discover
+interesting combinations of features through non-linearity, which
+significantly reduces the need for manual feature engineering, and
+allows high expressivity.
+%
+On the other hand, this makes the networks particularly difficult to
+interpret and also difficult to generalize unless vast amounts of
+training data are available.
+
 \emph{Training methodology} At this point we have feature vectors describing every term of every program in
@@
 In each case the most effective model we trained was a decision tree. (The other
-models we tried were Logistic Regression and MLP. Yaoyuan - anything to add to
-this section?)
+models we tried were Logistic Regression and MLP. Yao-Yuan - anything to add to
+this section? in particular, is it actually accurate to say we're doing
+cross-validation - the 3-fold thing we do is actually just to select parameters,
+and then there's no cross-validation on the final model, right?)
 
@@
 \caption{Based on the Top-3 Decision Tree on the full dataset. Feature Predictive Power here is the Gini Importance. TODO: add
-better-known measure like ANOVA or ReliefF.}
+better-known measure like ANOVA or ReliefF. TODO: base on Top-1 instead?}
 \label{tab:feature-predictive-power}
 \end{table}
+
+To measure how important each feature is, we use Gini Importance. In a decision
+tree,
 
@@
 
+\subsection{Threats to validity}
+
+\emph{Overfitting and other machine learning issues} Yao-Yuan, how do we avoid
+these?
+
+\emph{Language choice} We only showed that this technique works for Python 3.
+However, \cite{learning-to-blame} uses a similar technique to great effect in
+OCaml, and if it works on these two quite-different languages then we believe
+it would work on others as well. Additionally, Python is a common choice of
+language for introductory programming courses, so even a technique that worked
+only on Python would still be very useful for helping novices.
+
+\emph{Target population} We don't know exactly who our study population is -
+who wrote the programs in the dataset and what exactly they were trying to write.
+It is likely that they are primarily students and not professional software engineers,
+and thus possible that our technique does not generalize well to the kinds of
+bugs that would be found in an industrial setting rather than an educational one.
+
+\emph{Automatic labeling} Since our dataset is so large, we were forced to use an automatic
+scheme to label bugs rather than marking them by hand. We chose to use the diffs
+between each program that crashes and the next program that does not, but this method
+is not 100\% accurate because the next program may contain additional changes
+beyond what was needed to fix the bug, or it may indeed be an entirely unrelated
+program. We mitigate this problem by discarding as outliers program pairs that have
+changed too much. We also note that if our `true' accuracy is indeed a little
+smaller than our reported accuracy because of this effect, then the `true' baseline
+is also lower than the reported baseline for the same reason, so our
+\emph{margin} of effectiveness is unlikely to change much.
+
 \section{Related Work}

### 2018-08-20 paper/main.tex
rephrase the motex a bit

@@
 \usepackage{xcolor}
+\usepackage{listings}
+\usepackage{myref}
 \usepackage{pgfplots}
@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\newcommand\lt[1]{{\lstinline|#1|}}
+\lstset{language=python}
+\definecolor{dkgreen}{rgb}{0,0.5,0}
+\definecolor{dkred}{rgb}{0.5,0,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\lstset{basicstyle=\ttfamily\bfseries\footnotesize,
+  morekeywords={virtualinvoke},
+  keywordstyle=\color{blue},
+  ndkeywordstyle=\color{red},
+  commentstyle=\color{dkred},
+  stringstyle=\color{dkgreen},
+  numbers=left,
+  numberstyle=\ttfamily\footnotesize\color{gray},
+  stepnumber=1,
+  numbersep=10pt,
+  backgroundcolor=\color{white},
+  tabsize=4,
+  showspaces=false,
+  showstringspaces=false,
+  xleftmargin=.23in
+}
+
 \begin{document}
@@
 \begin{figure}
-\begin{verbatim}
+\begin{lstlisting}
 year = int(time.strftime("%Y"))
@@
 print(year + age)
-\end{verbatim}
-\caption{A buggy program}
+\end{lstlisting}
+\caption{
+\label{fig-motex} 
+A program with multiple fault localizations (e.g.,
+lines 1 and 2).} 
 \end{figure}
 
-Consider the program in Fig. 1, adapted from our dataset. The program
-attempts to use the current year and the user's age to compute their
-birth year and the year they can vote. However, while line 1 correctly
-turns the current year from a string into an int, this `int' call is
-missing from line 2 so `age' remains a string, and thus the addition
-on line 4 causes the program to crash.
+Consider the program in \figref{motex}, adapted from our dataset of
+human submissions. The program attempts to carry out some simple arithmetic
+based on a given number and the current year. When executed, however, the
+program raises an exception on line 4, related to the addition of an
+integer variable to a string variable.
 
-Notice that while a correct fix (and the one that the programmer
-actually used) is to add an `int' call to line 2, another possible ``fix''
-would be to \textit{remove} the `int' call from line 1 - both year and
-age would then be strings, and the addition on line 4 would be interpreted
-as string concatenation. This shows the necessity of a heuristic, data-driven
-approach, since without test cases there is no way to tell for sure which
-fix is better.
+One reasonable fix (and the one that the programmer actually used in this
+case) is to add \lt{int()} cast on line 2. However, another possible
+``fix'' would be to \textit{remove} the \lt{int()} cast from line 1, in
+which case both variables hold string values at runtime, and the addition
+on line 4 would be interpreted as string concatenation. While the second
+fix creates a well-typed program, it is less likely to correspond to
+developer intent, and is thus less likely to be helpful as a debugging aid.
 
-We can also start to see some features which will be
-important for finding bugs: we will need the \textit{type} of age, as
-well as some \textit{contextual} information about how it's used in the program.
+This simple example highlights a key fault localization challenge in such
+dynamically-typed settings: a runtime type error alone may not contain
+enough information to pinpoint the human-preferred localization from among
+many valid-but-less-helpful localizations. Our insight is that the relevant
+information is not in the types (e.g., string or integer) alone but in the
+context in which the values are used.  We propose an approach based on
+machine learning over a combination of static and dynamic contextual
+features.
 

### 2018-08-20 paper/main.tex
a little prose change near the feature predictive power table

@@
 \begin{table}[]
-\begin{tabular}{l|l|l}
-Feature Name & Feature Category & Feature Predictive Power \\ \hline
-F-Size-P & Contextual (Syntactic) & 0.112624 \\
-F-Stmt-Constr-C3= & Contextual (Syntactic) & 0.061552 \\
-F-Expr-Constr-P=List & Contextual (Syntactic) & 0.054632 \\
-F-PythonBlame & Dynamic (Error location) & 0.039189 \\
-F-Type=unknown & Dynamic (Type) & 0.037390 \\
-F-Type-P=unknown & Contextual (Type) & 0.037330 \\
-F-PythonMsg=IndexError & Dynamic (Error message) & 0.032749 \\
-F-Size & Syntactic & 0.028203 \\
-F-Expr-Constr-P=Dictionary & Contextual (Syntactic) & 0.026663 \\
-F-Size-C1 & Contextual (Syntactic) & 0.023662 \\
-F-Expr-Constr=Var & Syntactic & 0.020390 \\
-F-Type-C1=unknown & Contextual (Type) & 0.019815 \\
+\begin{tabular}{lll}
+Name & Category & Gini Importance \\ \bottomrule
+Size-P & Contextual (Syntactic)               & 0.112 \\
+Stmt-Constr-C3= & Contextual (Syntactic)      & 0.061 \\
+Expr-Constr-P=List & Contextual (Syntactic)   & 0.055 \\
+PythonBlame & Dynamic (Error location)        & 0.039 \\
+Type=unknown & Dynamic (Type)                 & 0.037 \\
+Type-P=unknown & Contextual (Type)            & 0.037 \\
+PythonMsg=IndexError & Dynamic (Error message)& 0.033 \\
+Size & Syntactic                              & 0.028 \\
+Expr-Constr-P=Dictionary & Contextual (Syntactic) & 0.027 \\
+Size-C1 & Contextual (Syntactic)              & 0.024 \\
+Expr-Constr=Var & Syntactic                   & 0.020 \\
+Type-C1=unknown & Contextual (Type)           & 0.020 \\
+\toprule
 \end{tabular}
-\caption{Based on the Top-3 Decision Tree on the full dataset. Feature Predictive Power here is the Gini Importance. TODO: add
+\caption{Feature predictive power (for a Top-3 Decision Tree 
+learned on the entire dataset). 
+TODO: add
 better-known measure like ANOVA or ReliefF. TODO: base on Top-1 instead?}
-\label{tab:feature-predictive-power}
+\label{tab-feature-predictive-power}
 \end{table}
 
-To measure how important each feature is, we use Gini Importance. In a decision
-tree,
+\tabref{feature-predictive-power} summarizes the relative importance 
+of the top features in our model. The features are ranked by their
+Gini importance (or mean decrease in impurity), a common measure
+for decision tree and random forest models~\cite{FIXME}. Informally, the
+Gini importance conveys a weighted count of the number of times a feature
+is used to split a node: a feature that is learned to guide more model
+classification decisions is more important.
 

### 2018-08-20 paper/main.tex
more slice info

@@
 
-\emph{Slice} (TODO - Leon can you expand this? What did you base our slicing on?)
-The goal of this feature is to help eliminate terms that cannot be the source of
-the crash. These include lines that are never run, lines that run successfully
-and don't affect any variables that are used later, and so on. More specifically, we
-build up a dependency graph of all lines in the program trace. Edges in the graph
-represent either def-use dependence or control-flow dependence. Then ideally,
-any node in the graph that is not a transitive dependency of the line that crashes
-can't be responsible for the crash. However dynamic slicing is not perfect (TODO
-include example?), so unlike in prior work, we do not treat membership in the slice
+\emph{Slice}
+\begin{figure}
+\begin{lstlisting}
+x = input() # The user inputs 0
+if x != 0:
+  x = 1
+print("One divided by your number is: %d" % (1 / x))
+\end{lstlisting}
+\caption{In this program, the programmer may have intended the placeholder
+  assignment \texttt{x = 1} to be used in case the user input 0, and used the
+  incorrect condition \texttt{x != 0}, rather than \texttt{x == 0}. However, the
+  condition expression will not be included in the slice (which would include
+  lines 1 and 4), because the presence of
+  the conditional does not affect the behavior of the program.
+}
+\label{slice-downside-example}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}
+while true:
+  x = input()
+  if x != 0:
+    break
+  print("One divided by your number is: %d" % (1 / x))
+\end{lstlisting}
+\caption{
+  This program's mistake is the same as before,
+  but instead they try to handle the
+  bad input by breaking out of the input loop. When computing the slice, our
+  method would record a control dependency between the condition on line 3 and
+  the print statement on line 5, placing the bad condition in the program slice.
+}
+\label{early-break}
+\end{figure}
+
+The goal of this feature, roughly equivalent in purpose to the type error slice in
+\cite {learning-to-blame}, is to help eliminate terms that cannot be the source of
+the crash. We compute a dynamic program slice \cite{KOREL1988155}, the set of
+terms that contributed at runtime to the observed exception, and turn this
+feature on for terms in the slice. We follow the basic approach of
+\cite{KOREL1988155, KOREL1990187}, building a graph of data and control
+dependencies. We then traverse the graph backwards, starting at the execution
+step where the error occurred, to collect the set of terms that the excepting
+line transitively depended on. This excludes lines that could not have caused the
+exception, such as lines that never ran, or lines that ran but had no connection
+to the step where the exception happened.
+
+A downside of using the dynamic slice, in the case of control dependencies, is
+that it excludes terms whose presence does not affect whether the exception
+happened, but which could have had an effect if the term was different. In
+\figref{slice-downside-example}, the conditional statement is never run, so the
+removal of the conditional would not have any effect on the program's behavior,
+and it would not be included in the program slice. The dynamic slice misses the
+incorrect test condition because the mistake itself caused the conditional
+statement and its body to become irrelevant.
+
+To overcome this issues in the ``early return'' or ``early break'' case
+(\figref{early-break}), we check if a break, return, or other statement for
+breaking out of structured control flow is present inside of a conditional
+statement. We then add dependencies in the execution trace's dependency graph
+between the enclosing conditional and the statements that would have been skipped
+by the break or return. This is insufficient to overcome all related problems,
+however, so unlike in prior work, we do not treat membership in the slice
 as a hard constraint.
@@
 \bibliographystyle{abbrv}
-\bibliography{nanomaly,sw,temp}
+\bibliography{nanomaly,sw,temp,slice}
 

### 2018-08-20 paper/main.tex
paper updates

@@
 
+\subsection{Label}
+
+The label we would ideally like to use would be whether or not a term is the
+true source of the crash; that is, whether the correct way to fix the crash is
+to edit this term. However, manually labeling terms in such a way would be
+prohibitive on such a large dataset. And the whole point of this work is that
+there does not yet exist a reliable way of pinpointing bugs automatically. Thus
+we are forced to use a proxy for our desired label as follows. Many PythonTutor
+users use the site in an iterative fashion: they start out by writing
+a program that crashes, and then edit it until it no longer crashes. Thus for our
+dataset, we consider only the crashing programs for which the same user later
+ran a program that did not crash. We then run a tree-diff \cite{tree-diff}
+between the program that crashes and the one that does not. Any terms in the
+crashing program that appear in the diff are labeled as bugs. This is of course
+not a perfect proxy for the label we want; see Threats to Validity.
+
 \subsection{Machine learning}
@@
 
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[
+    ybar,
+    symbolic x coords={Top-1,Top-3},
+    xtick=data,
+    ymin=0,
+    ylabel=Model Accuracy,
+    enlarge x limits=0.5,
+    legend style={at={(0.5,-0.15)},anchor=north}
+]
+    \addplot table[x=scoreName, y=All, col sep=comma]{defect-categories.csv};
+    \addplot table[x=scoreName, y=TypeError, col sep=comma]{defect-categories.csv};
+    \addplot table[x=scoreName, y=AttributeError, col sep=comma]{defect-categories.csv};
+\legend{All Errors,TypeError,AttributeError}
+\end{axis}
+\end{tikzpicture}
+\caption{Accuracy when using only a subset of the data that all has the same crash. TODO:
+get separate baselines?}
+\label{fig:defect-categories}
+\end{figure}
+
 prereq: define defect category. This could be either the raw python

### 2018-08-21 paper/main.tex
paper updates

@@
 buggy program by a Bag of Abstracted Terms. In our setting, a `term' means either
-a statment or an expression. For each term in the program we compute the following features:
+a statment or an expression.
+
+For each term in the program we compute static, dynamic,
+and contextual features. While it is easy to read some features off of the surface
+syntax, we expect that some features which are only accessible dynamically will also
+be useful, for example the type of each expression. Since programs are highly
+structured and expressions gain meaning in relation to their surroundings,
+we also expect contextual features to be useful.
+
+\subsection{Static features}
+
+\emph{Syntactic form} We expect certain kinds of terms to be more prone to bugs
+than others - for example, students might have more trouble with loop conditions
+than with simple assignments. Thus the first feature we consider is what kind of node the term
+is in the AST. For statements this has values like Return or Import; for expressions
+Variable or Application.
+
+\emph{Size} We also count the number of descendent nodes in the subtree rooted at that node.
 
@@
 
-\emph{Type} The possible values for the Type feature include all the basic Python
+\emph{Type} An expression's type should be a very useful feature. Not only are
+incompatible types a sign of an impending crash, but some types may be inherently
+suspect whenever they appear, e.g. there are very few reasons to have a variable
+of type NoneType. Our possible values for the Type feature include all the basic Python
 types (such as int and tuple) as well as a few special values:
@@
 relevant features filled out with a special NotApplicable value.
-
-\subsection{Static features}
-
-\emph{Syntactic form} The first feature we consider is what kind of node the term
-is in the AST. For statements this has values like Return or Import; for expressions
-Variable or Application.
-
-\emph{Size} We also count the number of descendent nodes in the subtree rooted at that node.
 
@@
     ybar,
-    symbolic x coords={Top-1,Top-3},
+    symbolic x coords={Baseline,Top-1,Top-3},
     xtick=data,
@@
 ]
-    \addplot table[x=scoreName, y=withEverything, col sep=comma]{fault-localization.csv};
-    \addplot table[x=scoreName, y=baseline, col sep=comma]{fault-localization.csv};
-\legend{Our best model,Python interpreter}
+    \addplot table[x=scoreName, y=score, col sep=comma]{fault-localization.csv};
 \end{axis}
 \end{tikzpicture}
-\caption{Based on a decision tree on the largest
-dataset (272534 pairs). TODO: also train for Top-2}
+\caption{Baseline is the normal Python interpreter. Other bars are from decision
+trees on the largest dataset (272534 pairs). TODO: also train for Top-2}
 \label{fig:full-dataset-acc}
@@
     ybar,
-    symbolic x coords={Top-1,Top-3},
+    symbolic x coords={noContext, noTypes, all},
     xtick=data,
@@
 ]
-    \addplot table[x=scoreName, y=all, col sep=comma]{removing-features.csv};
-    \addplot table[x=scoreName, y=noTypes, col sep=comma]{removing-features.csv};
-    \addplot table[x=scoreName, y=noContext, col sep=comma]{removing-features.csv};
-    \addplot table[x=scoreName, y=noSlice, col sep=comma]{removing-features.csv};
-\legend{All Features,No Types, No Context, No Slicing}
+    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2.csv};
 \end{axis}
@@
 \caption{Accuracy when features are removed, based on decision trees on a random subset of 20000 pairs.
-TODO: train better model and/or
-train on greater fraction of the data? TODO: Top-2. NOTE: removing slicing makes basically no difference...}
+TODO: Top-2}
 \label{fig:removing-features}
@@
     ybar,
-    symbolic x coords={Top-1,Top-3},
+    symbolic x coords={All, TypeError, AttributeError},
     xtick=data,
@@
 ]
-    \addplot table[x=scoreName, y=All, col sep=comma]{defect-categories.csv};
-    \addplot table[x=scoreName, y=TypeError, col sep=comma]{defect-categories.csv};
-    \addplot table[x=scoreName, y=AttributeError, col sep=comma]{defect-categories.csv};
-\legend{All Errors,TypeError,AttributeError}
+    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2.csv};
 \end{axis}
 \end{tikzpicture}
-\caption{Accuracy when using only a subset of the data that all has the same crash. TODO:
-get separate baselines?}
+\caption{Accuracy when the model is trained and tested on only the data exhibiting
+the error on the x-axis.}
 \label{fig:defect-categories}
@@
 
+% \begin{figure}
+% \foreach \method in {single, complete, average}
+% {
+% \begin{tikzpicture}
+%   \begin{axis}[
+%       xlabel=Threshold,
+%       ylabel=\# of clusters,
+%     %   ytick distance=2,
+%       title=Linkage: \method
+%     ]
+%     \addplot+[mark=none] table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+% \end{axis}
+% \end{tikzpicture}
+% }
+% \caption{The number of clusters when clustering a baseline of twenty dissimilar programs from
+%   Rosetta Code. We give no plot for the baseline of 5000 identical programs, as they were
+%   consistently placed in just 1 cluster.}
+% \label{fig:diversity-baseline}
+% \end{figure}
+
+% \begin{figure}
+% \foreach \method in {single, complete, average}
+% {
+% \begin{tikzpicture}
+%   \begin{axis}[
+%       xlabel=Threshold,
+%       ylabel=\# of clusters,
+%     %   ytick distance=1000,
+%       title=Linkage: \method
+%     ]
+%     \foreach \language in {ocaml, python}
+%     {
+%       \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+%     }
+% \legend{OCaml, Python}
+% \end{axis}
+% \end{tikzpicture}
+% }
+% \caption{Number of clusters when data is clustered by single, complete, and average linkage.
+%   A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
+%   into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
+%   of a link is its distance, z-scored against other nearby links.}
+% \label{fig:diversity}
+% \end{figure}
+
 \begin{figure}
-\foreach \method in {single, complete, average}
-{
 \begin{tikzpicture}
   \begin{axis}[
-      xlabel=Threshold,
-      ylabel=\# of clusters,
-    %   ytick distance=2,
-      title=Linkage: \method
-    ]
-    \addplot+[mark=none] table[x=threshold, y=python_twenty_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
-\end{axis}
-\end{tikzpicture}
-}
-\caption{The number of clusters when clustering a baseline of twenty dissimilar programs from
-  Rosetta Code. We give no plot for the baseline of 5000 identical programs, as they were
-  consistently placed in just 1 cluster.}
-\label{fig:diversity-baseline}
-\end{figure}
-
-\begin{figure}
-\foreach \method in {single, complete, average}
-{
-\begin{tikzpicture}
-  \begin{axis}[
-      xlabel=Threshold,
-      ylabel=\# of clusters,
+      xlabel=Clustering threshold parameter,
+      ylabel=\# of naturally occuring program clusters,
+      xmax=1.25,
+      legend style={at={(0.5,0.25)},anchor=north}
     %   ytick distance=1000,
-      title=Linkage: \method
+    %   title=Linkage: single
     ]
@@
     {
-      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_\method, col sep=comma]{cluster_counts.csv};
+      \addplot+[mark=none] table[x=threshold, y=\language_cluster_counts_single, col sep=comma]{cluster_counts.csv};
     }
-\legend{OCaml, Python}
+\legend{NATE OCaml data, Our Python data}
 \end{axis}
 \end{tikzpicture}
-}
-\caption{Number of clusters when data is clustered by single, complete, and average linkage.
+\caption{Number of clusters when data is clustered by single linkage.
   A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
   into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
-  of a link is its distance, z-scored against other nearby links.}
+  of a link is its distance, z-scored against other nearby links. TODO: find a way to include
+  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.}
 \label{fig:diversity}

### 2018-08-21 paper/main.tex
Wes pass over introduction

@@
 
-\emph{Placeholder Outline.} This text is an informal placeholder outline
-only.
-
-\emph{Identify an important problem.} Support the claim that Python fault
-localization for possibly-ill-typed student fragments is an important
-problem. Likely steps: fault localization is important, Python is
-important, incomplete code is important. First, software engineering is
-important and expensive; testing, maintenance, and debugging are the
-dominant activities in SE; localization and triage are key steps. Second,
-Python is increasingly used for pedagogy and industrial deployment, here
-are some GitHub numbers about the rate of growth in Python projects
-compared to C or Java, etc. Third, IDE tool use requires operating on code
-fragments, here are some numbers from PythonTutor~\cite{Guo2013-vu}, generality, etc.
-
-\emph{Identify the properties of a good solution.} Pick your favorite
-three. Some candidates: accuracy (cite Parnin/Orso paper an indicate that
-it must be found within top three, etc., since humans will not read long
-lists), generality (must apply to student code and incomplete fragments),
-scalability (must apply to hundreds of thousands of fragments, must present
-real-time responses suitable for web/IDE use), agreement with subjective
-judgments (must handle ``ambiguity'' the same way humans would when you
-don't know whether to blame the use or the definition, etc.), efficiency
-(must operate using a small number of dynamic runs, unlike Tarantula,
-etc.).
-
-\emph{Show that the current state of the art is inadequate.} For each $x$ of
-\{manual debugging, Python's default error messages,
-Mycroft/Sherrloc/Tarantula/whatever\}, show that $x$ lacks at least one of the
-desired properties listed above.
-
-\emph{Insights and special sauce.} What are the two or three good ideas
-that we bring together to solve this problem? Some candidates: the use of
-modern machine learning techniques (deep learning, etc.); the use of
-tree-based contextual features to hit the sweet spot between syntax and semantics;
-special handling of unsatisfied use-def type constraints (TBA); etc.
-
-\emph{What is our proposal?} We propose XYZ, an algorithm to do ABC. It
-takes as input DEF. It has stages P, Q and R. It produces as output GHI. It
-provides guarantees JKL.
-
-\emph{How will we evaluate it?} For each of the properties of a good
-solution listed above, indicate how we will evaluate our contribution. What
-evidence will we provide? Typically, for each property $x$, we support our
-claim to $x$ with either a formal proof (e.g., our algorithm has $x$ by
-construction), an empirical evaluation on software artifacts, or an
-empirical evaluation on human subjects. In each case, indicate both the
-evaluation metrics (e.g., if we care about efficiency, is it measured in
-lines-per-second? if we care about precision, what is our non-standard
-top-three metric) and success criteria (e.g., how will we define success?
-what is the baseline we want to beat). For empirical evaluations, briefly
-and implicitly argue that the results are likely to generalize by giving
-the size of the sampled set.
-
-\emph{Introduction wrap-up.} The contributions of this paper are as follows
-...
+Dynamically-typed languages are becoming increasingly common for rapid
+application development, full-scale software engineering, and
+pedagogy (e.g.,~\cite{FIXME,FIXME}). Languages like Python, Ruby and Lua
+often feature a gentle learning curve and attractive facilities, from
+familiar garbage collection and regular expressions to more expressive and
+exotic list comprehensions and higher-order function blocks. One side
+effect of popularity with novices and students, however, is that many
+programmers of such languages may lack expertise when presented with
+dynamic type errors, and may thus struggle to interpret and localize
+them~\cite{FIXME,FIXME}. We propose a machine learning approach based on
+static, dynamic and contextual information that effectively localizes
+beginner type errors.
+
+The relative costs and benefits of static vs. dynamic error detection are
+well-studied. Despite this, however, many currently-available techniques
+are ill-suited to address the problem of localization for type errors in
+dynamic languages. First, relying on the extant language interpreter
+requires test inputs~\cite{FIXME}, may involve error messages that are
+difficult for novices to understand~\cite{FIXME}, and may misleadingly
+implicate the symptom rather than the cause~\cite{FIXME}.  Second, a
+significant amount of development is carried out via IDE or web interfaces,
+which require tools that operate on code fragments~\cite{Guo2013-vu}.
+Third, attempts to retrofit static type systems into dynamic languages,
+such as TypeScript for JavaScript~\cite{FIXME} and FIXME for
+Python~\cite{FIXME} are promising but not yet as widely
+deployed~\cite{FIXME}. Finally, research tools and algorithms for fault
+localization either require rich static types (e.g., as in
+Mycroft~\cite{FIXME} or Sherrloc~\cite{FIXME}), require test cases (e.g.,
+as in Tarantula~\cite{FIXME}), or produce voluminous ranked output lists
+that have been found to be unhelpful in general and less useful to novices
+in particular~\cite[Sec.~5.1]{orso-parnin}.
+
+We focus on Python as an indicative, popular dynamically-typed language.
+We desire a fault localization algorithm for Python type errors that
+will agree with human intuition (\emph{accuracy}), will apply to
+off-the-shelf, unannotated beginner-written program fragments
+(\emph{generality}), and will scalably apply to industrial-strength
+deployments (\emph{scalability}). Our key insights are (1) that we can leverage
+modern supervised machine learning for accuracy and an agreement with human
+norms; (2) that we can use static, dynamic, contextual and slice-based
+features to generalize to off-the-shelf beginner programs; and (3) that we
+can use large beginner datasets, such as those from
+PythonTutor.com~\cite{Guo2012-vu}, to assess scalability.
+
+We propose an algorithm that takes an input a Python program fragment that 
+produces an uncaught runtime exception such as a type error. Via machine
+learning over a carefully-selected set of program-relevant features, we
+produce the expression most likely implicated in the exception according to
+our learned model. We evaluate on an entire year of buggy Python 3 program
+executions on PythonTutor.com --- over 270,000 instances --- measuring
+accuracy with respect to the human actions as the ground truth and
+comparing against the vanilla Python interpreter as a baseline.
+
+The contributions of this paper are as follows:
+\begin{enumerate}
+
+\item We present an algorithm for accurately localizing dynamic type errors
+in Python. We find that FIXME.
+
+\item We describe and evaluate static, dynamic, contextual and slice-based
+features for generalizing machine-learning fault localization to
+beginner-written Python fragments. We find that FIXME.
+
+\item We analyze the results of a thorough evaluation on over 270,000 real
+user instances, demonstrating the scalability of our algorithm.
+
+\end{enumerate}
 

### 2018-08-21 paper/main.tex
minor change to motex

@@
 \label{fig-motex}
-A program with multiple fault localizations (e.g.,
-lines 1 and 2).}
+A program with multiple fault localizations (lines 1 and 2).}
 \end{figure}
 
-Consider the program in \figref{motex}, adapted from our dataset of
-human submissions. The program attempts to carry out some simple arithmetic
-based on a given number and the current year. When executed, however, the
-program raises an exception on line 4, related to the addition of an
-integer variable to a string variable.
+One of the key difficulties in localizing type errors, especially for
+dynamically-typed programs, is that there are often multiple
+logically-valid ways to address a type error~\cite{FIXME} --- only one of
+which may be desired by the programmer~\cite{FIXME}.  Consider the program
+in \figref{motex}, adapted from our dataset of human submissions. The
+program attempts to carry out some simple arithmetic based on a given
+number and the current year. When executed, however, the program raises an
+exception on line 4, related to the addition of an integer variable to a
+string variable.
 

### 2018-08-21 paper/main.tex
approach section

@@
     T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+
+\newcommand{\fixme}[1]{\textcolor{red}{FIXME: #1}}
 
@@
 localization either require rich static types (e.g., as in
-Mycroft~\cite{FIXME} or Sherrloc~\cite{FIXME}), require test cases (e.g.,
-as in Tarantula~\cite{FIXME}), or produce voluminous ranked output lists
-that have been found to be unhelpful in general and less useful to novices
-in particular~\cite[Sec.~5.1]{orso-parnin}.
+Mycroft~\cite{FIXME}, Nate~\cite{FIXME} or Sherrloc~\cite{FIXME}), require
+test cases (e.g., as in Tarantula~\cite{FIXME}), or produce voluminous
+ranked output lists that have been found to be unhelpful in general and
+less useful to novices in particular~\cite[Sec.~5.1]{orso-parnin}.
 
@@
 \section{Motivating Example}
+\label{sec-motex}
 
@@
 
-\section{Overview and Approach}
+\section{Algorithm Overview}
 
 We present an algorithm for accurately localizing faults~\cite{tarantula} in
-dynamically-typed Python programs that exhibit non-trivial uncaught runtime
-exceptions. We do not consider syntax errors or references to undefined
-variables. Our algorithm uses machine learning models based on static and
-dynamic features to implicate suspicious expressions. Since studies have
-found that voluminous fault localization output is not useful to
-developers~\cite{orso-parnin,orso-parnin2015}, we focus on producing
-Top-1 and Top-3 rankings.
+dynamically-typed, beginner-written Python programs that exhibit
+non-trivial uncaught runtime exceptions. We do not consider syntax errors
+or references to undefined variables. Our algorithm uses machine learning
+models based on static, dynamic, contextual and slicing features to
+implicate suspicious expressions. Since studies have found that voluminous
+fault localization output is not useful to
+developers~\cite{orso-parnin,orso-parnin2015}, we focus on producing Top-1
+and Top-3 rankings.
 
@@
 
-\subsection{Feature Extraction}
+Drawing inspiration from localization algorithms such as
+Nate~\cite{learning-to-blame} and the natural language processing term
+frequency vector (or ``bag of words'') model~\cite{FIXME}, we represent
+each buggy program as a ``bag of abstracted terms''. A \emph{term} is
+either a statement of expression.
+
+\subsection{Model Feature Intuition and Extraction}
+
+We observe that many type errors admit multiple logically valid resolutions
+(see \secref{motex}): we thus cannot expect to match programmer desires
+through type constraints alone, and instead propose to leverage information
+from multiple modalities. We use static features to capture structured
+program meaning, contextual features to capture the relationship between a 
+program fragment and its environment, and dynamic features (including slice
+information) to reason about conditional behavior.
+
+For each term in the program we compute static, dynamic, and contextual
+features. While some features can be computed directly from syntactic
+information, others are only only accessible dynamically. In a
+dynamically-typed language such as Python, this includes types of each
+expression. Since programs are highly structured and expressions gain
+meaning in relation to their surroundings, we also include contextual
+features.
+
+While static features can be computed via parsing, dynamic features require
+executing the program fragment. To calculate dynamic features, we first
+convert each buggy program to an equivalent program that is closer to ANF.
+\fixme{WRW has no idea what ``ANF'' means in this context, and this is the
+first and only time it appears in the paper text. Someone else should
+rewrite this.} We define ANF to be FIXME and ``closer'' to mean FIXME in
+this context. We then run the resulting program through the PythonTutor
+backend~\cite{FIXME}.  This backend is based on the Python debigging
+library BDB~\cite{FIXME}; it returns a trace of the program along with the
+state of the heap at each execution step. That trace is used to compute
+dynamic features.
+
+\subsection{Model Features}
 \label{sec-features}
 
-Our key insight is that FIXME-INSIGHT. Are a result, we make use of
-static, dynamic and contextual information. Static features, such as FIXME,
-are effective at capturing FIXME-INFORMATION. By contrast, dynamic
-features, such as FIXME, support the modeling of FIXME. Finally, we also
-include contextual information, such as FIXME, which allows our algorithm
-to accurately handle FIXME. We consider the empirical justification of
-these features in Section~\ref{sec-eval}.
-
-We calculate features for each statement and expression in the program.
-
+\subsubsection{Syntactic Forms (Static)}
+
+We hypothesize that certain syntactic categories of terms may be more
+prone to bugs than others, especially for beginner programmers.  For
+example, students might have more trouble with loop conditions than with
+simple assignments. The first feature we consider is the syntactic
+category of node. This feature is categorical, using standard abstract
+syntax tree categories such as Return or Import for statements and Variable 
+or Application for expressions.
+
+\subsubsection{Expression Size (Static)}
+
+This numeric feature counts the number of descendent nodes in the subtree
+rooted at that node. Our intuition is that larger, more complex expressions
+may be more prone to faults.
+
+\subsubsection{Type (Dynamic)}
+
+Although type constraints alone cannot always match human intuition for 
+fault localization problems, types still contain very useful information.
+We observe that some types may be inherently suspect whenever they appear,
+especially for beginner-written code. For example, there are very few
+reasons to have a variable of type \lt{NoneType} in Python. We model
+type information as a categorical feature. Possible
+values include all the basic Python types (such as \lt{int}
+and \lt{tuple}) as well as three special values:
 \begin{itemize}
-
-\item \emph{Static / syntactic features}
-\begin{enumerate}
-    \item What kind of statement/expression it is, e.g. assignment / return
-     / import (for statements) or variable / literal / application (expressions).
-    \item Size (the number of AST nodes in the subtree rooted at this node)
-\end{enumerate}
-
-\item \emph{Dynamic features}
-We run each program through an instrumented interpreter (i.e. a modified PythonTutor
-backend, which itself is based on the BDB debugging library). This lets us
-compute the following features:
-\begin{enumerate}
-    \item What type(s) does the expression have at runtime? (The values of this feature
-    can be a type like int, or the special values "unknown" if the expression is
-    never successfully evaluated and "multiple" if it changes type.)
-    \item Is it part of the error slice?
-    \item Is it the node at which the program actually crashes?
-\end{enumerate}
-(Since this interpreter works at the granularity of lines and we want to work with
-expressions as well, we first convert each program to ANF.)
-
-We compute the error slice by running the program and building a dependency
-graph where node A depends on node B if either B defines a variable that A uses,
-or B affects control flow allowing A to run.
-
-We also check what uncaught exception is thrown by the program (TypeError,
-KeyError, etc), and add that as a constant feature to all vectors derived from that program.
-
-\item \emph{Contextual features} After all other features are computed so we have a
-vector $v_e$ representing expression $e$, we set $v_e^{+context} = v_e \circ
-v_p \circ v_{c_1} \circ v_{c_2} \circ v_{c_3}$ where $p$ is the expression's parent
-in the AST and $c_i$ are its children. This allows us to recover some of the
-program structure that would otherwise be completely lost when we convert the
-structured program into an unstructured bag of feature vectors.
-
-\item \emph{Label} Each feature vector is labeled by whether that expression/statement
-changed between the crashing and fixed versions. (This is our proxy for whether
-that node is to blame for the crash).
-
+    \item \lt{Statement} --- Given to all statements, since our terms
+    include both statments and expressions but only expressions 
+    have types
+    \item \lt{Unknown} --- Given to expressions that are never evaluated in the dynamic
+    trace
+    \item \lt{Multiple} --- Given to expressions which are evaluated multiple times in
+    a trace and do not always have the same type
 \end{itemize}
 
-Each vector is then one-hot encoded (e.g. feature "Type : Int" becomes set
-of features "Type-Int : 1", "Type-Bool : 0", "Type-Str : 0", etc.) for use with
-standard machine learning tools.
-
-\subsection{Machine Learning and Modeling}
-\label{sec-model}
-
-The models we train (inc. decision trees/forests,
-MLP, GBM) can then rank expressions of a test program in order of how likely they
-are to be the source of the bug, so we can compute Top-k accuracies: how often
-is the true program change located in the model's first k outputs.
-
-\section{Approach, in more detail}
-
-We follow the overall strategy of \cite{learning-to-blame} and represent each
-buggy program by a Bag of Abstracted Terms. In our setting, a `term' means either
-a statment or an expression.
-
-For each term in the program we compute static, dynamic,
-and contextual features. While it is easy to read some features off of the surface
-syntax, we expect that some features which are only accessible dynamically will also
-be useful, for example the type of each expression. Since programs are highly
-structured and expressions gain meaning in relation to their surroundings,
-we also expect contextual features to be useful.
-
-\subsection{Static features}
-
-\emph{Syntactic form} We expect certain kinds of terms to be more prone to bugs
-than others - for example, students might have more trouble with loop conditions
-than with simple assignments. Thus the first feature we consider is what kind of node the term
-is in the AST. For statements this has values like Return or Import; for expressions
-Variable or Application.
-
-\emph{Size} We also count the number of descendent nodes in the subtree rooted at that node.
-
-\subsection{Dynamic features}
-
-We first convert each buggy program to an equivalent program that is closer
-to ANF, then run it through the PythonTutor backend. This backend is based on the
-Python debigging library BDB; it returns to us a trace of the program along with
-the state of the heap at each execution step. We use this information to compute
-the following features for each term:
-
-\emph{Type} An expression's type should be a very useful feature. Not only are
-incompatible types a sign of an impending crash, but some types may be inherently
-suspect whenever they appear, e.g. there are very few reasons to have a variable
-of type NoneType. Our possible values for the Type feature include all the basic Python
-types (such as int and tuple) as well as a few special values:
-\begin{itemize}
-    \item IsStatement - Given to all statements, since our terms include both
-    statments and expressions but only expressions actually have types.
-    \item Unknown - Given to expressions that are never evaluated in the dynamic
-    trace.
-    \item Multiple - Given to expressions which are evaluated multiple times in
-    a trace and do not always have the same type.
-\end{itemize}
-
-\emph{Slice}
+\subsubsection{Slice (Dynamic)}
 \begin{figure}
@@
   x = 1
-print("One divided by your number is: %d" % (1 / x))
+print("One over your number is: %d" % (1 / x))
 \end{lstlisting}
 \caption{In this program, the programmer may have intended the placeholder
-  assignment \texttt{x = 1} to be used in case the user input 0, and used the
-  incorrect condition \texttt{x != 0}, rather than \texttt{x == 0}. However, the
+  assignment \lt{x = 1} to be used in case the user input 0, and used the
+  incorrect condition \lt{x != 0}, rather than \lt{x == 0}. However, the
   condition expression will not be included in the slice (which would include
@@
     break
-  print("One divided by your number is: %d" % (1 / x))
+  print("One over your number is: %d" % (1 / x))
 \end{lstlisting}
 \caption{
-  This program's mistake is the same as before,
-  but instead they try to handle the
-  bad input by breaking out of the input loop. When computing the slice, our
-  method would record a control dependency between the condition on line 3 and
-  the print statement on line 5, placing the bad condition in the program slice.
+  In this example, the programmer attempts to handle invalid input by
+  breaking out of the input loop. When computing the slice, our method
+  would record a control dependency between the condition on line 3 and the
+  print statement on line 5, placing the bad condition in the program
+  slice.
 }
@@
 
-The goal of this feature, roughly equivalent in purpose to the type error slice in
-\cite {learning-to-blame}, is to help eliminate terms that cannot be the source of
-the crash. We compute a dynamic program slice \cite{KOREL1988155}, the set of
-terms that contributed at runtime to the observed exception, and turn this
-feature on for terms in the slice. We follow the basic approach of
-\cite{KOREL1988155, KOREL1990187}, building a graph of data and control
-dependencies. We then traverse the graph backwards, starting at the execution
-step where the error occurred, to collect the set of terms that the excepting
-line transitively depended on. This excludes lines that could not have caused the
-exception, such as lines that never ran, or lines that ran but had no connection
-to the step where the exception happened.
+The goal of this feature, roughly equivalent in purpose to the type error
+slice in Nate~\cite{learning-to-blame}, is to help eliminate terms that
+cannot be the source of the crash. We compute a dynamic program slice
+\cite{KOREL1988155}, the set of terms that contributed at runtime to the
+observed exception, and turn this feature on for terms in the slice. We
+follow the basic approach of \cite{KOREL1988155, KOREL1990187}, building a
+graph of data and control dependencies. We then traverse the graph
+backwards, starting at the execution step where the error occurred, to
+collect the set of terms that the excepting line transitively depended on.
+This excludes lines that could not have caused the exception, such as lines
+that never ran, or lines that ran but had no connection to the step where
+the exception happened.
 
@@
 happened, but which could have had an effect if the term was different. In
-\figref{slice-downside-example}, the conditional statement is never run, so the
-removal of the conditional would not have any effect on the program's behavior,
-and it would not be included in the program slice. The dynamic slice misses the
-incorrect test condition because the mistake itself caused the conditional
-statement and its body to become irrelevant.
+\figref{slice-downside-example}, the conditional statement is never
+executed, so the removal of the conditional would not have any effect on
+the program's behavior, and it would not be included in the program slice.
+The dynamic slice misses the incorrect test condition because the mistake
+itself caused the conditional statement and its body to become irrelevant.
 
@@
 (\figref{early-break}), we check if a break, return, or other statement for
-breaking out of structured control flow is present inside of a conditional
+escaping structured control flow is present inside of a conditional
 statement. We then add dependencies in the execution trace's dependency graph
 between the enclosing conditional and the statements that would have been skipped
-by the break or return. This is insufficient to overcome all related problems,
-however, so unlike in prior work, we do not treat membership in the slice
-as a hard constraint.
-
-\emph{Crash location} As discussed later in the evaluation, the precise term that
-causes Python to crash is in fact the one the user decides to change about 30\% of
-the time, so we use it as a feature (and expect it to be a high-signal one, too.)
-
-\emph{Exception type} The type of error thrown by the program is useful context for
-finding the error. For example, if asked how likely a random `0` is of being a bug, you'd give
-a higher answer if you knew that the program eventually crashes with a division by
-zero exception. Thus we attach the exception type as a feature to \emph{every}
-vector derived from that program.
-
-\subsection{Contextual features}
-
-The approach of representing a program by a Bag of Abstracted Terms, if implemented
-naively, runs into the problem that the terms lose their context in the program
-as a whole. For example, the features for the term `0` include its type (int) and
-syntactic description (int literal), but there is no way that's enough to decide
-if it's innocuous, like the 0 in `x = 0`, or clearly the source of a bug, like the
-0 in `x / 0`. To restore some of this context, we introduce contextual features.
-This involves duplicating all our other features four times - once for the parent
-and once each for a term's first three children in the AST. So since `x / 0` has
-the feature SyntacticForm = DivisionOp, the `0` of `x / 0` has the feature
-ParentSyntacticForm = DivisionOp. Similarly, since `0` has SyntacticForm = IntLiteral,
-`x / 0` has Child1SyntacticForm = IntLiteral. In order to make all feature vectors
-the same length, nodes that have no parent or fewer than three children have the
-relevant features filled out with a special NotApplicable value.
-
-\subsection{Label}
-
-The label we would ideally like to use would be whether or not a term is the
-true source of the crash; that is, whether the correct way to fix the crash is
-to edit this term. However, manually labeling terms in such a way would be
-prohibitive on such a large dataset. And the whole point of this work is that
-there does not yet exist a reliable way of pinpointing bugs automatically. Thus
-we are forced to use a proxy for our desired label as follows. Many PythonTutor
-users use the site in an iterative fashion: they start out by writing
-a program that crashes, and then edit it until it no longer crashes. Thus for our
-dataset, we consider only the crashing programs for which the same user later
-ran a program that did not crash. We then run a tree-diff \cite{tree-diff}
-between the program that crashes and the one that does not. Any terms in the
-crashing program that appear in the diff are labeled as bugs. This is of course
-not a perfect proxy for the label we want; see Threats to Validity.
-
-\subsection{Machine learning}
+by the break or return. While this heuristic is effective in practice, it
+does not overcome all related problems: we thus treat dynamic slice
+information as one of many features rather than as a hard constraint.
+This boolean feature tracks whether or not the term is part of the final
+dynamic slice.
+
+\subsubsection{Crash Location (Dynamic)} We observe that the precise term that
+causes Python to crash is, in fact, the one the user decides to change
+much of the time. While this information is not sufficient on its own
+(cf. the vanilla Python interpreter, which largely uses this feature), 
+it can be a very helpful supplement. This boolean feature encodes whether
+or not the term is exactly the source location of the uncaught exception.
+
+\subsubsection{Exception Type (Dynamic)} The type of error thrown by the program
+contains useful information its localization. In particular, a term may be
+more or less suspicious depending on the ultimate uncaught exception. As an
+example, a \lt{v / 0} term may be more suspicious if the uncaught exception
+being considered is \lt{DivisionByZero}. We encode the exception type as
+as a categorical feature. 
+% to \emph{every} vector derived from that program.
+
+\fixme{The original draft contained this text: ``For example, if asked how
+likely a random `0` is of being a bug, you'd give a higher answer if you
+knew that the program eventually crashes with a division by zero exception.''
+This sort of rhetorical communication with the reader is \emph{too
+informal for scientific writing}. I am not saying that I disagree with the
+underlying argument. I am saying that you cannot really use the Socratic
+method or the ``Appeal to Common Practice'' logical fallacy or the ``No
+True Scotsman'' logical fallacy in scientific writing. If you wrote the
+quoted text, read this FIXME carefully, ask me if you have any questions
+about sloppy vs. formal arguments, and then delete this FIXME. --- Wes}
+
+\subsubsection{Contextual features}
+
+Term frequency vectors in general, and our bag of abstracted terms
+representation by extension, do not typically include contextual
+information such as ordering or relative positioning. As with many other
+analyses, such as contextual operational semantics~\cite{FIXME}, we observe 
+that the meaning of an expression may depend on its surrounding context.
+
+As an example, consider the \lt{0} terms in both \lt{x = 0} and \lt{ x/ 0}.
+Both instances of \lt{0} have the same syntactic form, size and type
+(etc.). We may prefer to implicate the \lt{0} in \lt{x / 0} as suspicious,
+especially for beginner-written programs, but cannot distinguish it without
+surrounding contextual information. Data structures such as abstract syntax
+trees and control flow graphics capture such contextual information, but
+are not immediately applicable to machine learning.
+
+We desire to encode such information while retaining the use of scalable,
+accurate off-the-shelf machine learning algorithms that operate on feature
+vectors. We thus propose to embed contextual information in a vector,
+borrowing insights from standard approaches in machine learning. We
+associated with term additional features that correspond to the features of
+its parent and child nodes. For representational regularity, we always
+model three children; terms without parents or children are given special
+values (e.g., zero or NotApplicable) for those contextual features.
+
+\subsection{Machine Learning Model}
+\label{sec-model}
 
@@
 \end{enumerate}
+
+\subsection{Label}
+
+The label we would ideally like to use would be whether or not a term is the
+true source of the crash; that is, whether the correct way to fix the crash is
+to edit this term. However, manually labeling terms in such a way would be
+prohibitive on such a large dataset. And the whole point of this work is that
+there does not yet exist a reliable way of pinpointing bugs automatically. Thus
+we are forced to use a proxy for our desired label as follows. Many PythonTutor
+users use the site in an iterative fashion: they start out by writing
+a program that crashes, and then edit it until it no longer crashes. Thus for our
+dataset, we consider only the crashing programs for which the same user later
+ran a program that did not crash. We then run a tree-diff \cite{tree-diff}
+between the program that crashes and the one that does not. Any terms in the
+crashing program that appear in the diff are labeled as bugs. This is of course
+not a perfect proxy for the label we want; see Threats to Validity.
 

### 2018-08-21 paper/main.tex
done with Wes pass for Tuesday

@@
 
+\pagestyle{plain}
+\pagenumbering{arabic}
+
 \begin{document}
@@
 
-\subsection{Machine Learning Model}
+\subsection{Machine Learning Model Generation}
 \label{sec-model}
 
-\emph{Pairing programs} For each PythonTutor user, we look at the sequence of
-programs they attempted to run. Each program that crashes is paired with the next
-program by that user which does not crash. Many programs do not make it to the
-end of our data pipeline; here are the most common reasons:
-\begin{itemize}
-    \item The user submitted only programs that crash or only programs that do
-    not crash, so we can't create any pairs for that user.
-    \item The program is out of scope for PythonTutor. For
-    example, the program imports forbidden libraries (like os), or the program
-    runs in over 1000 execution steps.
-    \item The proportion of the program that has changed between the buggy and
-    fixed versions is more than a standard deviation above the mean.
-\end{itemize}
-We end up with \textasciitilde270,000 usable pairs of programs.
-
+\fixme{
 \emph{Models} The following descriptions of Logistic Regression, Decision Trees,
 and MLP will have to be redone as they are copied from Learning to Blame.
+}
 
@@
 
-\item[RQ5]{FIXME clustering?}
-
 \end{enumerate}
 
-\subsection{Label}
-
-The label we would ideally like to use would be whether or not a term is the
-true source of the crash; that is, whether the correct way to fix the crash is
-to edit this term. However, manually labeling terms in such a way would be
-prohibitive on such a large dataset. And the whole point of this work is that
-there does not yet exist a reliable way of pinpointing bugs automatically. Thus
-we are forced to use a proxy for our desired label as follows. Many PythonTutor
-users use the site in an iterative fashion: they start out by writing
-a program that crashes, and then edit it until it no longer crashes. Thus for our
-dataset, we consider only the crashing programs for which the same user later
-ran a program that did not crash. We then run a tree-diff \cite{tree-diff}
-between the program that crashes and the one that does not. Any terms in the
-crashing program that appear in the diff are labeled as bugs. This is of course
-not a perfect proxy for the label we want; see Threats to Validity.
-
-\subsection{Data Set}
+\subsection{Dataset and Program Collection}
 
@@
 assumption (i.e., the program that does not crash is unrelated to the
-crashing program).
+crashing program). We also discard submissions that are out of scope given
+PythonTutor's policies (e.g., such as very long-running executions or the
+use of forbidden libraries).
+
+Ultimately, the dataset used in this evaluation contained just over
+270,000 usable pairs of programs.
+
+\subsection{Labeled Training and Ground Truth}
+
+Our algorithm is based on supervised machine learning and thus requires
+labeled training instances. In addition, our evaluations require a ground
+truth notion of which terms correspond to correct fault localizations.
+
+We use the terms changed involved in the fixes by actual users as our
+ground truth and training information.  Many PythonTutor users use the site
+in an iterative fashion: they start out by writing a program that crashes,
+and then edit it until it no longer crashes. Our dataset contains only only
+those crashing programs for which the same user later submitted a program
+that did not crash. We compute a tree-diff \cite{tree-diff} between the
+originaly, buggy submission and the first fixed submission. We define 
+the ground truth correct answer to be the set of terms in the crashing
+program that also appear in the diff. We discuss the implications of this
+choice in \secref{threats}.
 
@@
 \subsection{Threats to validity}
+\label{sec-threats}
 
@@
 \section{Related Work}
-FIXME: This must all be rewritten or dropped, as it is directly copied from
+\fixme{This must all be rewritten or dropped, as it is directly copied from
 my previous papers. It serves as a placeholder to indicate length and to
-remind us of some Software Engineering bits we might cite.
+remind us of some Software Engineering bits we might cite.}
 

### 2018-08-22 paper/main.tex
More slicing information

@@
 \subsubsection{Slice (Dynamic)}
+
+\begin{figure}
+\begin{lstlisting}
+x = input() # Input: 42
+if x < 20:
+  y = 5
+  z = 123
+else:
+  y = 0
+  z = 321
+assert(y != 0)
+\end{lstlisting}
+\begin{lstlisting}
+x = 42
+if x < 20:
+
+else:
+  y = 0
+
+assert(y != 0)
+\end{lstlisting}
+\caption{
+  a) A program being sliced. An assertion gets triggered when line 8 is run. b)
+  The lines included in the slice. This is not a valid python
+  program, but that does not pose a problem for using membership in the slice as
+  a learning feature.
+}
+\label{fig-slice-example}
+\end{figure}
+
 \begin{figure}
@@
 }
-\label{slice-downside-example}
+\label{fig-slice-downside-example}
 \end{figure}
@@
 }
-\label{early-break}
+\label{fig-early-break}
 \end{figure}
@@
 the exception happened.
+
+\figref{slice-example} is an example of the input and output of the slicing
+algorithm. The slicer takes as input a python source string and a list of inputs,
+to be fed into the program whenever \lt{input()} is invoked. The slicer outputs
+the list of line numbers for the lines in the slice.
+
+The slicer would first create an execution trace of the example program, which
+at each execution step includes the line number being run and the state of the
+heap and variables. In the example, lines $1, 2, 3, 4, 8$ would have been run.
+It then iterates forward through the trace, recording which
+execution steps have a define-use relation, and which execution steps
+have a test-control relation. The define-use relation is between a step where
+the value of a variable is used, and the step where that variable was last
+defined. In \figref{slice-example}, a define-use relationship exists
+between lines 1 and 2 and between 3 and 8. A test-control relation is between
+the execution of a control flow statement and any statements that were run as
+a result of it. In \figref{slice-example}, a test-control relationship exists
+between the if statement at line 2 and the assignments on lines 3 and 4. After
+building up these relations, the slicer starts at the execution step causing an
+exception (the execution of line 8), and builds the exception's set of
+dependencies by iteratively adding to the set any step that has a define-use
+or test-control relationship with a step already in the set. Finally, the
+algorithm takes all the execution steps, determines which line number each step
+was running, and then outputs this list of line numbers.
+
+Our slicing algorithm handles every syntax node supported by the Python3 standard
+library, \lt{ast}, except:
+\begin{itemize}
+\item Assignments where the left hand side is not a variable or a simple chain
+  of attribute indexing or subscripting
+\item Assignments where attribute indexing or subscripting on the left hand side
+  means something other than the default, which can happen if the operations are
+  overridden by a class
+\item Lambda expressions
+\item Set and dictionary comprehensions
+\item Generator expressions
+\item Await, yield, and yield from
+\item Ellipsis (the token representing the tail of the argument list given to a
+  varargs function)
+\item Starred expressions
+\item Async def functions and Asynchronous for
+\item Delete statements
+\item With statements
+\item Raise statements
+\end{itemize}
 

### 2018-08-22 paper/main.tex
more diversity info

@@
 
-In Nate there were 20 different functions. All the programs were doing the
-same 20 functions.
-
-Here, people are doing anything they want. Random stuff going on over here.
-
-We want to show that we get high accuracy *over a heterogenous set of
-programs*. Leon?
-
-prereq: cluster the programs based on similarity. Using agglomerative
-clustering plus python edit distance. How good the performance will be?
-right now it creates a clustering tree. binary tree, based on min distance.
-can give it a threshhold for distance (elements w/in a group). don't forget
-to rename all variables to X, perhaps normalize (edit distance size).
-
-leon: could also cluster by inconsistency coefficient
-
-PLOT Type: line graph (standard plot)
-
-PLOT X AXIS: clustering distance parameter
-
-PLOT Y AXIS: number of different clusters found
-
-PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''
-
-Goal: show that ``no matter how you decide program differences'', our
-dataset is always more diverse/heterogeneous than is previous work. So our
-success is amazing.
+%In Nate there were 20 different functions. All the programs were doing the
+%same 20 functions.
+%
+%Here, people are doing anything they want. Random stuff going on over here.
+%
+%We want to show that we get high accuracy *over a heterogenous set of
+%programs*. Leon?
+%
+%prereq: cluster the programs based on similarity. Using agglomerative
+%clustering plus python edit distance. How good the performance will be?
+%right now it creates a clustering tree. binary tree, based on min distance.
+%can give it a threshhold for distance (elements w/in a group). don't forget
+%to rename all variables to X, perhaps normalize (edit distance size).
+%
+%leon: could also cluster by inconsistency coefficient
+%
+%PLOT Type: line graph (standard plot)
+%
+%PLOT X AXIS: clustering distance parameter
+%
+%PLOT Y AXIS: number of different clusters found
+%
+%PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''
+%
+%Goal: show that ``no matter how you decide program differences'', our
+%dataset is always more diverse/heterogeneous than is previous work. So our
+%success is amazing.
 
@@
   of a link is its distance, z-scored against other nearby links. TODO: find a way to include
-  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.}
-\label{fig:diversity}
+  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.
+  LM: the y axis could be made into a percentage of the dataset size, but that might be
+  weird?}
+\label{fig-diversity}
 \end{figure}
+
+As a measure of the diversity of our dataset, we clustered the programs in
+our dataset and the NATE dataset, then compared the number of resultant clusters.
+To cluster programs, we transform the program ASTs into flat strings, and then
+perform agglomerative clustering on them, with string edit distance as a distance
+metric.
+
+\subsubsection{Transformation and Distance Metric}
+
+To measure the distance between programs, we flattened their ASTs into strings
+of tokens, and then computed the Levenshtein edit distance~\cite{levenshtein}.
+The decision to flatten the tree rather than compute a distance directly on the
+the ASTs is practical; the state of the art in tree edit distance has cubic
+asymptotic complexity~\cite{PAWLIK2016157}, while the levenshtein distance can be
+computed in time quadratic in the length of the strings~\cite{lev-quadratic}.
+When we attempted a test run using tree edit distance as the metric, it did not
+complete and we determined that it would not be practical to use.
+
+The flattening process, which is based on \cite{dist-site}, is defined
+recursively: the string for an AST
+begins with a unique token corresponding to the AST node type, which is
+followed by the concatenation of the transformations of the tree's subtrees.
+The string ends with an ``end'' token to denote the end of the tree. This
+preserves the tree structure of the AST.
+
+\subsubsection{Clustering Algorithm}
+
+We performed agglomerative clustering on the programs~\cite{modern-clustering}.
+This is a method of clustering in which every datapoint starts in its own
+cluster, and the two closest clusters are merged every step of the algorithm,
+until every datapoint is in the same cluster. This produces a binary tree, where
+each internal node corresponds to a cluster formed by merging two other clusters,
+and each leaf corresponds to an item in the dataset. We used single linkage,
+which means that the distance between clusters of multiple items was considered
+to be the minimum distance between their elements.
+
+This alone produces a tree rather than a list of disjoint clusters, which makes
+different clusterings difficult to compare. To produce a cluster count, we used
+the SciPy function \lt{scipy.cluster.hierarchy.fcluster}, which forms disjoint
+clusters by splitting the tree at internal nodes where the two joined clusters
+are more distant than some threshold. We used the \lt{'inconsistent'} method with
+depth set to 2, which specifies that the distances at each node should be
+z-scored against the distances at other nodes at the same level in the tree, and
+any nodes 1 or 2 levels below~\cite{inconsistent}. This adapts to differences in
+scale of edit distance that might arise between python and OCaml due to
+programming language expressivity or other factors, making cluster counts at each
+threshold value comparable.
+
+\figref{diversity} shows the results of clustering on our dataset and the NATE
+dataset. For all values of the inconsistency threshold, the algorithm computed a
+higher cluster count for the Python data, suggesting that it contains a more
+diverse set of programs than the OCaml dataset. The similarity in the graphs
+appears to confirm the comparibility of cluster counts between the two datasets.
 
@@
 \bibliographystyle{abbrv}
-\bibliography{nanomaly,sw,temp,slice}
+\bibliography{nanomaly,sw,temp,slice,cluster}
 

### 2018-08-22 paper/main.tex
paper updates

@@
 
-We propose an algorithm that takes an input a Python program fragment that 
+We propose an algorithm that takes as input a Python program fragment that
 produces an uncaught runtime exception such as a type error. Via machine
@@
 One reasonable fix (and the one that the programmer actually used in this
-case) is to add \lt{int()} cast on line 2. However, another possible
+case) is to add an \lt{int()} cast on line 2. However, another possible
 ``fix'' would be to \textit{remove} the \lt{int()} cast from line 1, in
@@
 
-Our algorithm first extracts static and dynamic features from Python
+Our algorithm first extracts static and dynamic features from a Python
 program (Section~\ref{sec-features}). Next, using a labeled training
@@
 each buggy program as a ``bag of abstracted terms''. A \emph{term} is
-either a statement of expression.
+either a statement or expression.
 
@@
 features. While some features can be computed directly from syntactic
-information, others are only only accessible dynamically. In a
+information, others are only accessible dynamically. In a
 dynamically-typed language such as Python, this includes types of each
@@
 executing the program fragment. To calculate dynamic features, we first
-convert each buggy program to an equivalent program that is closer to ANF.
-\fixme{WRW has no idea what ``ANF'' means in this context, and this is the
-first and only time it appears in the paper text. Someone else should
-rewrite this.} We define ANF to be FIXME and ``closer'' to mean FIXME in
-this context. We then run the resulting program through the PythonTutor
-backend~\cite{FIXME}.  This backend is based on the Python debigging
+convert each buggy program to an equivalent program in a form similar to
+A-normal form \cite{anf}: all intermediate expressions are split out and
+assigned to new temporary variables.
+We then run the resulting program through the PythonTutor
+backend~\cite{Guo2013-vu}.  This backend is based on the Python debugging
 library BDB~\cite{FIXME}; it returns a trace of the program along with the
@@
 
-To overcome this issues in the ``early return'' or ``early break'' case
+To overcome these issues in the ``early return'' or ``early break'' case
 (\figref{early-break}), we check if a break, return, or other statement for
@@
 \subsubsection{Exception Type (Dynamic)} The type of error thrown by the program
-contains useful information its localization. In particular, a term may be
+contains useful information for its localization. In particular, a term may be
 more or less suspicious depending on the ultimate uncaught exception. As an
-example, a \lt{v / 0} term may be more suspicious if the uncaught exception
+example, every division term may be more suspicious if the uncaught exception
 being considered is \lt{DivisionByZero}. We encode the exception type as
-as a categorical feature. 
+a categorical feature.
 % to \emph{every} vector derived from that program.
-
-\fixme{The original draft contained this text: ``For example, if asked how
-likely a random `0` is of being a bug, you'd give a higher answer if you
-knew that the program eventually crashes with a division by zero exception.''
-This sort of rhetorical communication with the reader is \emph{too
-informal for scientific writing}. I am not saying that I disagree with the
-underlying argument. I am saying that you cannot really use the Socratic
-method or the ``Appeal to Common Practice'' logical fallacy or the ``No
-True Scotsman'' logical fallacy in scientific writing. If you wrote the
-quoted text, read this FIXME carefully, ask me if you have any questions
-about sloppy vs. formal arguments, and then delete this FIXME. --- Wes}
 
@@
 
-As an example, consider the \lt{0} terms in both \lt{x = 0} and \lt{ x/ 0}.
+As an example, consider the \lt{0} terms in both \lt{x = 0} and \lt{x / 0}.
 Both instances of \lt{0} have the same syntactic form, size and type
@@
 borrowing insights from standard approaches in machine learning. We
-associated with term additional features that correspond to the features of
+associate with each term additional features that correspond to the features of
 its parent and child nodes. For representational regularity, we always
@@
 \emph{Models} The following descriptions of Logistic Regression, Decision Trees,
-and MLP will have to be redone as they are copied from Learning to Blame.
+Random Forests, and MLP will have to be redone as they are copied from Learning to Blame.
 }
@@
 well, though this can be somewhat mitigated by \emph{pruning} the tree.
+
+\emph{Random Forests}
+%
+Random forests improve generalization by training an
+\emph{ensemble} of distinct decision trees and using a majority
+vote to make a prediction.
+%
+The agreement among the trees forms a natural
+confidence score.
+%
+Since each classifier in the ensemble is a decision tree, this still
+allows for complex and expressive classifiers.
+
+The training process involves taking $N$ random subsets of the training
+data and training a separate decision tree on each subset --- the
+training process for the decision trees is often modified slightly to
+reduce correlation between trees, by forcing each tree to pick features
+from a random subset of all features at each node.
+%
+The diversity of the underlying models tends to make random forests less
+susceptible to the overfitting, but it
+also makes the learned model more difficult to interpret.
 
@@
 crashing program). We also discard submissions that are out of scope given
-PythonTutor's policies (e.g., such as very long-running executions or the
+PythonTutor's policies (e.g., very long-running executions or the
 use of forbidden libraries).
@@
 in an iterative fashion: they start out by writing a program that crashes,
-and then edit it until it no longer crashes. Our dataset contains only only
+and then edit it until it no longer crashes. Our dataset contains only
 those crashing programs for which the same user later submitted a program
@@
 
-For these evaluations, we employ machine learning algorithm FIXME, which
-provided the best balance of FIXME and FIXME (time? accuracy?) on our
-training set.
+For these evaluations, we employ decision trees as our machine learning
+algorithm, since they provided the best accuracy on our data given the
+available training time.
 
@@
 \subsection{RQ 1 --- Fault Localization Accuracy}
+
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[
+    ybar,
+    symbolic x coords={Baseline,Top-1,Top-2,Top-3},
+    xtick=data,
+    ymin=0,
+    ylabel=Model Accuracy,
+    enlarge x limits=0.3,
+    legend style={at={(0.5,-0.15)},anchor=north}
+]
+    \addplot table[x=scoreName, y=score, col sep=comma]{fault-localization.csv};
+\end{axis}
+\end{tikzpicture}
+\caption{Baseline is the normal Python interpreter. Other bars are from decision
+trees on the largest dataset (272534 pairs).}
+\label{fig-full-dataset-acc}
+\end{figure}
 
@@
     ylabel=Model Accuracy,
-    enlarge x limits=0.5,
+    enlarge x limits=0.3,
     legend style={at={(0.5,-0.15)},anchor=north}
 ]
-    \addplot table[x=scoreName, y=score, col sep=comma]{fault-localization.csv};
+    \addplot table[x=scoreName, y=score, col sep=comma]{fault-localization-random-forest.csv};
 \end{axis}
 \end{tikzpicture}
-\caption{Baseline is the normal Python interpreter. Other bars are from decision
-trees on the largest dataset (272534 pairs). TODO: also train for Top-2}
-\label{fig:full-dataset-acc}
+\caption{Baseline is the normal Python interpreter. Other bars are from random
+forests on the largest dataset (272534 pairs). TODO: get top-2, then change text
+to discuss this figure instead of decision-tree version}
+\label{fig-full-dataset-acc-random-forest}
 \end{figure}
 
-\begin{figure}
-\begin{tikzpicture}
-\begin{axis}[
-    ybar,
-    symbolic x coords={noContext, noTypes, all},
-    xtick=data,
-    ymin=0,
-    ylabel=Model Accuracy,
-    enlarge x limits=0.5,
-    legend style={at={(0.5,-0.15)},anchor=north}
-]
-    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2.csv};
-\end{axis}
-\end{tikzpicture}
-\caption{Accuracy when features are removed, based on decision trees on a random subset of 20000 pairs.
-TODO: Top-2}
-\label{fig:removing-features}
-\end{figure}
-
-prereq: define accuracy
-
-Goal: Establish that our algorithm works. This is the most basic question,
-but also the most relevant. If the reader takes away nothing else, they
-should conclude that our algorithm is effective.
-
-PLOT: See Figs. \ref{fig:full-dataset-acc} and \ref{fig:removing-features}
+We first use the full dataset to train and test decsion trees optimized for
+Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc},
+these produce a correct answer 55, 68, and
+75\% of the time (to two significant figures). As a baseline, we determine that
+the expression blamed by the standard Python interpreter (i.e. the one that
+causes it to crash) is only changed by the user 30\% of the time. Thus our
+most directly comparable model, the one for Top-1, significantly outperforms
+this baseline, and as we then exercise our power to give a ranked output rather
+than a single answer, our margin gets even better.
+
+% Goal: Establish that our algorithm works. This is the most basic question,
+% but also the most relevant. If the reader takes away nothing else, they
+% should conclude that our algorithm is effective.
 
@@
 \caption{Feature predictive power (for a Top-3 Decision Tree
-learned on the entire dataset).
-TODO: add
-better-known measure like ANOVA or ReliefF. TODO: base on Top-1 instead?}
+learned on the entire dataset).}
 \label{tab-feature-predictive-power}
 \end{table}
+
+\begin{table}[]
+\begin{tabular}{lll}
+Name & Category & F-score \\ \bottomrule
+Stmt-Constr-C3= & Contextual (Syntactic) & 69 \\
+Size & Syntactic                         & 63 \\
+PythonBlame & Dynamic (Error location)   & 50 \\
+Type-P=LIST & Contextual (Type)          & 36 \\
+Type-C1=bool & Contextual (Type)         & 30 \\
+Type-P=unknown & Contextual (Type)       & 28 \\
+Type-C1= & Contextual (Type)             & 25 \\
+Size-C3 & Contextual (Syntactic)         & 20 \\
+Op-Constr-C1= & Contextual (Syntactic)   & 20 \\
+Size-C1 & Contextual (Syntactic)         & 18 \\
+Type-P=DICT & Contextual (Type)          & 17 \\
+Type=str & Dynamic (Type)                & 12 \\
+\toprule
+\end{tabular}
+\caption{Feature predictive power as measured by ANOVA. F-score in
+thousands to two significant figures; all p-values less than .01}
+\label{tab-anova}
+\end{table}
+
+\begin{figure}
+\begin{tikzpicture}
+\begin{axis}[
+    ybar,
+    symbolic x coords={noContext, noTypes, all},
+    xtick=data,
+    ymin=0,
+    ylabel=Model Accuracy,
+    enlarge x limits=0.5,
+    legend style={at={(0.5,-0.15)},anchor=north}
+]
+    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2.csv};
+\end{axis}
+\end{tikzpicture}
+\caption{Accuracy when features are removed, based on decision trees on a
+random subset of 20000 program pairs.}
+\label{fig-removing-features}
+\end{figure}
 
@@
 
-prereq: define predictive power (leave-one-out, leave-one-in, Relief-F,
-ANOVA, whatever)
-
-PLOT: See Table \ref{tab:feature-predictive-power}
-
-Use as a rough guide: Table 3 on Page 8 of
-https://web.eecs.umich.edu/~weimerw/p/weimer-icsm2010.pdf
-
-Goal: Establish that we are smart for including all of these features. Note
-which individual features or feature categories were (not) included in
-previous work.
-
-Give a simple narrative about our effectiveness: implicitly, we were smart
-for deciding to include these features, better ingredients make better
-results, we have great results.
-
-TODO: what about turning on/off slicing?
-
-TODO: what about turning on/off types? (is that separate from slicing?)
+\tabref{anova} gives an alternate view of the relative feature importances
+as measured by ANOVA. In both tables we see that syntactic, dynamic, and
+contextual features are all important.
+
+We also try removing certain categories and retraining the model on what is
+left. \figref{removing-features} shows that there is a significant decrease in
+accuracy in each case, supporting the claim that all our feature categories are
+important.
+
+% prereq: define predictive power (leave-one-out, leave-one-in, Relief-F,
+% ANOVA, whatever)
+
+% Use as a rough guide: Table 3 on Page 8 of
+% https://web.eecs.umich.edu/~weimerw/p/weimer-icsm2010.pdf
+
+% Goal: Establish that we are smart for including all of these features. Note
+% which individual features or feature categories were (not) included in
+% previous work.
+
+% Give a simple narrative about our effectiveness: implicitly, we were smart
+% for deciding to include these features, better ingredients make better
+% results, we have great results.
+
+% TODO: what about turning on/off slicing?
+
+% TODO: what about turning on/off types? (is that separate from slicing?)
 
@@
 
-%In Nate there were 20 different functions. All the programs were doing the
-%same 20 functions.
-%
-%Here, people are doing anything they want. Random stuff going on over here.
-%
-%We want to show that we get high accuracy *over a heterogenous set of
-%programs*. Leon?
-%
-%prereq: cluster the programs based on similarity. Using agglomerative
-%clustering plus python edit distance. How good the performance will be?
-%right now it creates a clustering tree. binary tree, based on min distance.
-%can give it a threshhold for distance (elements w/in a group). don't forget
-%to rename all variables to X, perhaps normalize (edit distance size).
-%
-%leon: could also cluster by inconsistency coefficient
-%
-%PLOT Type: line graph (standard plot)
-%
-%PLOT X AXIS: clustering distance parameter
-%
-%PLOT Y AXIS: number of different clusters found
-%
-%PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''
-%
-%Goal: show that ``no matter how you decide program differences'', our
-%dataset is always more diverse/heterogeneous than is previous work. So our
-%success is amazing.
+% PLOT LINES: line 1 is ``Nate programs'', line 2 is ``25\%, at random, of our dataset''
 
@@
 
-As a measure of the diversity of our dataset, we clustered the programs in
-our dataset and the NATE dataset, then compared the number of resultant clusters.
+In the most similar prior work, Learning to Blame, each program in the dataset
+was a student's attempt at one of 23 different homework problems. Thus the
+dataset was comparatively homogenous, making it unclear whether the results
+would generalize to a more diverse dataset.
+
+In our PythonTutor data by comparision, users wrote whatever programs they
+wanted, so we
+hypothesized that the data were more heterogeneous. To confirm this
+quantitatively, we used agglomerative clustering both on our
+data and on the Learning to Blame data; whichever dataset naturally splits
+into more clusters is more heterogeneous.
+
+% The exact number of clusters returned
+% depends on the value we choose for the algorithm's threshold parameter, so we
+% checked all thresholds at .01 intervals, starting at 0 and ending at the point
+% where everything collapses into a single cluster. As shown in
+% \figref{diversity}, regardless of the choice of threshold, our data breaks into
+% more clusters than the NATE data. This confirms that our data is more
+% heterogeneous.
+
 To cluster programs, we transform the program ASTs into flat strings, and then

### 2018-08-22 paper/main.tex
added some to ml models

@@
 
-\emph{Decision Trees}
+\emph{Decision Tree}
 Decision tree algorithms learn a tree of binary predicates over the
@@
 %
-Random forests improve generalization by training an
+Random forests (RF)~\cite{breiman2001random} improve generalization by training an
 \emph{ensemble} of distinct decision trees and using a majority
 vote to make a prediction.
+For each tree, RF selects a random sample with replacement of the training
+set and trains the tree with these samples.
 %
@@
 also makes the learned model more difficult to interpret.
+\emph{Random Forest}
 
@@
 our dataset. The vectors from a random 30\% of the program pairs are set aside for testing and
-the other 70\% are used for training. Within the training data, we use 3-fold
+the other 70\% are used for training.
+Within the training data, we use 3-fold
 cross-validation to select the best parameters for our models, before training
 using those parameters on the full training set and then scoring the models
-using the testing set. Each trained model takes in any vector representing a term
+using the testing set.
+Each trained model takes in any vector representing a term
 in a buggy program, and returns a confidence score representing how likely it is that
@@
 confidence, where the top result is the one that the model deems most likely
-to be changed. For
-a given k, we score the model based on Top-k accuracy: for what proportion of the
-programs is a correct answer (i.e. a term that is actually changed) present in the
-top k results. We retrain the models for each value of k, since a model optimized
-to produce the single best result might be different from one optimized to get a
-good result into the top three.
+to be changed.
+For a given k, we score the model based on Top-k accuracy: for what
+proportion of the programs is a correct answer (i.e. a term that is actually
+changed) present in the top k results.
+We retrain the models for each value of k, since a model optimized to produce
+the single best result might be different from one optimized to get a good
+result into the top three.
+
+For decision tree and random forest, we use the implementation from
+\textsc{scikit-learn}~\cite{scikit-learn}.
+Due to the large amount of data points, we did not use \textsc{scikit-learn}'s
+implementation for logistic regression and multilayer perceptron.
+We implementation both model using
+\textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.
+
+Because of the imbalance nature of the data set, where not-blamed lines
+(negative examples) appears more frequently than blamed lines (positive
+examples).
+The weight for each example during training is re-weighted to the reciprocal
+of the frequency of corresponding class.
+
+\paragraph{Hyper-parameters}
+All hyper-parameters are tuned using three-fold cross-validation on the
+intended score (Top-k accuracy).
+For RF, the maximum depth of each tree is searched from $25$ to $35$.
+For decision tree maximum depth of the tree is searched from $25$ to $35$ with interval of $5$,
+the minimum impurity decrease for each split is searched from $10^{-7}$ to
+$10^{-3}$ with multiple of $10^{-2}$ and the minimum samples in each leaf is set to $200$.
+For logistic regression and MLP, binary crossentropy is set as the loss
+function and Adam \cite{kingma2014adam} optimizer is used.
+Learning rate is set to $0.1$.
+L2 regularization is applied to the weights of both models,
+the weight for L2 term is searched from $10^0$ to $10^{-5}$ with a multiple of $10^{-1}$.
+Other parameters that are not mentioned are kept with the default setting of the
+original packages.
 
@@
 cross-validation - the 3-fold thing we do is actually just to select parameters,
-and then there's no cross-validation on the final model, right?)
+and then there's no cross-validation on the final model, right? 
+%
+Correct
+)
 
@@
 \bibliographystyle{abbrv}
-\bibliography{nanomaly,sw,temp,slice,cluster}
+\bibliography{nanomaly,sw,temp,slice,cluster,ml}
 

### 2018-08-23 paper/main.tex
restructure the slicing description

@@
 \usepackage{graphicx}
+\usepackage{subcaption} 
 \usepackage{booktabs}
@@
 \usepackage{myref}
+\usepackage{url}
 \usepackage{pgfplots}
@@
 
+\newsavebox{\verbsavebox}
+
 \begin{document}
@@
 \subsubsection{Slice (Dynamic)}
+The goal of this feature, roughly equivalent in purpose to the type error
+slice in Nate~\cite{learning-to-blame}, is to help eliminate terms that
+cannot be the source of the crash. We compute a dynamic program
+slice~\cite{KOREL1988155}: a set of terms that contributed at runtime to
+the observed exception. This boolean feature encodes whether the term
+is a member of the slice. Our slicing algorithm, which balances scalable
+simplicity with coverage for beginner-written programs, is presented in
+\secref{slice-algorithm}.
+
+\subsubsection{Crash Location (Dynamic)} We observe that the precise term that
+causes Python to crash is, in fact, the one the user decides to change
+much of the time. While this information is not sufficient on its own
+(cf. the vanilla Python interpreter, which largely uses this feature),
+it can be a very helpful supplement. This boolean feature encodes whether
+or not the term is exactly the source location of the uncaught exception.
+
+\subsubsection{Exception Type (Dynamic)} The type of error thrown by the program
+contains useful information for its localization. In particular, a term may be
+more or less suspicious depending on the ultimate uncaught exception. As an
+example, every division term may be more suspicious if the uncaught exception
+being considered is \lt{DivisionByZero}. We encode the exception type as
+a categorical feature.
+% to \emph{every} vector derived from that program.
+
+\subsubsection{Contextual features}
+
+Term frequency vectors in general, and our bag of abstracted terms
+representation by extension, do not typically include contextual
+information such as ordering or relative positioning. As with many other
+analyses, such as contextual operational semantics~\cite{FIXME}, we observe
+that the meaning of an expression may depend on its surrounding context.
+
+As an example, consider the \lt{0} terms in both \lt{x = 0} and \lt{x / 0}.
+Both instances of \lt{0} have the same syntactic form, size and type
+(etc.). We may prefer to implicate the \lt{0} in \lt{x / 0} as suspicious,
+especially for beginner-written programs, but cannot distinguish it without
+surrounding contextual information. Data structures such as abstract syntax
+trees and control flow graphics capture such contextual information, but
+are not immediately applicable to machine learning.
+
+We desire to encode such information while retaining the use of scalable,
+accurate off-the-shelf machine learning algorithms that operate on feature
+vectors. We thus propose to embed contextual information in a vector,
+borrowing insights from standard approaches in machine learning. We
+associate with each term additional features that correspond to the features of
+its parent and child nodes. For representational regularity, we always
+model three children; terms without parents or children are given special
+values (e.g., zero or NotApplicable) for those contextual features.
+
+\subsection{Dynamic Slicing Algorithm}
+\label{sec-slice-algorithm}
 
 \begin{figure}
-\begin{lstlisting}
-x = input() # Input: 42
+\begin{lrbox}{\verbsavebox} 
+\begin{lstlisting}[xrightmargin=0.5\linewidth]
+x = input() # 42
 if x < 20:
@@
 \end{lstlisting}
-\begin{lstlisting}
+\end{lrbox} 
+~ \hfill
+\subcaptionbox{Pre-Slicing}{\usebox{\verbsavebox}}
+\hfill
+\begin{lrbox}{\verbsavebox}
+\begin{lstlisting}[xrightmargin=0.5\linewidth]
 x = 42
@@
 \end{lstlisting}
+\end{lrbox} 
+\subcaptionbox{Post-Slicing}{\usebox{\verbsavebox}}
+\hfill
+~
 \caption{
-  a) A program being sliced. An assertion gets triggered when line 8 is run. b)
-  The lines included in the slice. This is not a valid python
-  program, but that does not pose a problem for using membership in the slice as
-  a learning feature.
+  A program being sliced (left) and the resulting slice (right) with
+  respect to the assertion on line 8. 
 }
@@
 
-The goal of this feature, roughly equivalent in purpose to the type error
-slice in Nate~\cite{learning-to-blame}, is to help eliminate terms that
-cannot be the source of the crash. We compute a dynamic program slice
-\cite{KOREL1988155}, the set of terms that contributed at runtime to the
-observed exception, and turn this feature on for terms in the slice. We
-follow the basic approach of \cite{KOREL1988155, KOREL1990187}, building a
+Slicing information can prevent our model from implicating irrelevant
+nodes in the fault localization. Program slicing is a well-studied
+field with many explored tradeoffs (e.g., see Xu et al. for a
+survey~\cite{xu2005}). We desire a slicing algorithm that can be computed
+efficiently (to scale to hundreds of thousands of instances) but that will
+also admit high accuracy: we achieve this by focusing on features relevant
+to beginner-written programs. We
+follow the basic approach of Korel and Laski~\cite{KOREL1988155,
+KOREL1990187}, building a
 graph of data and control dependencies. We then traverse the graph
@@
 to be fed into the program whenever \lt{input()} is invoked. The slicer outputs
-the list of line numbers for the lines in the slice.
-
-The slicer would first create an execution trace of the example program, which
-at each execution step includes the line number being run and the state of the
-heap and variables. In the example, lines $1, 2, 3, 4, 8$ would have been run.
+the program elements included in the slice.
+
+The slicer proceeds by creating an execution trace of the example program. 
+At each execution step the trace includes the line number being run and the state of the
+heap and variables. In the example in \figref{slice-example}, 
+lines $1, 2, 3, 4$ and $8$ are executed. 
 It then iterates forward through the trace, recording which
@@
 or test-control relationship with a step already in the set. Finally, the
-algorithm takes all the execution steps, determines which line number each step
-was running, and then outputs this list of line numbers.
-
-Our slicing algorithm handles every syntax node supported by the Python3 standard
-library, \lt{ast}, except:
+algorithm maps all of the execution steps to their corresponding program 
+element locations and returns the result.
+
+To balance ease of prototype implementation against coverage for
+beginner-written programs, our slicing algorithm handles every syntax node
+supported by the Python3 standard \lt{ast} library except:
 \begin{itemize}
@@
 \item Assignments where attribute indexing or subscripting on the left hand side
-  means something other than the default, which can happen if the operations are
-  overridden by a class
-\item Lambda expressions
+  means something other than the default, (e.g., if the operations are
+  overridden by a class)
+\item Lambda, generator and starred expressions
 \item Set and dictionary comprehensions
-\item Generator expressions
 \item Await, yield, and yield from
-\item Ellipsis (the token representing the tail of the argument list given to a
-  varargs function)
-\item Starred expressions
-\item Async def functions and Asynchronous for
-\item Delete statements
-\item With statements
-\item Raise statements
+\item Variable argument ellipsis 
+\item Coroutine definitions and asynchronous loops
+\item Delete, with and raise statements
 \end{itemize}
 
-A downside of using the dynamic slice, in the case of control dependencies, is
-that it excludes terms whose presence does not affect whether the exception
-happened, but which could have had an effect if the term was different. In
+A downside of using a dynamic (as opposed to static) slice, in the case of
+control dependencies, is that it excludes terms whose presence does not
+affect whether the exception happened, but which could have had an effect
+if the term was different. In
 \figref{slice-downside-example}, the conditional statement is never
 executed, so the removal of the conditional would not have any effect on
-the program's behavior, and it would not be included in the program slice.
-The dynamic slice misses the incorrect test condition because the mistake
-itself caused the conditional statement and its body to become irrelevant.
+the program's behavior, and it would not be included in a more precise
+program slice. In this example, the dynamic slice misses the incorrect
+test condition because the very defect we are trying to localize
+caused the conditional statement and its body to become irrelevant.
 
@@
 dynamic slice.
-
-\subsubsection{Crash Location (Dynamic)} We observe that the precise term that
-causes Python to crash is, in fact, the one the user decides to change
-much of the time. While this information is not sufficient on its own
-(cf. the vanilla Python interpreter, which largely uses this feature),
-it can be a very helpful supplement. This boolean feature encodes whether
-or not the term is exactly the source location of the uncaught exception.
-
-\subsubsection{Exception Type (Dynamic)} The type of error thrown by the program
-contains useful information for its localization. In particular, a term may be
-more or less suspicious depending on the ultimate uncaught exception. As an
-example, every division term may be more suspicious if the uncaught exception
-being considered is \lt{DivisionByZero}. We encode the exception type as
-a categorical feature.
-% to \emph{every} vector derived from that program.
-
-\subsubsection{Contextual features}
-
-Term frequency vectors in general, and our bag of abstracted terms
-representation by extension, do not typically include contextual
-information such as ordering or relative positioning. As with many other
-analyses, such as contextual operational semantics~\cite{FIXME}, we observe
-that the meaning of an expression may depend on its surrounding context.
-
-As an example, consider the \lt{0} terms in both \lt{x = 0} and \lt{x / 0}.
-Both instances of \lt{0} have the same syntactic form, size and type
-(etc.). We may prefer to implicate the \lt{0} in \lt{x / 0} as suspicious,
-especially for beginner-written programs, but cannot distinguish it without
-surrounding contextual information. Data structures such as abstract syntax
-trees and control flow graphics capture such contextual information, but
-are not immediately applicable to machine learning.
-
-We desire to encode such information while retaining the use of scalable,
-accurate off-the-shelf machine learning algorithms that operate on feature
-vectors. We thus propose to embed contextual information in a vector,
-borrowing insights from standard approaches in machine learning. We
-associate with each term additional features that correspond to the features of
-its parent and child nodes. For representational regularity, we always
-model three children; terms without parents or children are given special
-values (e.g., zero or NotApplicable) for those contextual features.
 

### 2018-08-23 paper/main.tex
readability changes to the experiments

@@
 \end{lstlisting}
-\caption{In this program, the programmer may have intended the placeholder
-  assignment \lt{x = 1} to be used in case the user input 0, and used the
-  incorrect condition \lt{x != 0}, rather than \lt{x == 0}. However, the
+\caption{In this program, the programmer intended the placeholder
+  assignment \lt{x = 1} to be used in case the user inputs 0. 
+  The defect is that the 
+  incorrect condition \lt{x != 0} (rather than \lt{x == 0}) was employed. However, the
   condition expression will not be included in the slice (which would include
@@
   In this example, the programmer attempts to handle invalid input by
-  breaking out of the input loop. When computing the slice, our method
-  would record a control dependency between the condition on line 3 and the
-  print statement on line 5, placing the bad condition in the program
-  slice.
+  breaking out of the input loop. Our method
+  records a control dependency between the condition on line 3 and the
+  print statement on line 5, correctly placing the buggy condition in the
+  program slice.
 }
@@
 \end{tikzpicture}
-\caption{Baseline is the normal Python interpreter. Other bars are from decision
-trees on the largest dataset (272534 pairs).}
+\caption{Fault localization accuracy. 
+Baseline is the normal Python interpreter. Other bars represent our
+approach, using decision trees, 
+on the largest dataset (272534 pairs).}
 \label{fig-full-dataset-acc}
@@
 \end{tikzpicture}
-\caption{Baseline is the normal Python interpreter. Other bars are from random
-forests on the largest dataset (272534 pairs). TODO: get top-2, then change text
-to discuss this figure instead of decision-tree version}
+\caption{Fault localization accuracy. 
+Baseline is the normal Python interpreter. Other bars represent
+our approach, using random forests, 
+on the largest dataset (272534 pairs). \fixme{get top-2, then change text
+to discuss this figure instead of decision-tree version}}
 \label{fig-full-dataset-acc-random-forest}
@@
 \begin{table}[]
-\begin{tabular}{lll}
+\begin{tabular}{llc}
 Name & Category & Gini Importance \\ \bottomrule
@@
 \begin{table}[]
-\begin{tabular}{lll}
+\begin{center}
+\begin{tabular}{llc}
 Name & Category & F-score \\ \bottomrule
@@
 \end{tabular}
-\caption{Feature predictive power as measured by ANOVA. F-score in
-thousands to two significant figures; all p-values less than .01}
+\end{center}
+\caption{Feature predictive power as measured by ANOVA.  The F-score in
+shown in 
+thousands to two significant figures; all p-values are less than 0.01.}
 \label{tab-anova}
@@
 \caption{Accuracy when features are removed, based on decision trees on a
-random subset of 20000 program pairs.}
+random subset of 20000 program pairs.
+\fixme{The x-axis labels are not perfectly clear: does ``all'' mean that
+everything is still in or that everything is out?}
+\fixme{Indicate that this is a top-1 model.}
+\fixme{Normal the y-axis so that ``all'' is 1.0 to make the subset effects
+more evident.} 
+}
 \label{fig-removing-features}
@@
 is used to split a node: a feature that is learned to guide more model
-classification decisions is more important.
-
-\tabref{anova} gives an alternate view of the relative feature importances
-as measured by ANOVA. In both tables we see that syntactic, dynamic, and
-contextual features are all important.
-
-We also try removing certain categories and retraining the model on what is
-left. \figref{removing-features} shows that there is a significant decrease in
-accuracy in each case, supporting the claim that all our feature categories are
-important.
+classification decisions is more important. Similarly, \tabref{anova} gives
+an alternate view of the relative feature importances, as measured by an
+analysis of variance (ANOVA).
+
+We also present the results of a leave-one-out analysis in which 
+certain categories of features are removed and the model is trained
+and tested only on those that remain. \figref{removing-features} shows
+that there is a significant decrease in accuracy in each case. We note,
+however, that leave-one-out analyses underestimate importance in cases of
+feature overlap. For example, if a small program contains a string bug but
+only a few string variables, both type information and slice information
+are likely to implicate similar terms.
+
+In all cases we see that syntactic, dynamic, and contextual features are
+all important. Features such as FIXME, FIXME and FIXME are found to be of
+primary importance. We interpret this to mean FIXME with respect to the
+application of our algorithm to beginner-written code.
 
@@
 \caption{Accuracy when the model is trained and tested on only the data exhibiting
-the error on the x-axis.}
+the error on the x-axis.
+\fixme{Spin this for generality. Indicate that our model gives high-quality
+fault localization regardless of defect type. To put it another way, we
+can view this as a sensitivity analysis: our algorithm is not sensitive to
+defect type (that's good!).} 
+}
 \label{fig:defect-categories}

### 2018-08-23 paper/main.tex
more suggestions for the evaluation section

@@
 \end{tikzpicture}
-\caption{Number of clusters when data is clustered by single linkage.
-  A linkage tree is built using agglomerative clustering, and the resulting tree is broken up
-  into clusters by thresholding the inconsistency coefficient. The inconsistency coefficient
-  of a link is its distance, z-scored against other nearby links. TODO: find a way to include
-  the Rosetta Code baseline in this graph? better not to devote multiple graphs to this issue.
-  LM: the y axis could be made into a percentage of the dataset size, but that might be
-  weird?}
+\caption{Number of natural program clusters in each dataset as a function
+of clustering parameter (informally, how ``strict'' or ``tight'' the
+clustering is). \fixme{Talk about how you sample to be fair to the smaller
+datasets.} At all points our dataset demonstrates more diversity than 
+that of previous work.} 
 \label{fig-diversity}
@@
 
-In the most similar prior work, Learning to Blame, each program in the dataset
-was a student's attempt at one of 23 different homework problems. Thus the
-dataset was comparatively homogenous, making it unclear whether the results
-would generalize to a more diverse dataset.
-
-In our PythonTutor data by comparision, users wrote whatever programs they
-wanted, so we
-hypothesized that the data were more heterogeneous. To confirm this
-quantitatively, we used agglomerative clustering both on our
-data and on the Learning to Blame data; whichever dataset naturally splits
-into more clusters is more heterogeneous.
+To demonstrate that our evaluation dataset is not only larger but also more
+diverse than those used in previous work, we compare the diversity of
+programs present here to those in a relevant balance. The Nate
+algorithm~\cite{learning-to-blame} also provides type error localization
+and also focuses on a machine-learning approach backed by slicing. However,
+Nate targets strongly statically typed OCaml programs. Nate's evaluation
+also focused on beginner-written programs: specifically, it considered
+student submissions to 23 different university homework problems.  Thus the
+Nate dataset is comparatively homogeneous, raising concerns about whether
+evaluation results would generalize to more diverse settings.
+
+In our PythonTutor dataset, by comparison, users were not constrained to
+specific university assignments. We hypothesize that the evaluation data are
+more heterogeneous. To assess this quantitatively, we used agglomerative
+clustering to discover the number of ``natural'' program categories
+present in both our dataset and also in the Nate dataset. Datasets with
+more distinct program categories are more heterogeneous.
 
@@
 
-To cluster programs, we transform the program ASTs into flat strings, and then
-perform agglomerative clustering on them, with string edit distance as a distance
-metric.
+% To cluster programs, we transform the program ASTs into flat strings, and
+% then perform agglomerative clustering on them, with string edit distance
+% as a distance metric.
 
@@
 
-To measure the distance between programs, we flattened their ASTs into strings
-of tokens, and then computed the Levenshtein edit distance~\cite{levenshtein}.
-The decision to flatten the tree rather than compute a distance directly on the
-the ASTs is practical; the state of the art in tree edit distance has cubic
-asymptotic complexity~\cite{PAWLIK2016157}, while the levenshtein distance can be
-computed in time quadratic in the length of the strings~\cite{lev-quadratic}.
-When we attempted a test run using tree edit distance as the metric, it did not
-complete and we determined that it would not be practical to use.
-
-The flattening process, which is based on \cite{dist-site}, is defined
-recursively: the string for an AST
-begins with a unique token corresponding to the AST node type, which is
-followed by the concatenation of the transformations of the tree's subtrees.
-The string ends with an ``end'' token to denote the end of the tree. This
-preserves the tree structure of the AST.
+Many clustering algorithms depend on distance metrics. 
+To measure the distance between programs, we flattened their ASTs into
+strings of tokens, and then computed the Levenshtein edit
+distance~\cite{levenshtein}. The decision to flatten the tree rather than
+compute a distance directly on the ASTs relates to scalability for our
+large datasets. The state of the art in tree edit distance has cubic
+asymptotic complexity~\cite{PAWLIK2016157}, while the Levenshtein distance
+can be computed in time quadratic in the length of the
+strings~\cite{lev-quadratic}. We did implement and attempt a tree edit
+distance metric, but it did not terminate for this analysis in a practical
+amount of time.
+
+The flattening process, which is based on standard approaches for
+tree-structured data~\cite{dist-site}, is defined recursively: the string
+for an AST begins with a unique token corresponding to the AST node type,
+which is followed by the concatenation of the transformations of the tree's
+subtrees.  The string ends with an ``end'' token to denote the end of the
+tree. This embeds information related to the tree structure of the AST in
+the flattened string.
 
@@
 
-We performed agglomerative clustering on the programs~\cite{modern-clustering}.
+We performed agglomerative clustering on the 
+datasets of flattened programs~\cite{modern-clustering}.
 This is a method of clustering in which every datapoint starts in its own
@@
 each internal node corresponds to a cluster formed by merging two other clusters,
-and each leaf corresponds to an item in the dataset. We used single linkage,
-which means that the distance between clusters of multiple items was considered
+and each leaf corresponds to an item in the dataset. We used a single linkage
+approach in which the distance between clusters of multiple items was considered
 to be the minimum distance between their elements.
 
-This alone produces a tree rather than a list of disjoint clusters, which makes
-different clusterings difficult to compare. To produce a cluster count, we used
-the SciPy function \lt{scipy.cluster.hierarchy.fcluster}, which forms disjoint
-clusters by splitting the tree at internal nodes where the two joined clusters
-are more distant than some threshold. We used the \lt{'inconsistent'} method with
-depth set to 2, which specifies that the distances at each node should be
-z-scored against the distances at other nodes at the same level in the tree, and
-any nodes 1 or 2 levels below~\cite{inconsistent}. This adapts to differences in
-scale of edit distance that might arise between python and OCaml due to
-programming language expressivity or other factors, making cluster counts at each
-threshold value comparable.
-
-\figref{diversity} shows the results of clustering on our dataset and the NATE
-dataset. For all values of the inconsistency threshold, the algorithm computed a
-higher cluster count for the Python data, suggesting that it contains a more
-diverse set of programs than the OCaml dataset. The similarity in the graphs
-appears to confirm the comparibility of cluster counts between the two datasets.
+To produce a cluster count from the agglomerative tree, we followed the
+standard approach of splitting the tree at internal nodes where the two
+joined clusters are more distant than some threshold. To account for
+potential programming languages differences between Python and OCaml (such
+as expressivity), we z-score nodes against not only others at the same tree
+depth, but also against against others one or two levels
+below~\cite{inconsistent}. This makes cluster counts at each threshold
+value comparable. Our implementation uses the standard
+SciPy library (e.g., \lt{scipy.cluster.hierarchy.fcluster} with the 
+\lt{inconsistent} method).
+
+\fixme{We need to say something about how we controlled for number, lest
+reviewers think we have more clusters simply because we have hundreds of
+thousands of datapoints.}
+
+\figref{diversity} shows the results of clustering on our dataset and the
+Nate dataset. At a high level, the structural similarity in the graphs
+confirms the comparability of cluster counts between the two datasets.  For
+all values of the inconsistency threshold, there are more clusters in our
+Python dataset than in the baseline OCaml dataset. This suggests that our
+dataset contains a more diverse set of programs. Note that we are not
+claiming any novelty or advances in clustering accuracy in this
+determination (indeed, scalability concerns limited us to coarser
+approaches); instead, our claim is that even with simple clustering,
+it is clear that our dataset contains a greater diversity of programs than
+were considered by previously-published evaluations.  
+We view it as an advantage of our algorithm that it can apply to multiple
+different program categories.
 

### 2018-08-23 learning/decisionpath.py
update visualizer; remove obsolete learning folder

@@ -1,186 +1,0 @@
-import math
-import os.path
-import random
-random.seed()
-
-from sklearn import tree
-from sklearn.ensemble import RandomForestClassifier
-import numpy as np
-import pandas as pd
-
-
-# import plotly.plotly as py
-# import plotly.graph_objs as go
-import pydotplus
-# from IPython.display import Image
-
-import sys
-
-import input_old
-
-FIRST_NON_SLICE_FEATURE = 'F-Is-Var'
-
-model = sys.argv[1]
-test, fs, ls = input_old.load_csv(sys.argv[2], filter_no_labels=True, only_slice=True)
-
-
-# print test
-test_samps = test.loc[:,FIRST_NON_SLICE_FEATURE:]
-test_labels = test.loc[:,'L-DidChange']
-test_span = test.loc[:,'SourceSpan']
-# print test.iloc[1]
-# print test.values[1]
-feature_names = fs[1:]
-
-
-#------- Loading
-from sklearn.externals import joblib
-estimator = joblib.load(model)
-
-
-# # -----PLOTTING
-dot_data = tree.export_graphviz(estimator, out_file=None,
-	feature_names=feature_names,
-	filled=True,
-	rounded=True)
-graph = pydotplus.graph_from_dot_data(dot_data)
-# graph.write_png(model+'.png')
-# ---------
-
-# graph.render(filename='img/g1')
-graph.write_pdf(model+".pdf")
-# --------------
-
-X_test = test_samps.values
-
-
-# Using those arrays, we can parse the tree structure:
-
-n_nodes = estimator.tree_.node_count
-children_left = estimator.tree_.children_left
-children_right = estimator.tree_.children_right
-feature = estimator.tree_.feature
-threshold = estimator.tree_.threshold
-
-# The tree structure can be traversed to compute various properties such
-# as the depth of each node and whether or not it is a leaf.
-# node_depth = np.zeros(shape=n_nodes)
-# is_leaves = np.zeros(shape=n_nodes, dtype=bool)
-# stack = [(0, -1)]  # seed is the root node id and its parent depth
-# while len(stack) > 0:
-#     node_id, parent_depth = stack.pop()
-#     node_depth[node_id] = parent_depth + 1
-
-#     # If we have a test node
-#     if (children_left[node_id] != children_right[node_id]):
-#         stack.append((children_left[node_id], parent_depth + 1))
-#         stack.append((children_right[node_id], parent_depth + 1))
-#     else:
[…]

### 2018-08-23 learning/decisiontree2.py

@@ -1,188 +1,0 @@
-import math
-import os.path
-import random
-random.seed()
-
-from sklearn import tree
-from sklearn.ensemble import RandomForestClassifier
-import numpy as np
-import pandas as pd
-
-import input_old
-
-csvs2 = [f for f in os.listdir('ml2/data/fa15/op+type+size') if f.endswith('.csv')]
-csvs = [f for f in os.listdir('ml2/data/sp14/op+type+size') if f.endswith('.csv')]
-
-random.shuffle(csvs)
-dfs = []
-test = []
-train = []
-
-for csv in csvs2:
-	df, fs, ls = input_old.load_csv(os.path.join('ml2/data/fa15/op+type+size', csv), filter_no_labels=True, only_slice=False)
-
-	if df is None:
-		continue
-	if df.shape[0] == 0:
-		continue
-
-	test.append(df)
-
-for csv in csvs:
-	df2, fs2, ls2 = input_old.load_csv(os.path.join('ml2/data/sp14/op+type+size', csv), filter_no_labels=True, only_slice=False)
-
-	if df2 is None:
-		continue
-	if df2.shape[0] == 0:
-		continue
-
-	train.append(df2)
-
-
-
-train = pd.concat(train)
-test = pd.concat(test)
-
-#print (len(test))  
-#print (len(train))  
-
-
-# print (df.shape)
-# print df
-classes = list(train.groupby(ls2))
-#print(ls)
-max_samples = max(len(c) for _, c in classes)
-train = pd.concat(c.sample(max_samples, replace=True) for _, c in classes)
-#print (len(train))  
-#print df.shape
-#print type(df)
-#list_keys = [ k for k in df ]
-#print list_keys
-# print samps
-
-#print sum(df['L-DidChange'].values)
-# print df['L-DidChange'].index
-
-
-
-train_samps = train.loc[:,'F-InSlice':]
-train_labels = train.loc[:,'L-DidChange']
-
-# print test
-test_samps = test.loc[:,'F-InSlice':]
-test_labels = test.loc[:,'L-DidChange']
-test_span = test.loc[:,'SourceSpan']
-# print test.iloc[1]
-# print test.values[1]
-
-
-# dflist = []
-# keylist = []
-# for key, value in df.iteritems():
-#     temp = value
-#     tempk = key
-#     dflist.append(temp)
-#     keylist.append(tempk)
[…]

### 2018-08-23 learning/hidden.py

@@ -1,234 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import os.path
-import pandas as pd
-import tensorflow as tf
-
-import json
-import math
-from time import time
-
-import util
-
-def build_model(features, labels, hidden,
-                learn_rate=0.1, beta=0.01, beta_out=0.01, model_dir=None):
-    '''Build a linear classifier with fully-connected hidden layers.
-
-    @param features: A list of feature names.
-    @param labels: A list of label names.
-    @param hidden: A list of ints indicating how many neurons to put in each hidden layer.
-    @param learn_rate: The training rate, defaults to 0.1.
-    @param beta: The regularization rate, defaults to 0.01.
-    @param beta_out: The regularization rate for the output layer, defaults to 0.01.
-    @param model_dir: A directory to store the model summaries in.
-
-    @return: A 4-tuple of training, testing, plotting, and closing functions.
-    '''
-
-    #print 'building MLP with alpha=%f, beta=%f' % (learn_rate, beta)
-
-    n_in = len(features)
-    n_out = len(labels)
-
-    x = tf.placeholder(tf.float64, [None, n_in], name='x')
-    keep_prob = tf.placeholder(tf.float64, name='keep_prob')
-
-    global_step = tf.Variable(0, name='global_step', trainable=False)
-
-    h = x
-    n = n_in
-    Ws = []
-    bs = []
-    for i, n_hidden in enumerate(hidden):
-        n_hidden = int(n_hidden)
-        with tf.name_scope('hidden-' + str(i)):
-            W = tf.Variable(
-                tf.truncated_normal([n, n_hidden],
-                                    # we're using a ReLU, so initialize weights with
-                                    # a variance of 2/n (according to He et al 2015)
-                                    stddev=math.sqrt(2.0/float(n)),
-                                    dtype=tf.float64),
-                name='W_' + str(i))
-            b = tf.Variable(tf.constant(0.0, shape=[n_hidden], dtype=tf.float64),
-                            name='b_' + str(i))
-            util.variable_summaries(W,'W_' + str(i))
-            util.variable_summaries(b,'b_' + str(i))
-            Ws.append(W)
-            bs.append(b)
-            # h = tf.nn.dropout(tf.nn.relu(tf.matmul(h, W) + b), keep_prob)
[…]

### 2018-08-23 learning/input.py

@@ -1,58 +1,0 @@
-import pandas as pd
-import numpy as np
-
-def load_csv(path, filter_no_labels=False, balance_labels=True, only_slice=False, no_slice=False):
-    '''Load feature vectors from a csv file.
-
-    Expects a header row with feature columns prefixed with 'F-' and
-    label columns prefixed with 'L-'.
-
-    @param filter_no_labels: if True, filter out samples where all labels are 0.
-
-    @param balance_labels: if True, balance the count of samples from
-    each class of labels, by duplicating samples from under-represented
-    classes.
-
-    @return: (dataframe, feature names, label names)
-
-    '''
-    df = pd.read_csv(path)
-    label_names = [c for c in df.columns if c[0] == 'L']
-    feature_names = [c for c in df.columns if c[0] == 'F']
-
-    if filter_no_labels:
-        # print df.shape
-        # filter out vectors with no predictions
-        criteria = (df[l] == 1.0 for l in label_names)
-        df = df[reduce(lambda x, acc: x | acc, criteria)]
-        # print df.shape
-
-
-    if only_slice:
-        if len(df[df['L-DidChange'] == 1]) == 0:
-            print 'no changes', path
-            df = None
-            return (df, feature_names, label_names)
-        df = df[df['F-InSlice'] == 1].reset_index(drop=True)
-        if len(df) == 1:
-            print '1 sliced', path
-            df = None
-            return (df, feature_names, label_names)
-        if len(df[df['L-DidChange'] == 1]) == 0:
-            print 'no overlap', path
-            df = None
-
-    if no_slice or only_slice:
-        #del df['F-InSlice']
-        feature_names = [f for f in feature_names if f != 'F-InSlice']
-
-
-    # if balance_labels:
-    #     print df.shape
-    #     classes = df.groupby(label_names)
-    #     max_samples = max(len(c) for _, c in classes)
-    #     print max_samples
-    #     df = pd.concat(c.sample(max_samples, replace=True) for _, c in classes)
-    #     print df.shape
-
-    return (df, feature_names, label_names)

### 2018-08-23 learning/input_old.py

@@ -1,55 +1,0 @@
-import pandas as pd
-import numpy as np
-import ntpath
-
-
-def load_csv(path, filter_no_labels=False, balance_labels=True, only_slice=False):
-    '''Load feature vectors from a csv file.
-
-    Expects a header row with feature columns prefixed with 'F-' and
-    label columns prefixed with 'L-'.
-
-    @param filter_no_labels: if True, filter out samples where all labels are 0.
-
-    @param balance_labels: if True, balance the count of samples from
-    each class of labels, by duplicating samples from under-represented
-    classes.
-
-    @return: (dataframe, feature names, label names)
-
-    '''
-    df = pd.read_csv(path)
-    filenum = (ntpath.basename(path)).split('.')[0]
-    df['new_index'] = int(filenum)*(np.ones((len(df.index),1)))
-    df = df.set_index('new_index')
-    label_names = [c for c in df.columns if c[0] == 'L']
-    feature_names = [c for c in df.columns if c[0] == 'F']
-
-    if filter_no_labels:
-        # print df.shape
-        # filter out vectors with no predictions
-        criteria = (df[l] == 1.0 for l in label_names)
-        df = df[reduce(lambda x, acc: x | acc, criteria)]
-        # print df.shape
-
-    if only_slice:
-        if len(df[df['L-DidChange'] == 1]) == 0:
-            df = None
-            return (df, feature_names, label_names)
-        df = df[df['F-InSlice'] == 1].reset_index(drop=True)
-        if len(df) == 0 or len(df) == 1:
-            df = None
-            return (df, feature_names, label_names)
-        if len(df[df['L-DidChange'] == 1]) == 0:
-            print path
-            df = None
-
-    # if balance_labels:
-    #     print df.shape
-    #     classes = df.groupby(label_names)
-    #     max_samples = max(len(c) for _, c in classes)
-    #     print max_samples
-    #     df = pd.concat(c.sample(max_samples, replace=True) for _, c in classes)
-    #     print df.shape
-
-    return (df, feature_names, label_names)

### 2018-08-23 learning/learn.py

@@ -1,246 +1,0 @@
-from __future__ import print_function
-
-import math
-import os.path
-import random
-import sys
-import time
-
-from hyperopt import fmin, tpe, hp
-import numpy as np
-import pandas as pd
-from sklearn.model_selection import KFold, ShuffleSplit
-import tensorflow as tf
-
-# local modules
-import hidden
-import input
-import linear
-import svm
-import util
-
-flags = tf.app.flags
-FLAGS = flags.FLAGS
-
-flags.DEFINE_bool("verbose", False, "Print verbose output.")
-flags.DEFINE_string("data", "", "Path to the data.")
-flags.DEFINE_string("test_data", "", "Path to test data (if separate from training).")
-flags.DEFINE_string("model", "linear",
-                    "Valid model types: {'linear', 'svm', 'hidden'}.")
-flags.DEFINE_string("model_dir", None, '')
-flags.DEFINE_string("hidden_layers", "", "")
-flags.DEFINE_integer("batch_size", 100, "Size of each training minibatch.")
-flags.DEFINE_integer("n_batches", None, "Number of training rounds.")
-flags.DEFINE_integer("n_epochs", 1, "Number of training rounds.")
-flags.DEFINE_float("learn_rate", 0.1, "Learning rate.")
-flags.DEFINE_float("reg_rate", 0.01, "Regularization rate.")
-flags.DEFINE_float("reg_out_rate", None, "Regularization rate for output (defaults to reg_rate).")
-flags.DEFINE_integer("n_folds", 1, "Number of folds for cross-validation.")
-flags.DEFINE_bool("plot", False, "Plot the learned model.")
-flags.DEFINE_bool("only_slice", False, "Only look at exprs in the error slice.")
-flags.DEFINE_bool("no_slice", False, "Ignore the error slice feature.")
-flags.DEFINE_bool("store_predictions", False,
-                  "Whether to store predictions for offline analysis.")
-flags.DEFINE_bool("hyperopt", False, "Whether to use hyperopt to tune parameters.")
-flags.DEFINE_integer("seed", None, '')
-
-
-def load_data(path):
-    csvs = sorted([f for f in os.listdir(path) if f.endswith('.csv')])
-    random.shuffle(csvs)
-    dfs = []
-    for i, csv in enumerate(csvs):
-        f = os.path.join(path, csv)
[…]

### 2018-08-23 learning/linear.py

@@ -1,181 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import os
-import tensorflow as tf
-
-import math
-from time import time
-
-import util
-
-def build_model(features, labels, learn_rate=0.1, beta=0.01, model_dir=None):
-    '''Build a linear classifier.
-
-    @param features: A list of feature names.
-    @param labels: A list of label names.
-    @param learn_rate: The training rate, defaults to 0.1.
-    @param beta: The regularization rate, defaults to 0.01.
-    @param model_dir: A directory to store the model summaries in.
-
-    @return: A 4-tuple of training, testing, plotting, and closing functions.
-    '''
-
-    n_in = len(features)
-    n_out = len(labels)
-
-    x = tf.placeholder(tf.float64, [None, n_in], name='x')
-
-    global_step = tf.Variable(0, name='global_step', trainable=False)
-
-    with tf.name_scope('linear'):
-        W = tf.Variable(
-            tf.truncated_normal([n_in, n_out],
-                                stddev=1.0 / math.sqrt(float(n_in)),
-                                dtype=tf.float64),
-            name='W')
-        b = tf.Variable(tf.zeros([n_out], dtype=tf.float64), name='b')
-        util.variable_summaries(W,'W')
-        util.variable_summaries(b,'b')
-        y = tf.matmul(x, W) + b
-
-    y_ = tf.placeholder(tf.float64, [None, n_out], name='y_')
-
-    with tf.name_scope('cross_entropy'):
-        cross_loss = tf.reduce_mean(tf.nn.softmax_cross_entropy_with_logits(logits=y, labels=y_))
-        tf.summary.scalar('cross_loss', cross_loss)
-        regularizers = beta * tf.nn.l2_loss(W)
-        tf.summary.scalar('l2_loss', regularizers)
-        loss = cross_loss + regularizers
-        tf.summary.scalar('loss', loss)
-    with tf.name_scope('train'):
-        train_step = tf.train.AdamOptimizer(learn_rate).minimize(loss, global_step=global_step)
-
-    sess = tf.InteractiveSession()
-    merged = tf.summary.merge_all()
-    if model_dir:
-        summary_writer = tf.summary.FileWriter(model_dir, sess.graph)
-
-    if n_out >= 2:
-        correct_prediction = tf.equal(tf.argmax(tf.nn.softmax(y),1), tf.argmax(y_,1))
-    else:
[…]

### 2018-08-23 learning/plot.py

@@ -1,9 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-
-def plot_weights(w, feature_names, label_names):
-    plt.matshow(w, cmap='hot', interpolation='nearest')
-    plt.xticks(np.arange(len(feature_names)), feature_names, rotation=90)
-    plt.yticks(np.arange(len(label_names)), label_names)
-    # plt.legend()
-    plt.show()

### 2018-08-23 learning/randomforest2.py

@@ -1,179 +1,0 @@
-import math
-import os.path
-import random
-random.seed()
-
-from sklearn import tree
-from sklearn.ensemble import RandomForestClassifier
-import numpy as np
-import pandas as pd
-
-import input_old
-
-csvs2 = [f for f in os.listdir('ml2/data/fa15/op+context-count+type+size') if f.endswith('.csv')]
-csvs = [f for f in os.listdir('ml2/data/sp14/op+context-count+type+size') if f.endswith('.csv')]
-
-random.shuffle(csvs)
-dfs = []
-test = []
-train = []
-
-for csv in csvs:
-	df, fs, ls = input_old.load_csv(os.path.join('ml2/data/sp14/op+context-count+type+size', csv), filter_no_labels=True, only_slice=False)
-
-	if df is None:
-		continue
-	if df.shape[0] == 0:
-		continue
-
-	train.append(df)
-
-for csv in csvs2:
-	df2, fs2, ls2 = input_old.load_csv(os.path.join('ml2/data/fa15/op+context-count+type+size', csv), filter_no_labels=True, only_slice=False)
-
-	if df2 is None:
-		continue
-	if df2.shape[0] == 0:
-		continue
-
-	test.append(df2)
-
-
-
-train = pd.concat(train)
-test = pd.concat(test)
-
-# print (len(test))  
-# print (len(train))  
-
-
-# print (df.shape)
-# print df
-classes = list(train.groupby(ls2))
-#print(ls)
-max_samples = max(len(c) for _, c in classes)
-train = pd.concat(c.sample(max_samples, replace=True) for _, c in classes)
-# print (len(train))  
-#print df.shape
-#print type(df)
-#list_keys = [ k for k in df ]
-#print list_keys
-# print samps
-
-#print sum(df['L-DidChange'].values)
-# print df['L-DidChange'].index
-
-
-
-train_samps = train.loc[:,'F-InSlice':]
-train_labels = train.loc[:,'L-DidChange']
-
-# print test
-test_samps = test.loc[:,'F-InSlice':]
-test_labels = test.loc[:,'L-DidChange']
-test_span = test.loc[:,'SourceSpan']
-
-# print test.iloc[1]
-# print test.values[1]
-
-# dflist = []
-# keylist = []
-# for key, value in df.iteritems():
-#     temp = value
-#     tempk = key
[…]

### 2018-08-23 learning/simple.py

@@ -1,227 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import pandas as pd
-import tensorflow as tf
-import tensorflow.contrib.learn.python.learn as learn
-
-import csv
-import json
-import math
-import random
-random.seed()
-
-def read_data(path):
-    df = pd.read_csv('data/type-inference+vars.csv')
-    label_names = [c for c in df.columns if c[0] == 'L']
-    feature_names = [c for c in df.columns if c[0] == 'F']
-
-    print df.shape
-    # filter out vectors with no predictions
-    criteria = (df[l] == 1.0 for l in label_names)
-    df = df[reduce(lambda x, acc: x | acc, criteria)]
-    print df.shape
-
-    return (df, feature_names, label_names)
-
-def plot_weights(w, x_labels, y_labels):
-    plt.matshow(w, cmap='hot', interpolation='nearest')
-    plt.xticks(np.arange(len(x_labels)), x_labels, rotation=90)
-    plt.yticks(np.arange(len(y_labels)), y_labels)
-    # plt.legend()
-    plt.show()
-
-
-df, feature_names, label_names = read_data('data/type-inference+vars.csv')
-
-N_CATS = len(feature_names)
-N_OUTS = len(label_names)
-
-N_FOLDS = 1
-
-# select an equal number of good and bad samples
-# N_SAMPLES = 100000
-# good=0
-# bad=0
-# samples = []
-# for l in orig:
-#     try:
-#         d = json.loads(l)
-#         # if d['kind'] == 'Good':
-#         #     good+=1
-#         #     if good > N_SAMPLES:
-#         #         continue
-#         # else:
-#         #     bad+=1
-#         #     if bad > N_SAMPLES:
-#         #         continue
-#         fs = [float(f) for f in d['features']]
-#         ls = [float(l) for l in d['kind']]
-#         if ls == [0. for l in d['kind']]:
-#             #print 'skipping expr with unknown type'
-#             continue
-#         # if N_OUTS == 1:
-#         #     l = [1.] if d['kind'] == 'Bad' else [-1.]
-#         # else:
-#         #     l = [1., 0.] if d['kind'] == 'Bad' else [0., 1.]
-#     except:
-#         continue
[…]

### 2018-08-23 learning/svm.py

@@ -1,49 +1,0 @@
-import matplotlib.pyplot as plt
-import numpy as np
-import tensorflow as tf
-
-import math
-
-import util
-
-def build_model(features, labels, learn_rate=0.1, model_dir=None):
-    '''Build a linear classifier.
-
-    @param features: A list of feature names.
-    @param labels: A list of label names.
-    @param learn_rate: The training rate, defaults to 0.1.
-    @param model_dir: A directory to store the model summaries in.
-
-    @return: A 4-tuple of training, testing, plotting, and closing functions.
-    '''
-
-    ## TODO: construct actual model
-
-    def train(data, i, validation=None, verbose=False):
-        summary_str, _ = sess.run([merged, train_step],
-                                  feed_dict={x: data[features], y_: data[labels]})
-        summary_writer.add_summary(summary_str, i)
-        if validation is not None:
-            acc = sess.run(accuracy,
-                           feed_dict={x: validation[features], y_: validation[labels]})
-            if verbose and i % 100 == 0:
-                print('accuracy at step {}: {}'.format(i, acc))
-
-
-    def test(data):
-        acc = sess.run(accuracy, {x: data[features], y_: data[labels]})
-        print('accuracy: %f' % acc)
-
-    def plot():
-        w = sess.run(tf.transpose(W))
-        plt.matshow(w, cmap='hot', interpolation='nearest')
-        plt.xticks(np.arange(len(features)), features, rotation=90)
-        plt.yticks(np.arange(len(labels)), labels)
-        # plt.legend()
-        plt.show()
-
-    def close():
-        sess.close()
-        tf.reset_default_graph()
-
-    return train, test, plot, close

### 2018-08-23 learning/trees.py

@@ -1,242 +1,0 @@
-import math
-import os.path
-import sys
-
-import sklearn
-from sklearn import tree, neural_network
-from sklearn.ensemble import RandomForestClassifier
-import numpy as np
-import pandas as pd
-
-import input_old
-
-model = sys.argv[1]
-if model not in ['mlp', 'decision-tree', 'random-forest']:
-        print 'python trees.py [decision-tree|random-forest] <train> <test>'
-        sys.exit(1)
-train_dir = sys.argv[2]
-test_dir = sys.argv[3]
-
-FIRST_NON_SLICE_FEATURE = 'F-Is-Var'
-
-RANDOM_SEED = 0
-np.random.seed(RANDOM_SEED)
-prng=np.random.RandomState(RANDOM_SEED)
-
-# FIXME: sort??
-train_csvs = sorted([f for f in os.listdir(train_dir) if f.endswith('.csv')])
-test_csvs = sorted([f for f in os.listdir(test_dir) if f.endswith('.csv')])
-
-sklearn.utils.shuffle(train_csvs, random_state=prng)
-dfs = []
-test = []
-train = []
-
-for csv in train_csvs:
-	df, fs, ls = input_old.load_csv(os.path.join(train_dir, csv), filter_no_labels=True, only_slice=False)
-
-	if df is None:
-		continue
-	if df.shape[0] == 0:
-		continue
-
-	train.append(df)
-
-for csv in test_csvs:
-	df2, fs2, ls2 = input_old.load_csv(os.path.join(test_dir, csv), filter_no_labels=True, only_slice=False)
-
-	if df2 is None:
-		continue
-	if df2.shape[0] == 0:
-		continue
-
-        df2['SOURCE_FILE'] = csv
-	test.append(df2)
-
-# print len(test)
-
-train = pd.concat(train)
-test = pd.concat(test)
-
-#print (len(test))
-#print (len(train))
-
-
-# print (df.shape)
-# print df
-classes = list(train.groupby(ls2))
-#print(ls)
-max_samples = max(len(c) for _, c in classes)
-train = pd.concat(c.sample(max_samples, replace=True) for _, c in classes)
-print train.shape
-#print (len(train))
-#print df.shape
-#print type(df)
-#list_keys = [ k for k in df ]
-#print list_keys
-# print samps
-
-#print sum(df['L-DidChange'].values)
-# print df['L-DidChange'].index
-
-
-
-train_samps = train.loc[:,FIRST_NON_SLICE_FEATURE:]
-print train_samps.shape
[…]

### 2018-08-23 learning/util.py

@@ -1,13 +1,0 @@
-import tensorflow as tf
-
-def variable_summaries(var, name):
-  """Attach a lot of summaries to a Tensor."""
-  with tf.name_scope('summaries'):
-    mean = tf.reduce_mean(var)
-    tf.summary.scalar('mean/' + name, mean)
-    with tf.name_scope('stddev'):
-      stddev = tf.sqrt(tf.reduce_mean(tf.square(var - mean)))
-    tf.summary.scalar('stddev/' + name, stddev)
-    tf.summary.scalar('max/' + name, tf.reduce_max(var))
-    tf.summary.scalar('min/' + name, tf.reduce_min(var))
-    tf.summary.histogram(name, var)

### 2018-08-23 pipeline/decisionpath.py

@@ -1,0 +1,111 @@
+import sys
+import joblib
+import numpy as np
+import csv
+import sklearn.externals.joblib
+
+
+file_path = sys.argv[3]
+encfile = sys.argv[1]
+enc = joblib.load(encfile)
+
+def load_csv(file_path):
+    ret = []
+    spans = []
+    with open(file_path, "r") as csvfile:
+        reader = csv.reader(csvfile)
+        for _, row in enumerate(reader):
+            ret.append(row[1:])
+            spans.append(row[0])
+        header = ret[0]
+        ret = ret[1:]
+        spans = spans[1:]
+    return ret, header, spans
+
+ret = {}
+nominals = ['F-Size', 'F-Size-P', 'F-Size-C1', 'F-Size-C2', 'F-Size-C3']
+ignore = ['X-UD', 'X-DU']
+v, header, spans = load_csv(file_path)
+categories = [i for i in range(len(header[2:])) \
+    if header[2:][i] not in nominals and header[2:][i] not in ignore]
+nominals = [i for i in range(len(header[2:])) if header[2:][i] in nominals]
+X = []
+X += [np.array([r[2:][i] for i in categories], object) for r in v]
+
+X = np.asarray(X)
+
+l = np.array(v)[:, :2].astype(float)
+n = np.array(v)[:, 2:][:, nominals]
+n[n == ''] = 0
+ret = np.hstack((l, enc.transform(np.array(v)[:, 2:][:, categories]).toarray(), n)).astype(float)
+
+feature_names = []
+for i, cates in enumerate(enc.categories_):
+    for cat in cates:
+        feature_names.append(header[2:][categories[i]] + '_' + str(cat))
+for i in nominals:
+    feature_names.append(header[2:][i])
+
+# print(ret)
+# print('------------------------------------------')
+# print(header)
+# print('------------------------------------------')
+# print(ret_categories)
+
+modelfile = sys.argv[2]
+estimator = sklearn.externals.joblib.load(modelfile)
+X_test = ret[:,2:]
+test_labels = ret[:,1]
+node_indicator = estimator.decision_path(X_test)
+
+
+n_nodes = estimator.tree_.node_count
+children_left = estimator.tree_.children_left
+children_right = estimator.tree_.children_right
+feature = estimator.tree_.feature
+threshold = estimator.tree_.threshold
+
+leave_id = estimator.apply(X_test)
+
+for ind, _ in enumerate(X_test):
[…]

### 2018-08-23 pipeline/makeHtml.py

@@ -2,9 +2,9 @@ import json, sys, ast, os
 from subprocess import run, PIPE
 import functools, html
 
-ML_DIR = 'foo'
+ML_DIR = 'v4-partial/vectors'
 ML_EXT = '.ml'
-CSV_DIR = 'goodCsvs'
+CSV_DIR = 'v4-partial/vectors/blah+context+slice+size'
 CSV_EXT = '.csv'
 
 HTML_BEGIN = '''<head>
@@ -20,6 +20,8 @@ Key:
 <br/>
 <div class="box change">change</div> / <div class="box">no change</div>
 <br/>
+<br/>
+<br/>
 '''
 HTML_FIX_BEGIN = '''<div class="container">'''
 HTML_END = '''
@@ -27,15 +29,6 @@ HTML_END = '''
 </div>
 </body>
 '''
-#<script src="https://ajax.googleapis.com/ajax/libs/jquery/2.1.3/jquery.min.js"></script>
-#<script src="foo.js"></script>
-# $(".box").children().hover(function () {
-#     $(this).parent().css("border-color", "grey");
-#     $(this).css("border-color", "red");
-# }, function () {
-#     $(this).parent().css("border-color", "");
-#     $(this).css("border-color", "");
-# });
 
 # TODO: handle changes outside of slice
 
@@ -72,6 +65,7 @@ def factComparator(fact1, fact2):
 def visualize(code, facts):
     code = code.split('\n')
     outBuffer = ""
+    z = 0
     for (row,line) in enumerate(code):
         for (col,char) in enumerate(line):
             for fact in facts:
@@ -84,11 +78,17 @@ def visualize(code, facts):
                     if confidence >= 0.5:
                         classes.append("blame")
                     color = int((1-confidence)*120)
-                    outBuffer += '<div class="%s" style="border-color:hsl(%d, 100%%, 50%%)" title="%f\n%s">' % (" ".join(classes),color,confidence,rules)
+                    z += 1
+                    if z > 9:
+                        print("warning: can't handle z > 9")
+                    classes.append("b%d" % z)
+                    tooltipText = "Confidence: %f<br/>%s" % (confidence,'<br/>'.join(rules.split('\n')))
+                    outBuffer += '<div class="%s" style="border-color:hsl(%d, 100%%, 50%%)"><span class="span%d">%s</span>' % (" ".join(classes),color,z,tooltipText)
[…]

### 2018-08-23 paper/main.tex
graph updates

@@
 \begin{figure}
+
+\caption{\fixme{This figure has been removed; this is a placeholder so the
+numbering doesn't change while I handle some TODOs}}
+\label{fig-full-dataset-acc}
+\end{figure}
+
+\begin{figure}
 \begin{tikzpicture}
@@
     symbolic x coords={Baseline,Top-1,Top-2,Top-3},
-    xtick=data,
-    ymin=0,
-    ylabel=Model Accuracy,
-    enlarge x limits=0.3,
-    legend style={at={(0.5,-0.15)},anchor=north}
-]
-    \addplot table[x=scoreName, y=score, col sep=comma]{fault-localization.csv};
-\end{axis}
-\end{tikzpicture}
-\caption{Fault localization accuracy. 
-Baseline is the normal Python interpreter. Other bars represent our
-approach, using decision trees, 
-on the largest dataset (272534 pairs).}
-\label{fig-full-dataset-acc}
-\end{figure}
-
-\begin{figure}
-\begin{tikzpicture}
-\begin{axis}[
-    ybar,
-    symbolic x coords={Baseline,Top-1,Top-3},
     xtick=data,
@@
 our approach, using random forests,
-on the largest dataset (272534 pairs). \fixme{get top-2, then change text
-to discuss this figure instead of decision-tree version}}
+on the largest dataset (272534 pairs).}
 \label{fig-full-dataset-acc-random-forest}
@@
 
-We first use the full dataset to train and test decsion trees optimized for
+We first use the full dataset to train and test decsion forests optimized for
 Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc},
-these produce a correct answer 55, 68, and
-75\% of the time (to two significant figures). As a baseline, we determine that
+these produce a correct answer 60, 72, and
+79\% of the time (to two significant figures). As a baseline, we determine that
 the expression blamed by the standard Python interpreter (i.e. the one that
@@
     ybar,
-    symbolic x coords={All, TypeError, AttributeError},
+    symbolic x coords={All, TypeError, AttributeError, ValueError},
     xtick=data,

### 2018-08-23 paper/main.tex
some more updates

@@
 
-\emph{Logistic Regression}
-The simplest classifier we investigate is logistic regression:
-a linear model where the goal is to learn a set of weights $W$
-that describe the following model for predicting a label
-$b$ from a feature vector $v$:
-%
-\[ \Pr(b = 1 | v) = \frac{1}{1 + e^{-W^{\top} v}} \]
-%
-The weights $W$ are learnt from training data, and the value of
-$\Pr(b | v)$ naturally leads to a confidence score $\mathcal{C}$.
-%
-Logistic regression is a widely used classification algorithm, preferred
-for its simplicity, ease of generalization, and interpretability.
+\fixme{how the data is represented}
+
+\fixme{is program pair the correct term?}
+
+For each line of a program pair, we can extracted the features mentioned in Section
+\ref{sec-features}.
+By performing one-hot encoding on categorical features, we can represent the
+extracted features into a length $M$ feature vector $\mathbf{x} \in
+\mathcal{R}^{M}$.
+Let the number of examples be $N$.
+The problem can be formulated as a binary classification problem.
+Given a data set $\mathcal{D} \in \{\mathbf{x}^{(i)}, y^{(i)}\}^{N}_{i=1}$,
+where $\mathbf{x}^{(i)} \in \mathcal{R}^{M}$ and $y^{(i)} \in \{0, 1\}$,
+a learning model will learn a function $f: \mathbf{R}^{M} \to \{0, 1\}$.
+Due to the large amount of data provided in this task, we only considers the
+following models which are scalable to large data set.
+
+\paragraph{Logistic Regression}
+Logistic regression~(LR) is a popular linear model for classification task.
+LR models the probability of a label $b$ being positive as
+\[ \Pr(b = 1 | \mathbf{x}) = \frac{1}{1 + e^{-\mathbf{w}^{\top} \mathbf{x}}} \]
+Where $\mathbf{w} \in \mathcal{R}^M$ is a weight vector and
+$\mathbf{x} \in \mathcal{R}^M$ is the feature vector to be predicted.
+LR learns a weight vector $\mathbf{w}$ by minimizing the binary crossentropy loss
+over the given data set $\mathcal{D}$.
+The prediction value of $\Pr(b | v)$ naturally leads to a confidence score.
+%$\mathcal{C}$.
+%
+%Logistic regression is a widely used classification algorithm, preferred
+%for its simplicity, ease of generalization, and interpretability.
 %
@@
 %
-While this can be somewhat mitigated by adding higher order (quadratic
-or cubic) features, this often requires substantial domain knowledge.
-
-\emph{Decision Tree}
-Decision tree algorithms learn a tree of binary predicates over the
-features, recursively partitioning the input space until a final
-classification can be made.
-%
-Each node in the tree contains a single predicate of the form
-$v_j \leq t$ for some feature $v_j$ and threshold $t$, which determines
+%While this can be somewhat mitigated by adding higher order (quadratic or
+%cubic) features, this often requires substantial domain knowledge and
+%computation.
+
+\paragraph{Decision Tree}
+Decision tree~(DT) learns a tree of binary predicates over the features,
+recursively partitioning the feature space until a final classification can be
+made.
+%
+Each node in the tree contains a single thresholding
+$\mathbf{x}_i \leq t$ for some feature $\mathbf{x}_i$ and threshold $t$, which determines
 whether a given input should proceed down the left or right subtree.
@@
 %
-This leads to a prediction rule that can be quite expressive depending
-on the data used to build it.
-
-Training a decision tree entails finding both a set of good partitioning
-predicates and a good ordering of the predicates based on data.
-%
-This is usually done in a top-down greedy manner, and there are several
-standard training algorithms such as C4.5 \cite{Quinlan1993-de} and
-CART \cite{Breiman1984-qy}.
-
-Another advantage of decision trees is their ease of interpretation ---
-the decision rule is a white-box model that can be readily described to
-a human, especially when the tree is small.
-%
-However, the main limitation is that these trees often do not generalize
-well, though this can be somewhat mitigated by \emph{pruning} the tree.
-
-\emph{Random Forests}
-%
-Random forests (RF)~\cite{breiman2001random} improve generalization by training an
-\emph{ensemble} of distinct decision trees and using a majority
+This leads to a non-linear prediction rule which can be more expressive
+then linear model like LR depending on the data used to build it.
+%Training a decision tree entails finding both a set of good partitioning
+%predicates and a good ordering of the predicates based on data.
+%
+The training of a decision tree is usually done in a top-down greedy manner,
+and there are several implementations of DT such as C4.5
+\cite{quinlan2014c4} and CART \cite{breiman2017classification}.
+
+%Another advantage of decision trees is their ease of interpretation ---
+%the decision rule is a white-box model that can be readily described to
+%a human, especially when the tree is small.
+%%
+%However, the main limitation is that these trees often do not generalize
+%well, though this can be somewhat mitigated by \emph{pruning} the tree.
+
+\emph{Random Forest}
+%
+Random forest (RF)~\cite{breiman2001random} improve generalization by
+training an \emph{ensemble} of distinct decision trees and using a majority
 vote to make a prediction.
@@
 %
-The agreement among the trees forms a natural
-confidence score.
+The agreement among the trees forms a natural confidence score.
 %
@@
 also makes the learned model more difficult to interpret.
-\emph{Random Forest}
-
-\emph{Neural Networks}
+
+\paragraph{Neural Networks}
 %
 The last (and most complex) model we use is a type of neural network
-called a \emph{multi-layer perceptron} (see \cite{Nielsen2015-pu} for
+called a \emph{multi-layer perceptron~(MLP)} (see \cite{Nielsen2015-pu} for
 an introduction to neural networks).
 %
-A multi-layer perceptron can be represented as a directed acyclic
-graph whose nodes are arranged in layers that are fully connected by
-weighted edges.
+A MLP can be represented as a directed acyclic graph whose nodes are arranged
+in layers that are fully connected by weighted edges.
 %
@@
   unit} (ReLU), defined as $g(x) = \mathsf{max}(0,x)$
-\cite{Nair2010-xg}.
+\cite{nair2010rectified}.
 %
@@
 intended score (Top-k accuracy).
-For RF, the maximum depth of each tree is searched from $25$ to $35$.
+For RF, the maximum depth of each tree is searched from $25$ to $35$ and the
+number of tree are fixed to $500$.
 For decision tree maximum depth of the tree is searched from $25$ to $35$ with interval of $5$,
@@
 $10^{-3}$ with multiple of $10^{-2}$ and the minimum samples in each leaf is set to $200$.
-For logistic regression and MLP, binary crossentropy is set as the loss
+For logistic regression and MLP, binary cross-entropy is set as the loss
 function and Adam \cite{kingma2014adam} optimizer is used.
-Learning rate is set to $0.1$.
+Learning rate is set to $0.1$ and the number of epochs are $10$.
 L2 regularization is applied to the weights of both models,
@@
 Gini importance (or mean decrease in impurity), a common measure
-for decision tree and random forest models~\cite{FIXME}. Informally, the
+for decision tree and random forest models~\cite{breiman2001random}. Informally, the
 Gini importance conveys a weighted count of the number of times a feature

### 2018-08-23 paper/main.tex
remove descriptions of unused model types

@@
 \fixme{
-\emph{Models} The following descriptions of Logistic Regression, Decision Trees,
-Random Forests, and MLP will have to be redone as they are copied from Learning to Blame.
+\emph{Models} The following descriptions of Decision Trees and
+Random Forests will have to be redone as they are copied from Learning to Blame.
 }
@@
 
-\paragraph{Logistic Regression}
-Logistic regression~(LR) is a popular linear model for classification task.
-LR models the probability of a label $b$ being positive as
-\[ \Pr(b = 1 | \mathbf{x}) = \frac{1}{1 + e^{-\mathbf{w}^{\top} \mathbf{x}}} \]
-Where $\mathbf{w} \in \mathcal{R}^M$ is a weight vector and
-$\mathbf{x} \in \mathcal{R}^M$ is the feature vector to be predicted.
-LR learns a weight vector $\mathbf{w}$ by minimizing the binary crossentropy loss
-over the given data set $\mathcal{D}$.
-The prediction value of $\Pr(b | v)$ naturally leads to a confidence score.
-%$\mathcal{C}$.
-%
-%Logistic regression is a widely used classification algorithm, preferred
-%for its simplicity, ease of generalization, and interpretability.
-%
-Its main limitation is that the prediction rule is constrained to be a
-linear combination of the features, and hence relatively simple.
-%
-%While this can be somewhat mitigated by adding higher order (quadratic or
-%cubic) features, this often requires substantial domain knowledge and
-%computation.
-
-\paragraph{Decision Tree}
+\emph{Decision Tree}
 Decision tree~(DT) learns a tree of binary predicates over the features,
@@
 
-\paragraph{Neural Networks}
-%
-The last (and most complex) model we use is a type of neural network
-called a \emph{multi-layer perceptron~(MLP)} (see \cite{Nielsen2015-pu} for
-an introduction to neural networks).
-%
-A MLP can be represented as a directed acyclic graph whose nodes are arranged
-in layers that are fully connected by weighted edges.
-%
-The first layer corresponds to the input features, and the final to the
-output.
-%
-%The output of a node $v$ in an internal layer is given by:
-The output of an internal node $v$ is
-%
-\[ h_v = g(\sum_{j \in N(v)} W_{jv} h_j ) \]
-%
-where $N(v)$ is the set of nodes in the previous layer that are adjacent
-to $v$, $W_{jv}$ is the weight of the $(j, v)$ edge and $h_j$ is the
-output of node $j$ in the previous layer.
-%
-Finally $g$ is a non-linear function, called the activation function,
-which in recent work is commonly chosen to be the \emph{rectified linear
-  unit} (ReLU), defined as $g(x) = \mathsf{max}(0,x)$
-\cite{nair2010rectified}.
-%
-The number of layers, the number of neurons per layer, and the
-connections between layers constitute the \emph{architecture} of a
-neural network.
-%
-In this work, we use relatively simple neural networks which have an
-input layer, a single hidden layer and an output layer.
-
-A major advantage of neural networks is their ability to discover
-interesting combinations of features through non-linearity, which
-significantly reduces the need for manual feature engineering, and
-allows high expressivity.
-%
-On the other hand, this makes the networks particularly difficult to
-interpret and also difficult to generalize unless vast amounts of
-training data are available.
-
 \emph{Training methodology} At this point we have feature vectors describing every term of every program in
@@
 
-\paragraph{Hyper-parameters}
+\emph{Hyper-parameters}
 All hyper-parameters are tuned using three-fold cross-validation on the

### 2018-08-23 paper/main.tex
update graph

@@
     ybar,
-    symbolic x coords={All, TypeError, AttributeError, ValueError},
+    symbolic x coords={baseline, Type, Attribute, Value, Key},
     xtick=data,
@@
     ylabel=Model Accuracy,
-    enlarge x limits=0.5,
     legend style={at={(0.5,-0.15)},anchor=north}

### 2018-08-23 paper/main.tex
feature importance tables

@@
 Name                      & Category                & Gini Importance \\ \bottomrule
-Size-P & Contextual (Syntactic)               & 0.112 \\
-Stmt-Constr-C3= & Contextual (Syntactic)      & 0.061 \\
-Expr-Constr-P=List & Contextual (Syntactic)   & 0.055 \\
-PythonBlame & Dynamic (Error location)        & 0.039 \\
-Type=unknown & Dynamic (Type)                 & 0.037 \\
-Type-P=unknown & Contextual (Type)            & 0.037 \\
-PythonMsg=IndexError & Dynamic (Error message)& 0.033 \\
+P size                    & Contextual (Syntactic)  & 0.112 \\
+C3 is statement           & Contextual (Syntactic)  & 0.061 \\
+P is list literal         & Contextual (Syntactic)  & 0.055 \\
+Program crashes here      & Dynamic (Error location)& 0.039 \\
+Type is unknown           & Dynamic (Type)          & 0.037 \\
+P type is unknown         & Contextual (Type)       & 0.037 \\
+Err message is IndexError & Dynamic (Error message) & 0.033 \\
 Size                      & Syntactic               & 0.028 \\
-Expr-Constr-P=Dictionary & Contextual (Syntactic) & 0.027 \\
-Size-C1 & Contextual (Syntactic)              & 0.024 \\
-Expr-Constr=Var & Syntactic                   & 0.020 \\
-Type-C1=unknown & Contextual (Type)           & 0.020 \\
+P is dictionary literal   & Contextual (Syntactic)  & 0.027 \\
+C1 size                   & Contextual (Syntactic)  & 0.024 \\
+Is variable               & Syntactic               & 0.020 \\
+C1 type is unknown        & Contextual (Type)       & 0.020 \\
 \toprule
@@
 \caption{Feature predictive power (for a Top-3 Decision Tree
-learned on the entire dataset).}
+learned on the entire dataset). P and C1-C3 refer to the parent and
+first three children of the node in question.}
 \label{tab-feature-predictive-power}
@@
 Name                     & Category                & F-score \\ \bottomrule
-Stmt-Constr-C3= & Contextual (Syntactic) & 69 \\
+C3 is statement          & Contextual (Syntactic)  & 69 \\
 Size                     & Syntactic               & 63 \\
-PythonBlame & Dynamic (Error location)   & 50 \\
-Type-P=LIST & Contextual (Type)          & 36 \\
-Type-C1=bool & Contextual (Type)         & 30 \\
-Type-P=unknown & Contextual (Type)       & 28 \\
-Type-C1= & Contextual (Type)             & 25 \\
-Size-C3 & Contextual (Syntactic)         & 20 \\
-Op-Constr-C1= & Contextual (Syntactic)   & 20 \\
-Size-C1 & Contextual (Syntactic)         & 18 \\
-Type-P=DICT & Contextual (Type)          & 17 \\
-Type=str & Dynamic (Type)                & 12 \\
+Program crashes here     & Dynamic (Error location)& 50 \\
+P type is list           & Contextual (Type)       & 36 \\
+C1 type is bool          & Contextual (Type)       & 30 \\
+P type is unknown        & Contextual (Type)       & 28 \\
+C1 is expression         & Contextual (Type)       & 25 \\
+C3 size                  & Contextual (Syntactic)  & 20 \\
+C1 is binary or unary op & Contextual (Syntactic)  & 20 \\
+C1 size                  & Contextual (Syntactic)  & 18 \\
+P type is dictionary     & Contextual (Type)       & 17 \\
+Type is string           & Dynamic (Type)          & 12 \\
 \toprule

### 2018-08-24 paper/main.tex
update

@@
 
-For each line of a program pair, we can extracted the features mentioned in Section
-\ref{sec-features}.
+To apply machine learning to analyze the data, we have to formulate the 
+problem into standard classification problem.
+For each line of a program pair, we extracted the features mentioned in
+Section \ref{sec-features}.
 By performing one-hot encoding on categorical features, we can represent the
@@
 \mathcal{R}^{M}$.
+For the corresponding label, is represented as $1$ and $0$, which means whether
+the line blamed or not.
 Let the number of examples be $N$.
-The problem can be formulated as a binary classification problem.
 Given a data set $\mathcal{D} \in \{\mathbf{x}^{(i)}, y^{(i)}\}^{N}_{i=1}$,
@@
 a learning model will learn a function $f: \mathbf{R}^{M} \to \{0, 1\}$.
-Due to the large amount of data provided in this task, we only considers the
+Due to the large amount of data provided in this task, we only consider the
 following models which are scalable to large data set.
@@
 \emph{Decision Tree}
-Decision tree~(DT) learns a tree of binary predicates over the features,
-recursively partitioning the feature space until a final classification can be
-made.
+Decision tree~(DT) learns a tree of binary predicates over the features and
+recursively partitioning the feature space until a final classification can
+be made.
 %
-Each node in the tree contains a single thresholding
-$\mathbf{x}_i \leq t$ for some feature $\mathbf{x}_i$ and threshold $t$, which determines
-whether a given input should proceed down the left or right subtree.
+Each node in the tree contains a single thresholding $\mathbf{x}_i \leq t$, where
+applying threshold $t \in \mathcal{R}$ on the $i$-th feature of $\mathbf{x}$,
+which determines whether a given input should proceed down the left or right
+subtree.
 %
@@
 %
-Random forest (RF)~\cite{breiman2001random} improve generalization by
-training an \emph{ensemble} of distinct decision trees and using a majority
-vote to make a prediction.
-For each tree, RF selects a random sample with replacement of the training
-set and trains the tree with these samples.
+Random forest (RF)~\cite{breiman2001random} improves generalization by
+learning an \emph{ensemble} of decision trees.
+For prediction, each tree in the forest votes on one of the classes.
+The percentage of vote for each class forms a natural confidence score.
+During training, RF selects a subset of training data by sampling with
+replacement and use the subset to train each of the tree.
+By doing so, RF is able to generate a forest with each tree trained on
+a different subset of data.
+The diversity of the underlying models tends to make RF less susceptible to
+the overfitting, but it also makes the learned model more difficult to
+interpret.
 %
-The agreement among the trees forms a natural confidence score.
+%Since each classifier in the ensemble is a decision tree, this still
+%allows for complex and expressive classifiers.
+
+%The training process involves taking $N$ random subsets of the training
+%data and training a separate decision tree on each subset --- the
+%training process for the decision trees is often modified slightly to
+%reduce correlation between trees, by forcing each tree to pick features
+%from a random subset of all features at each node.
 %
-Since each classifier in the ensemble is a decision tree, this still
-allows for complex and expressive classifiers.
-
-The training process involves taking $N$ random subsets of the training
-data and training a separate decision tree on each subset --- the
-training process for the decision trees is often modified slightly to
-reduce correlation between trees, by forcing each tree to pick features
-from a random subset of all features at each node.
-%
-The diversity of the underlying models tends to make random forests less
-susceptible to the overfitting, but it
-also makes the learned model more difficult to interpret.
-
-\emph{Training methodology} At this point we have feature vectors describing every term of every program in
-our dataset. The vectors from a random 30\% of the program pairs are set aside for testing and
+
+\paragraph{Training methodology}
+At this point we have feature vectors describing every term of every program in
+our data set. The vectors from a random 30\% of the program pairs are set aside for testing and
 the other 70\% are used for training.
@@
 to be changed.
-For a given k, we score the model based on Top-k accuracy: for what
+For a given $k$, we score the model based on Top-$k$ accuracy: for what
 proportion of the programs is a correct answer (i.e. a term that is actually
-changed) present in the top k results.
+changed) present in the top $k$ results.
 We retrain the models for each value of k, since a model optimized to produce
@@
 
-\emph{Hyper-parameters}
+\paragraph{Hyper-parameters}
 All hyper-parameters are tuned using three-fold cross-validation on the
-intended score (Top-k accuracy).
+intended score (Top-$k$ accuracy).
 For RF, the maximum depth of each tree is searched from $25$ to $35$ and the
@@
 
-In each case the most effective model we trained was a decision tree. (The other
-models we tried were Logistic Regression and MLP. Yao-Yuan - anything to add to
-this section? in particular, is it actually accurate to say we're doing
-cross-validation - the 3-fold thing we do is actually just to select parameters,
-and then there's no cross-validation on the final model, right?
-%
-Correct
-)
+%In each case the most effective model we trained was a decision tree. (The other
+%models we tried were Logistic Regression and MLP. Yao-Yuan - anything to add to
+%this section? in particular, is it actually accurate to say we're doing
+%cross-validation - the 3-fold thing we do is actually just to select parameters,
+%and then there's no cross-validation on the final model, right?
+%%
+%Correct
+%)
 

### 2018-08-24 paper/main.tex
qualitative evaluation examples; graph

@@
     ybar,
-    symbolic x coords={baseline, Type, Attribute, Value, Key},
+    symbolic x coords={baseline, Type, Attribute, Value, Key, Index},
     xtick=data,
@@
 
+\fixme{
 Pick out one example where we did poorly. Walk through why it is ``out of
@@
 future work.
-
-Pick out a few examples where we did well and the baseline did poorly. Walk
-through how interesting and indicative they are and explain why we did well.
+}
+
+%%% Original for fig-win-example-1; index 19615 of v4
+% def word_checker(food):
+%     w_alphabet_ned = ['a', 'b', 'c', 'd', 'e', 'f', 'g', 'h', 'i', 'j',
+%                 'k', 'l', 'm', 'n', 'o', 'p', 'q', 'r', 's', 't',
+%                 'u', 'v', 'w', 'x', 'y', 'z', 'ë', 'ï', 'é', 'è', 'ö', 'ê',
+%                 'ü', 'ó', 'ç', 'á', 'à', 'ä', 'û', 'î', 'í', 'ô', 'ú', 'ñ', 'â']
+%
+%     w_list_food = list(food)
+%
+%     for letters in w_list_food:
+%         if letters not in w_alphabet_ned:
+%             w_list_food = w_list_food.remove(letters)
+%         food = ''.join(w_list_food)
+%     print(food)
+%
+%
+%
+%
+% word_checker('sl^* ( ) 0a')
+
+\begin{figure}
+\begin{lstlisting}
+def devowel(word):
+    word = 'foo'
+    w_list = list(word)
+    vowels = ['a', 'e', 'i', 'o', 'u']
+    for letter in w_list:
+        if letter not in vowels:
+            w_list = w_list.remove(letter)
+    word = ''.join(w_list)
+print(devowel('foo'))
+\end{lstlisting}
+\caption{A program with a bug on line 7.}
+\label{fig-win-example-one}
+\end{figure}
+
+%%% Original for fig-win-example-2; index 50219 of v4
+% GeorgiaAreaCodes = [404, 770, 678, 706, 762, 229, 423, 912, 478]
+% for i in range(0, len(GeorgiaAreaCodes) + 1):
+%     print(GeorgiaAreaCodes[i])
+
+\begin{figure}
+\begin{lstlisting}
+areaCodes = [800, 555]
+for i in range(0, len(areaCodes) + 1):
+    print(areaCodes[i])
+\end{lstlisting}
+\caption{A program with an off-by-one bug.}
+\label{fig-win-example-two}
+\end{figure}
+
+We now go through a couple examples where our method does well. These examples
+are adapted slightly for simplicity of
+presentation and to protect the anonymity of the programmers, but they retain
+the essential character of the originals from our dataset.
+
+\paragraph{NoneType}
+The function in \figref{win-example-one} attempts to remove all vowels from a
+given word. However the
+\lt{remove} method in line 7 actually modifies the list in place and returns
+\lt{None}, so the correct version forgoes the assignment and just has
+\lt{w_list.remove(letter)}. Python does not crash until line 8, where its
+message is \lt{TypeError: can only join an iterable}. However, our decision
+tree flags the correct statement, using as some of its rules that it is an
+assignment statement whose second child has type \lt{NoneType}. This seems to
+capture the intuitive rule that there is rarely a good reason to assign the
+value \lt{None} to a variable. Note that these rules use both the syntactic
+form of the statement (an assignment) and the dynamic type of one of its
+children, so all our categories of features - syntactic, dynamic, and
+contextual - were useful here.
+
+\paragraph{Off-by-one bugs}
+In \figref{win-example-two}, the programmer incorrectly adds one to the high
+end of their range, causing an \lt{IndexError: list index out of range} on line
+3 during the final iteration of the for loop. The correct expression to blame
+is the addition \lt{len(areaCodes) + 1}. Some of the rules that our
+decision tree uses to correctly localize this bug include that the error is an
+index error, the type of the parent, and the fact that it's an addition
+expression. Thus once again we see all three categories of features being used.
 

### 2018-08-24 pipeline/make_trace/anfToSlice.py

@@ -1,16 +1,15 @@
 from subprocess import run, PIPE, CalledProcessError, TimeoutExpired
 from process_raw import process_one
-from collections import defaultdict
 import json
 import ast
 
 ANF_BIN="../../.stack-work/install/x86_64-osx/lts-8.14/8.0.2/bin/make-anf-raw"
 ANF_TIMEOUT=3
 
-def contains(superSpan, subSpan):
-    (a,b,c,d) = superSpan
-    (w,x,y,z) = subSpan
-    return (a,b) <= (w,x) and (y,z) <= (c,d)
+# def contains(superSpan, subSpan):
+#     (a,b,c,d) = superSpan
+#     (w,x,y,z) = subSpan
+#     return (a,b) <= (w,x) and (y,z) <= (c,d)
 
 def getSpan(line):
     parts = line.split("#  interaction:")

### 2018-08-24 paper/main.tex
graphs

@@
     ybar,
-    symbolic x coords={noContext, noTypes, all},
+    symbolic x coords={No Context, No Types},
     xtick=data,
     ymin=0,
-    ylabel=Model Accuracy,
+    ymax=1.1,
+    ylabel=Normalized Model Accuracy,
     enlarge x limits=0.5,
@@
 ]
-    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2.csv};
+    \addplot table[x=scoreName, y=score, col sep=comma]{removing-features-2-normalized.csv};
+    \draw [red] ({rel axis cs:0,0}|-{axis cs:No Types,1}) -- ({rel axis cs:1,0}|-{axis cs:No Types,1}) node [pos=0.33, above] {};
 \end{axis}
 \end{tikzpicture}
-\caption{Accuracy when features are removed, based on decision trees on a
-random subset of 20000 program pairs.
-\fixme{The x-axis labels are not perfectly clear: does ``all'' mean that
-everything is still in or that everything is out?}
-\fixme{Indicate that this is a top-1 model.}
-\fixme{Normal the y-axis so that ``all'' is 1.0 to make the subset effects
-more evident.}
+\caption{Accuracy when features are removed, based on Top-1 decision trees on a
+random subset of 20000 program pairs, as compared to a Top-1 decision tree trained
+on all features.
 }
@@
     ybar,
-    symbolic x coords={baseline, Type, Attribute, Value, Key, Index},
+    symbolic x coords={Index, Value, Key, Type, Attribute},
     xtick=data,
     ymin=0,
-    ylabel=Model Accuracy,
+    ylabel=Normalized Model Accuracy,
     legend style={at={(0.5,-0.15)},anchor=north}
 ]
-    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2.csv};
+    \addplot table[x=scoreName, y=score, col sep=comma]{defect-categories-2-normalized.csv};
+    \draw [red] ({rel axis cs:0,0}|-{axis cs:Index,1}) -- ({rel axis cs:1,0}|-{axis cs:Index,1}) node [pos=0.33, above] {};
 \end{axis}
+% \begin{axis}
+%     \addplot{1};
+% \end{axis}
 \end{tikzpicture}
-\caption{Accuracy when the model is trained and tested on only the data exhibiting
-the error on the x-axis.
+\caption{Accuracy when a Top-1 decision tree is trained and tested on only the data exhibiting
+the error on the x-axis, as compared to a Top-1 decision tree on all the data.
 \fixme{Spin this for generality. Indicate that our model gives high-quality

### 2018-08-24 paper/main.tex
graphs
paper

@@
 
-\paragraph{Training methodology}
+\emph{Training methodology}
 At this point we have feature vectors describing every term of every program in
@@
 \textsc{scikit-learn}~\cite{scikit-learn}.
-Due to the large amount of data points, we did not use \textsc{scikit-learn}'s
-implementation for logistic regression and multilayer perceptron.
-We implementation both model using
-\textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.
-
-Because of the imbalance nature of the data set, where not-blamed lines
-(negative examples) appears more frequently than blamed lines (positive
-examples).
-The weight for each example during training is re-weighted to the reciprocal
+% Due to the large amount of data points, we did not use \textsc{scikit-learn}'s
+% implementation for logistic regression and multilayer perceptron.
+% We implementation both model using
+% \textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.
+
+Unsurprisingly this is an imbalanced dataset in that negative examples, i.e.
+terms that are not changed by the programmer, are much more common than
+positive examples, so during training we re-weight each example to the reciprocal
 of the frequency of corresponding class.
 
-\paragraph{Hyper-parameters}
+\emph{Hyper-parameters}
 All hyper-parameters are tuned using three-fold cross-validation on the
@@
 
-\item[RQ3]{Which features are the most important for various categories of
-Python defects?}
+% \item[RQ3]{Which features are the most important for various categories of
+% Python defects?}
+
+\item[RQ3]{How well does our algorithm handle different kinds of Python errors?}
 
@@
 \subsection{RQ 1 --- Fault Localization Accuracy}
-
-\begin{figure}
-
-\caption{\fixme{This figure has been removed; this is a placeholder so the
-numbering doesn't change while I handle some TODOs}}
-\label{fig-full-dataset-acc}
-\end{figure}
 
@@
 We first use the full dataset to train and test decsion forests optimized for
-Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc},
+Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc-random-forest},
 these produce a correct answer 60, 72, and
@@
 the expression blamed by the standard Python interpreter (i.e. the one that
-causes it to crash) is only changed by the user 30\% of the time. Thus our
+causes it to crash) is only changed by the user 45\% of the time. Thus our
 most directly comparable model, the one for Top-1, significantly outperforms
@@
 on all features.
+\fixme{will add No Syntax bar}
 }
@@
 \end{axis}
-% \begin{axis}
-%     \addplot{1};
-% \end{axis}
 \end{tikzpicture}
@@
 the error on the x-axis, as compared to a Top-1 decision tree on all the data.
-\fixme{Spin this for generality. Indicate that our model gives high-quality
-fault localization regardless of defect type. To put it another way, we
-can view this as a sensitivity analysis: our algorithm is not sensitive to
-defect type (that's good!).}
 }
-\label{fig:defect-categories}
+\label{fig-defect-categories}
 \end{figure}
 
-prereq: define defect category. This could be either the raw python
-exception name or it could be clusters that we have manually created.
-Ranjit notes: result could be ``we need everything for everything'', at
-which point this is useless and should be skipped. Per Ben's exam: may find
-that for different categories of errors we should use different ML
-classifiers.
-
-PLOT Type: Bar Graph (three bars per point on the X axis)
-
-PLOT Y Axis: Leave-One-In Accuracy (normalized per X axis point)
-Ranjit notes: on smaller subsets we could rebuild the classifier each time.
-Wes notes that if we do, it takes us longer. If we re-use the monolithic
-model we can do this very quickly.
-
-PLOT X AXis: Defect Categories
-
-PLOT Bars: Static, Dynamic, Context
-
-Goal: Demonstrate that each category of feature is essential for a certain
-piece of the problem. We need ABC feature to handle DEF class of defects,
-but ABC feature does not work well on GHI class of defects, for those we
-need JKL features. Implicit: we were smart for including all of these
-categories.
+We next try training and testing Top-1 decision trees on only the subsets of
+the data
+corresponding to each of the five most common exceptions: \lt{TypeError},
+\lt{IndexError}, \lt{AttributeError}, \lt{ValueError}, and \lt{KeyError}.
+As shown in \figref{defect-categories}, these models have accuracy between
+86\% and 115\% of a comparable model trained on the data as a whole. This shows
+that our algorithm is robust, able to give high accuracy fault localizations on
+a wide variety of defect types.
+
+% prereq: define defect category. This could be either the raw python
+% exception name or it could be clusters that we have manually created.
+% Ranjit notes: result could be ``we need everything for everything'', at
+% which point this is useless and should be skipped. Per Ben's exam: may find
+% that for different categories of errors we should use different ML
+% classifiers.
+
+% PLOT Type: Bar Graph (three bars per point on the X axis)
+
+% PLOT Y Axis: Leave-One-In Accuracy (normalized per X axis point)
+% Ranjit notes: on smaller subsets we could rebuild the classifier each time.
+% Wes notes that if we do, it takes us longer. If we re-use the monolithic
+% model we can do this very quickly.
+
+% PLOT X AXis: Defect Categories
+
+% PLOT Bars: Static, Dynamic, Context
+
+% Goal: Demonstrate that each category of feature is essential for a certain
+% piece of the problem. We need ABC feature to handle DEF class of defects,
+% but ABC feature does not work well on GHI class of defects, for those we
+% need JKL features. Implicit: we were smart for including all of these
+% categories.
 
@@
 
-\subsection{RQ 5 --- Wishlist: Partitioning Kinds of Errors}
-
-Can we partition the kinds of errors that people make? What are common
-misconceptions? Do these correspond to paths of decision trees.
-
-If we don't get to the source-sink thing, it could go here.
-
-I don't think we have time, in this paper, to actually implement a
-two-level machine learning.
+% \subsection{RQ 5 --- Wishlist: Partitioning Kinds of Errors}
+
+% Can we partition the kinds of errors that people make? What are common
+% misconceptions? Do these correspond to paths of decision trees.
+
+% If we don't get to the source-sink thing, it could go here.
+
+% I don't think we have time, in this paper, to actually implement a
+% two-level machine learning.
 

### 2018-08-24 paper/main.tex
introduction, etc.

@@
 dynamic type errors, and may thus struggle to interpret and localize
-them~\cite{FIXME,FIXME}. We propose a machine learning approach based on
-static, dynamic and contextual information that effectively localizes
-beginner type errors.
+them~\cite{orso-parnin,Zhang2014-lv,Christiansen2014-qc,Pavlinovic2014-mr,Chen2014-gd,Neubauer2003-xv,Stuckey2004-ne}.
+We propose a machine learning approach based on static, dynamic and
+contextual information that effectively localizes beginner type errors.
 
@@
 dynamic languages. First, relying on the extant language interpreter
-requires test inputs~\cite{FIXME}, may involve error messages that are
-difficult for novices to understand~\cite{FIXME}, and may misleadingly
-implicate the symptom rather than the cause~\cite{FIXME}.  Second, a
+requires test inputs
+(cf.~\cite{Godefroid2005-am,Naylor2007-mi,Cadar2008-kg,PAcheco2007-at}),
+may involve error messages that are difficult for novices to
+understand~\cite{Marceau2011-ok,Marceau2011-cy,Nienaltowski2008-bv,Jadud2006-ly,Ishii2014-nf},
+and may misleadingly implicate the symptom rather than the
+cause~\cite{zeller05,Jose2011}.  Second, a
 significant amount of development is carried out via IDE or web interfaces,
@@
 Third, attempts to retrofit static type systems into dynamic languages,
-such as TypeScript for JavaScript~\cite{FIXME} and FIXME for
-Python~\cite{FIXME} are promising but not yet as widely
-deployed~\cite{FIXME}. Finally, research tools and algorithms for fault
+such as TypeScript~\cite{typescript2014,refscriptpldi16,mollerOOPSLA14} for
+JavaScript and various
+gradual and static type approaches for 
+Python~\cite{mypy,gradual-typing-python} or
+Ruby~\cite{static-typing-ruby,an11,foster09,foster11}
+are promising but not always widely
+deployed. Finally, research tools and algorithms for fault
 localization either require rich static types (e.g., as in
-Mycroft~\cite{FIXME}, Nate~\cite{FIXME} or Sherrloc~\cite{FIXME}), require
-test cases (e.g., as in Tarantula~\cite{FIXME}), or produce voluminous
+Mycroft~\cite{FIXME}, Nate~\cite{learning-to-blame} or Sherrloc~\cite{FIXME}), require
+test cases (e.g., as in Tarantula~\cite{tarantula}), or produce voluminous
 ranked output lists that have been found to be unhelpful in general and
@@
 can use large beginner datasets, such as those from
-PythonTutor.com~\cite{Guo2012-vu}, to assess scalability.
+PythonTutor.com~\cite{Guo2013-vu}, to assess scalability.
 
@@
 \item We present an algorithm for accurately localizing dynamic type errors
-in Python. We find that FIXME.
+in Python. We find that our approach is 60--79\% accurate, compared to
+the interpreter baseline of 45\%.
 
@@
 features for generalizing machine-learning fault localization to
-beginner-written Python fragments. We find that FIXME.
+beginner-written Python fragments. We find such features are all critical
+to our model, but that our model is not sensitive to error type and
+performs similarly across defect classes.
 
 \item We analyze the results of a thorough evaluation on over 270,000 real
-user instances, demonstrating the scalability of our algorithm.
+user instances, show that it is a more heterogeneous dataset than those used
+in previous work, and demonstrate the scalability of our algorithm.
 

### 2018-08-24 paper/main.tex
weimer pass continues

@@
 \section{Algorithm Overview}
+\label{sec-algorithm}
 
@@
     \item \lt{Statement} --- Given to all statements, since our terms
-    include both statments and expressions but only expressions
+    include both statements and expressions but only expressions
     have types
@@
 crashing program). We also discard submissions that are out of scope given
-PythonTutor's policies (e.g., very long-running executions or the
-use of forbidden libraries).
-
-Ultimately, the dataset used in this evaluation contained just over
-270,000 usable pairs of programs.
+PythonTutor's policies (e.g., very long-running executions or those that 
+use forbidden libraries).
+
+Ultimately, the dataset used in this evaluation contained 
+272,534 usable pairs of programs.
 
@@
 Our algorithm is based on supervised machine learning and thus requires
-labeled training instances. In addition, our evaluations require a ground
-truth notion of which terms correspond to correct fault localizations.
+labeled training instances.  For these experiments, we employ a random
+forest our machine learning algorithm, finding it to provide the best
+accuracy on our dataset with a scalable training and testing time.  In
+addition, our evaluations require a ground truth notion of which terms
+correspond to correct fault localizations.
 
@@
 those crashing programs for which the same user later submitted a program
-that did not crash. We compute a tree-diff \cite{tree-diff} between the
+that did not crash. We compute a tree-diff~\cite{tree-diff} between the
 originaly, buggy submission and the first fixed submission. We define
-the ground truth correct answer to be the set of terms in the crashing
-program that also appear in the diff. We discuss the implications of this
-choice in \secref{threats}.
+the \emph{ground truth} correct answer to be the set of terms in the
+crashing program that also appear in the diff. We discuss the implications
+of this choice in \secref{threats}.
+
+Given that notion of ground truth, a candidate fault localization answer is
+\emph{accurate} if it is contained in the ground truth set. That is, if the
+human user changed lines $X$ and $Y$, a technique (either our algorithm 
+or a baseline) is given credit for returning either $X$ or $Y$. When
+considering ranked lists, a response list is \emph{top-$k$ accurate} if any one
+of the top $k$ answers is contained in the ground truth set. \fixme{Read
+this paragraph and correct it for truth; it is currently simply WRW's best
+guess and I don't actually know what you did.}
 
@@
 the potential threat of overfitting.
-
-\subsection{Methodology}
-
-For these evaluations, we employ decision trees as our machine learning
-algorithm, since they provided the best accuracy on our data given the
-available training time.
-
-We report Top-1 and Top-3...
-
-We define FIXME-ACCURACY to be FIXME.
 
@@
 \caption{Fault localization accuracy.
-Baseline is the normal Python interpreter. Other bars represent
+Baseline is the normal Python interpreter. The Top-$k$ bars represent
 our approach, using random forests,
-on the largest dataset (272534 pairs).}
+on the largest dataset (272534 pairs). The strong performance of our
+algorithm on a large, real-world dataset is the primary result of this paper. 
+}
 \label{fig-full-dataset-acc-random-forest}
@@
 
-We first use the full dataset to train and test decsion forests optimized for
+We first use the full dataset to train and test decision forests optimized for
 Top-1, Top-2, and Top-3 accuracy. As shown in \figref{full-dataset-acc-random-forest},
-these produce a correct answer 60, 72, and
+these produce a correct answer 60\%, 72\%, and
 79\% of the time (to two significant figures). As a baseline, we determine that
-the expression blamed by the standard Python interpreter (i.e. the one that
-causes it to crash) is only changed by the user 45\% of the time. Thus our
-most directly comparable model, the one for Top-1, significantly outperforms
-this baseline, and as we then exercise our power to give a ranked output rather
-than a single answer, our margin gets even better.
+the expression blamed by the standard Python interpreter (i.e., the one that
+raises the uncaught exception) is only changed by the user 45\% of the
+time.
+
+Thus, our most directly comparable model (Top-1), significantly outperforms
+this baseline. Users who are only willing to look at a single error message
+would have been better-served by our Top-1 model on this historical data.
+
+In addition, previous studies have shown that developers are willing to use
+very short ranked lists~\cite[Sec.~5.6]{orso-parnin}, but not voluminous
+ones. Our Top-3 model accuracy of 79\% dramatically improves upon the
+current state of practice for scalable type-error localization in Python.
 
@@
 Name                      & Category                & Gini Importance \\ \bottomrule
-P size                    & Contextual (Syntactic)  & 0.112 \\
-C3 is statement           & Contextual (Syntactic)  & 0.061 \\
-P is list literal         & Contextual (Syntactic)  & 0.055 \\
+Parent size                    & Contextual (Syntactic)  & 0.112 \\
+Child3 is statement           & Contextual (Syntactic)  & 0.061 \\
+Parent is list literal         & Contextual (Syntactic)  & 0.055 \\
 Program crashes here      & Dynamic (Error location)& 0.039 \\
 Type is unknown           & Dynamic (Type)          & 0.037 \\
-P type is unknown         & Contextual (Type)       & 0.037 \\
+Parent type is unknown         & Contextual (Type)       & 0.037 \\
 Err message is IndexError & Dynamic (Error message) & 0.033 \\
 Size                      & Syntactic               & 0.028 \\
-P is dictionary literal   & Contextual (Syntactic)  & 0.027 \\
-C1 size                   & Contextual (Syntactic)  & 0.024 \\
+Parent is dictionary & Contextual (Syntactic)  & 0.027 \\
+Child1 size                   & Contextual (Syntactic)  & 0.024 \\
 Is variable               & Syntactic               & 0.020 \\
-C1 type is unknown        & Contextual (Type)       & 0.020 \\
+Child1 type is unknown        & Contextual (Type)       & 0.020 \\
 \toprule
@@
 \caption{Feature predictive power (for a Top-3 Decision Tree
-learned on the entire dataset). P and C1-C3 refer to the parent and
+learned on the entire dataset). Parent and Child1--3 refer to the parent and
 first three children of the node in question.}
@@
 Name                     & Category                & F-score \\ \bottomrule
-C3 is statement          & Contextual (Syntactic)  & 69 \\
+Child3 is statement          & Contextual (Syntactic)  & 69 \\
 Size                     & Syntactic               & 63 \\
 Program crashes here     & Dynamic (Error location)& 50 \\
-P type is list           & Contextual (Type)       & 36 \\
-C1 type is bool          & Contextual (Type)       & 30 \\
-P type is unknown        & Contextual (Type)       & 28 \\
-C1 is expression         & Contextual (Type)       & 25 \\
-C3 size                  & Contextual (Syntactic)  & 20 \\
-C1 is binary or unary op & Contextual (Syntactic)  & 20 \\
-C1 size                  & Contextual (Syntactic)  & 18 \\
-P type is dictionary     & Contextual (Type)       & 17 \\
+Parent type is list           & Contextual (Type)       & 36 \\
+Child1 type is bool          & Contextual (Type)       & 30 \\
+Parent type is unknown        & Contextual (Type)       & 28 \\
+Child1 is expression         & Contextual (Type)       & 25 \\
+Child3 size                  & Contextual (Syntactic)  & 20 \\
+Child1 is binary or unary op & Contextual (Syntactic)  & 20 \\
+Child1 size                  & Contextual (Syntactic)  & 18 \\
+Parent type is dictionary     & Contextual (Type)       & 17 \\
 Type is string           & Dynamic (Type)          & 12 \\
@@
 shown in
-thousands to two significant figures; all p-values are less than 0.01.}
+thousands to two significant figures; all p-values are less than 0.01.
+Parent and Child1--3 refer to the parent and first three children of
+the node in question.}
 \label{tab-anova}
@@
 \end{tikzpicture}
-\caption{Accuracy when features are removed, based on Top-1 decision trees on a
-random subset of 20000 program pairs, as compared to a Top-1 decision tree trained
-on all features.
+\caption{Normalized accuracy when categories of features are removed, based
+on Top-1 decision trees on a random subset of 20,000 program pairs, as
+compared to a Top-1 decision tree trained on all features.
 \fixme{will add No Syntax bar}
@@
 
+Having established the efficacy of our approach, we now investigate which
+elements of our algorithmic design (\secref{algorithm}) contributed to that
+success. 
 \tabref{feature-predictive-power} summarizes the relative importance
@@
 classification decisions is more important. Similarly, \tabref{anova} gives
-an alternate view of the relative feature importances, as measured by an
-analysis of variance (ANOVA).
-
-We also present the results of a leave-one-out analysis in which
-certain categories of features are removed and the model is trained
+an alternate view of the relative feature importances, as measured by a
+standard analysis of variance (ANOVA).
+
+Some features, such as ``program crashes here'' and features
+related to term size, are fairly direct. Others, such as ``Child3 is
+statement'', a contextual, syntactic feature that was found to be very
+important, merit additional explanation. In practice, we found it to
+encode FIXME, since Python terms with statements in the third child node
+position are FIXME. Similarly, features that determine if the parent node
+is a list or dictionary are relevant because of FIXME. In beginner-written
+Python we find that defects and typos often in the immediate child position
+of a list or dictionary term because of FIXME. Finally, we note the
+importance of our \lt{Unknown} type feature for expressions that are never
+evaluated in the given trace: we find that they are relevant because of
+FIXME. \fixme{Ben (etc.) must actually fill in real explanations here.}
+
+More broadly, we also present the results of a leave-one-out analysis in which
+entire categories of features are removed and the model is trained
 and tested only on those that remain. \figref{removing-features} shows
@@
 
-In all cases we see that syntactic, dynamic, and contextual features are
-all important. Features such as FIXME, FIXME and FIXME are found to be of
-primary importance. We interpret this to mean FIXME with respect to the
-application of our algorithm to beginner-written code.
+In all cases we see that syntactic, dynamic, and contextual features (i.e.,
+the design decisions of our algorithm) are crucial to our algorithm's
+accuracy.
 
@@
 
-We next try training and testing Top-1 decision trees on only the subsets of
-the data
-corresponding to each of the five most common exceptions: \lt{TypeError},
-\lt{IndexError}, \lt{AttributeError}, \lt{ValueError}, and \lt{KeyError}.
-As shown in \figref{defect-categories}, these models have accuracy between
-86\% and 115\% of a comparable model trained on the data as a whole. This shows
-that our algorithm is robust, able to give high accuracy fault localizations on
-a wide variety of defect types.
+We investigate the sensitivity of our algorithm to different categories of
+Python errors: is our accuracy sensitive to certain kinds of user defects? 
+We investigate training and testing Top-1 decision trees on only those
+subsets of the dataset corresponding to each of the five most common
+uncaught exceptions: \lt{TypeError}, \lt{IndexError}, \lt{AttributeError},
+\lt{ValueError}, and \lt{KeyError}. Together, these five exceptions make
+up FIXME\% of the errors in our dataset.
+
+As shown in \figref{defect-categories}, these per-defect models have
+normalized accuracy between 86\% and 115\% of a comparable model trained on
+the dataset as a whole. This shows that our algorithm is robust and able to
+give high-accuracy fault localizations on a wide variety of defect types. 
+Having consistent, rather than defect-type-sensitive, performance is
+important for debugging-tool
+usability~\cite{orso-parnin,Bessey2010,ayewah10}.
 
@@
 diverse than those used in previous work, we compare the diversity of
-programs present here to those in a relevant balance. The Nate
+programs used here to those in a relevant baseline. The Nate
 algorithm~\cite{learning-to-blame} also provides type error localization
@@
 also focused on beginner-written programs: specifically, it considered
-student submissions to 23 different university homework problems.  Thus the
-Nate dataset is comparatively homogeneous, raising concerns about whether
-evaluation results would generalize to more diverse settings.
+student submissions to 23 different university homework problems.  Such a
+dataset is comparatively homogeneous, raising concerns about whether
+associated evaluation results would generalize to more diverse settings.
 
@@
 approaches); instead, our claim is that even with simple clustering,
-it is clear that our dataset contains a greater diversity of programs than
-were considered by previously-published evaluations.
-We view it as an advantage of our algorithm that it can apply to multiple
-different program categories.
+it is clear that our dataset contains a greater diversity of programs, even
+when controlling for size, than were considered by previously-published
+evaluations. We view it as an advantage of our algorithm that it can apply
+to multiple different program categories.
 
@@
 \subsection{Qualitative Analysis}
-
-\fixme{
-Pick out one example where we did poorly. Walk through why it is ``out of
-scope'', in some sense: we would need features that capture XYZ to handle
-it correctly, but those are either expensive or undecideable. Hint at
-future work.
-}
 
@@
 
+We now highlight a few indicative localization examples in detail to
+demonstrate how our algorithm accurately localizes faults. 
+These examples are simplified slightly for presentation and to protect the
+anonymity of the programmers, but they retain the essential character of
+the originals from our dataset.
+
+\paragraph{NoneType}
+The function in \figref{win-example-one} attempts to remove all vowels from
+a given word. However the \lt{remove} method in line 7 actually modifies
+the list in place and returns \lt{None}. By contrast, the user-corrected
+version forgoes the assignment instead uses \lt{w_list.remove(letter)}. In
+the buggy case, Python does not crash until line 8, where its message is
+the somewhat-misleading \lt{TypeError: can only join an iterable}. However,
+our decision tree flags the correct statement, based on the features 
+that it is an assignment statement whose second child has type
+\lt{NoneType}. This seems to capture the intuitive reasoning that there is
+rarely a good reason to assign the value \lt{None} to a variable.
+
+Note that our algorithm uses both the syntactic form of the statement (an
+assignment) and the dynamic type of one of its children, so all our
+categories of features --- syntactic, dynamic, and contextual --- were
+useful here.
+
 \begin{figure}
@@
 \end{lstlisting}
-\caption{A program with an off-by-one bug.}
+\caption{A program with an off-by-one bug on line 2.}
 \label{fig-win-example-two}
 \end{figure}
-
-We now go through a couple examples where our method does well. These examples
-are adapted slightly for simplicity of
-presentation and to protect the anonymity of the programmers, but they retain
-the essential character of the originals from our dataset.
-
-\paragraph{NoneType}
-The function in \figref{win-example-one} attempts to remove all vowels from a
-given word. However the
-\lt{remove} method in line 7 actually modifies the list in place and returns
-\lt{None}, so the correct version forgoes the assignment and just has
-\lt{w_list.remove(letter)}. Python does not crash until line 8, where its
-message is \lt{TypeError: can only join an iterable}. However, our decision
-tree flags the correct statement, using as some of its rules that it is an
-assignment statement whose second child has type \lt{NoneType}. This seems to
-capture the intuitive rule that there is rarely a good reason to assign the
-value \lt{None} to a variable. Note that these rules use both the syntactic
-form of the statement (an assignment) and the dynamic type of one of its
-children, so all our categories of features - syntactic, dynamic, and
-contextual - were useful here.
 
@@
 3 during the final iteration of the for loop. The correct expression to blame
-is the addition \lt{len(areaCodes) + 1}. Some of the rules that our
+is the addition \lt{len(areaCodes) + 1}. Some of the features that our
 decision tree uses to correctly localize this bug include that the error is an
-index error, the type of the parent, and the fact that it's an addition
-expression. Thus once again we see all three categories of features being used.
+index error, the type of the parent, and the fact that it is an addition
+expression. This example also highlights the use of all three categories of
+features simultaneously to capture a notion of root cause that more closely
+aligns with human expectations.
 

### 2018-08-24 paper/main.tex
evaluation summary

@@
 \usepackage{url}
+\usepackage{framed}
 \usepackage{pgfplots}
@@
 \subsection{Labeled Training and Ground Truth}
+\label{sec-training}
 
@@
 
-\subsection{Evaluation Summary}
-
-Summarize the evaluation. One sentence per subsection.
-
 \subsection{Threats to validity}
@@
 
-\emph{Overfitting and other machine learning issues} Yao-Yuan, how do we avoid
-these?
-
-\emph{Language choice} We only showed that this technique works for Python 3.
-However, \cite{learning-to-blame} uses a similar technique to great effect in
-OCaml, and if it works on these two quite-different languages then we believe
-it would work on others as well. Additionally, Python is a common choice of
-language for introductory programming courses, so even a technique that worked
-only on Python would still be very useful for helping novices.
-
-\emph{Target population} We don't know exactly who our study population is -
-who wrote the programs in the dataset and what exactly they were trying to write.
-It is likely that they are primarily students and not professional software engineers,
-and thus possible that our technique does not generalize well to the kinds of
-bugs that would be found in an industrial setting rather than an educational one.
-
-\emph{Automatic labeling} Since our dataset is so large, we were forced to use an automatic
-scheme to label bugs rather than marking them by hand. We chose to use the diffs
-between each program that crashes and the next program that does not, but this method
-is not 100\% accurate because the next program may contain additional changes
-beyond what was needed to fix the bug, or it may indeed be an entirely unrelated
-program. We mitigate this problem by discarding as outliers program pairs that have
-changed too much. We also note that if our `true' accuracy is indeed a little
-smaller than our reported accuracy because of this effect, then the `true' baseline
-is also lower than the reported baseline for the same reason, so our
-\emph{margin} of effectiveness is unlikely to change much.
+Although our evaluation demonstrates that our algorithm scales to
+accurately localize Python errors in lage datasets, our results may not
+generalize to all use cases.
+
+\emph{Overfitting.} Since our algorithm makes use of supervised machine
+learning, one threat to validity is overfitting (i.e., learning a model
+that is too complex with respect to the data and thus fails to generalize).
+We mitigate this threat by using cross-validation, as discussed in
+\secref{training}.
+
+\emph{Language choice.} We have only demonstrated that our technique works
+for Python 3. We hypothesize that it should apply to similar
+dynamically-typed languages, such as Ruby, but such evaluations remain
+future work. We mitigate this threat slightly by constructing our algorithm 
+so that it does not depend on Python-specific features or node types
+(e.g., in \secref{slice-algorithm} we explicitly omit relatively ``exotic'' 
+features such as generators, ``yield'' and coroutines that may not always
+be present in other languages).
+
+\emph{Target population.} Unlike many other studies that focus specifically
+on students, we have very little information about the makeup of our
+subject population. The general popularity of PythonTutor is an advantage
+for collecting a large, indicative dataset, but it does mean that we have
+no specific information about the programmers or the programs they were
+trying to write. In general the website targets beginners rather than
+professional software engineers. As a result, our technique may not
+generalize well to the kinds of bugs that would be found in an industrial
+setting rather than a novice or educational one.
+
+\emph{Ground truth.} The size of our dataset precluded the manual annotation 
+of each buggy program. Instead, we used historical successful
+edits from actual users as our ground truth notion of the desired fault
+localization. This has the advantage of aligning our algorithm with user
+intuitions in cases where there are multiple logically-consistent answers
+(see \secref{motex}), and thus increasing the utility of our tool. However,
+this definition of ground truth may be overly permissive: the next
+correct program in the historical sequence may contain additional spurious
+changes beyond those strictly needed to fix the bug.
+We mitigate this problem by discarding as outliers program pairs that have
+changed too much. \fixme{Really? This is the first time WRW is hearing
+about this. If this is true, you have to go back and mention this earlier
+where we talk about ground truth definitions and data set selection.}
+
+\subsection{Evaluation Summary}
+
+We evaluated our results on over 270,000 pairs of beginner-written Python
+programs. Our primary research result relates to accuracy: 
+\begin{framed}
+\noindent Our algorithm is very accurate, implicating the correct program
+terms 60--79\% of the time (for Top-1 to Top-3 lists), compared to the
+Python interpreter's 45\% accuracy. 
+\end{framed} 
+In addition, we investigated the design decisions of our algorithm:
+\begin{framed}
+\noindent All feature categories (i.e., static, dynamic, and contextual)
+were critical, as measured by Gini importance, ANOVA F-score
+and a leave-one-out analysis. 
+\end{framed}
+Finally, we investigated the sensitivity of our algorithm to various
+contexts:
+\begin{framed}
+\noindent Our algorithm generalizes, with similar accuracy across the top
+error types and on a dataset that is more diverse than that of related
+work.
+\end{framed}
 

### 2018-08-24 paper/main.tex
rewrite related work

@@
 \label{fig-motex}
-A program with multiple fault localizations (lines 1 and 2).}
+A simple program with multiple fault localizations (lines 1 and 2).}
 \end{figure}
@@
 \section{Related Work}
-\fixme{This must all be rewritten or dropped, as it is directly copied from
-my previous papers. It serves as a placeholder to indicate length and to
-remind us of some Software Engineering bits we might cite.}
-
-There are several fault localization techniques that rely much more heavily
-on dynamic information\footnote{Note: we do not directly compare to these
-state-of-the-art fault localization techniques for several reasons.  First,
-the benchmarks used in this paper do not always have available the test
-suites needed by dynamic techniques and, conversely, those used in the
-previous work do not have the publicly available bug reports required
-by our technique.  Additionally, we return an answer only when confident in
-our result set and also employ a top list of all available results, while
-previous work is typically evaluated using a score metric.  We consider the
-techniques complimentary.}~\cite
-{harrold05,Renieris03,cleve05,wang09}.
-In general,
-these techniques leverage differences between passing and failing program
-executions.  While effective, this type of approach requires program traces
-for not only the specific fault at hand but also from a comprehensive regression
-test suite for comparison.
-A hybrid approach that considers both dynamic
-traces and static natural language similarity was recently proposed by
-Medini \textit{et al.}~\cite{Medini11}.
-In contrast to these approaches, our technique relies purely on static
-information that is readily available in most commercial systems with minimal
-additional developer effort.
-
-Prabhakararao and Ruthruff \emph{et al.} performed two human studies
-to gauge the effectiveness of an interactive fault localization tool
-developed for end users with little to no
-experience~\cite{Prabhakararao03,ruthruff05}.  The goal of their
-studies was to evaluate the use of feedback when locating faults and
-to generally study the process of fault localization, especially by
-users with no expert domain knowledge of the source.  By comparison,
-our human study also examines the fault localization process but for
-the purpose of evaluating software quality metrics.  We are less
-interested in the specific process and more concerned with the
-resulting accuracy and the human intuitions about the code in
-question.  Additionally, our human study is of a much broader scope and
-thus we hope it is more generalizable.
-%Additionally, their studies included 10 and 20
-%participants respectively using a spreadsheet-based visualization tool.
-%Our study of 61 humans looking at actual source code potentially provides
-%more generalizable results and targets specifically the area of software
-%maintenance completed by trained programmers.
-
-Ashok \textit{et al.} propose a similar natural language search technique
-in which users can match an incoming report to previous reports,
-programmers and source code~\cite{Ashok09}.  By comparison, our technique
-is more lightweight and focuses only on searching the code and the
-defect report.
-
-Jones \textit{et al.} developed Tarantula, a technique that performs
-fault localization based on the insight that statements executed often
-during failed test cases likely account for potential fault
-locations~\cite{harrold05}. Their approach is quite effective when
-a rich, indicative test suite is available and can be run as part of
-the fault localization process. It thus requires the fault-inducing
-input but not any natural language defect report. By contrast,
-our approach is lightweight, does not require an indicative test
-suite or fault-inducing input, but does require a natural language
-defect report. Both approaches will yield comparable performance, and
-could even be used in tandem.
-
-Cleve and Zeller localize faults by finding differences between
-correct and failing program execution states, limiting the scope of
-their search to only variables and values of interest to the fault in
-question~\cite{cleve05}. Notably, they focus on those variable and
-values that are relevant to the failure and to those program execution
-points where transitions occur and those variables become causes of
-failure. Their approach is in a strong sense finer-grained than ours:
-while nothing prevents our technique from being applied at the level
-of methods instead of files, their technique can give very precise
-information such as ``the transition to failure happened when $x$
-became 2.'' Our approach is lighter-weight and does not require
-that the program be run, but it does require defect reports.
-
-Renieris and Rice use a ``nearest neighbor'' technique in their
-Whither tool to identify faults based on exposing differences in
-faulty and non-faulty runs that take very similar executions
-paths~\cite{Renieris03}. They assume a large number of correct runs
-(e.g., normal test cases) and one failing run. Their approach uses a
-distance criterion to select the correct run that is closest to the
-failing run and produces a report of ``suspicious'' parts of the
-program. By comparison, we chose to limit the programmatic information
-used by our technique to only that which was reported by users: we
-do not use test case runs but do need natural language.
-
-Liblit \textit{et al.} use Cooperative Bug Isolation, a statistical
-approach to isolate multiple bugs within a program given a deployed user
-base. By analyzing large amounts of collected execution data from real
-users, they can successfully differentiate between different causes of
-faults in failing software~\cite{liblit05}. Their technique produces
-a ranked list of very specific fault localizations (e.g., ``the fault
-occurs when $i > arrayLen$ on line 57''). In general, their technique
-can produce more precise results than ours, but it requires a set of
-deployed users and works best on those bugs experienced by many users.
-By contrast, we do not require that the program be runnable, much less
-deployed, and use only natural language defect report text.
+
+Broadly, the two most relevant areas of related work are 
+software engineering approaches to fault localization (typically based on
+dynamic test information) and programming languages approaches to fault
+localization (typically based on static type information).
+
+A significant body of work in fault localization follows from the Tarantula
+project~\cite{tarantula}. Jones et al. proposed that statements executed
+often in failing test runs but rarely in successful test runs were likely
+to be implicated in the defect. Such spectrum-based approaches gather
+coverage or other dynamic information and rank statements by a
+mathematically-computed suspiciousness score. Subsequent algorithms
+such as Pinpoint~\cite{chen02} or Ochiai~\cite{Abreu2006} have improved
+upon Tarantula in various ways. For example, the Cooperative Bug Isolation
+approach gathers richer dynamic information (i.e., invariants) and also
+combines information from multiple end-user sources~\cite{liblit05}. 
+Spectrum-based fault localization has been
+the subject of significant empirical evaluation~\cite{Abreu2009}. 
+Research on spectrum-based methods continues to this day, with
+refinements, and even optimality results, to the associated
+mathematical formulae~\cite{yoo2017}. While human studies have demonstrated
+that users dislike large ranked lists~\cite{orso-parnin}, fault
+localization has only increased in relevance with the rise of automated
+program repair~\cite{genprog-tse-journal,semfix13}, where many techniques
+depend critically on accurate fault localization~\cite{gpfaultloc}.
+
+Our approach shares the overall goal of spectrum-based fault localization:
+pinpointing suspicious statements to reducing debugging effort. However,
+where spectrum-based traditionally focus on larger, industrial-scale 
+programs, we target beginner-written software. Spectrum methods require
+multiple test cases (ideally very many of them); we use just one program
+execution. Spectrum methods focus heavily on dynamic features; we make
+critical use of syntactic, contextual and type information as well. 
+In addition, some human studies have focused specifically on accuracy and
+expertise for fault localization~\cite{fry2010,Prabhakararao03,ruthruff05};
+our decision to use particular features to support beginning programmers is
+informed by such insights.
+
+Zeller's popular Delta Debugging algorithm~\cite{zeller99}, interpreted
+generally, efficiently finds a minimal ``interesting'' subset from among a
+large set of elements. When the set of elements represents changes made to
+a source code version control system and interesting is defined with
+respect to failing a test suite, it can quickly locate program edits that
+cause regressions. Alternatively, when considering correct and failing
+execution states, such approaches can help focus on variable values that
+cause failures~\cite{zeller05}. While our dynamic slicing information can
+be viewed as a coarse approximation to the precise, fine-grained
+localization such an approach can provide, a key difference is our use
+of machine learning to agree with human judgments in cases where multiple
+causes are equally logically valid.
+
+In the programming languages community, a large body of work has focused on 
+localizing faults and providing better error messages, typically through
+the use of type information. \fixme{Sherrloc, Mycroft, Nate, etc.}
 

### 2018-08-24 paper/main.tex
abstract, conclusion, etc.

@@
 \begin{abstract}
-Abstract
+As dynamically-typed languages continue to grow in popularity, especially
+among beginning programmers, there is an increased need to provide precise
+localization for defects. In many instances, such localization is
+ambiguous: multiple program locations are equally formally implicated in
+the bug, but not all are equally relevant to human programmers. We propose
+a scalable fault localization approach for dynamic languages that agrees
+with human judgments. We base our approach on a combination of static,
+dynamic, and contextual features, guided by machine learning. We evaluate
+our approach on over 270,000 diverse real user instances from the popular
+PythonTutor.com website. We find that our approach is scalable, general,
+and quite accurate: up to 79\% for top-three responses, compared to 45\%
+for the default interpreter. 
 \end{abstract}
@@
 
+We present an approach for accurately localizing type errors in
+off-the-shelf beginner-written Python programs. Our approach makes use of a
+combination of static, dynamic and contextual features. Static features,
+such as syntactic forms and expression sizes, are a particularly powerful
+heuristic for novice programmer defects. Dynamic features can both 
+implicate relevant terms (partial typing information) and rule out
+irrelevant program regions (slicing). Contextual features allow our
+approach to gain the benefits of precise AST- or CFG-style information
+while retaining the scalable performance of a linear representation. 
+We use off-the-shelf machine learning to accurately combine those disparate
+features in a way that captures and models human judgments of ground-truth
+correct answers --- a notion that is especially relevant when multiple 
+program locations are equally formally implicated but not equally useful to
+the user.
+
+We evaluate our approach with respect to historical user defects and fixes.
+All of the feature categories (static, dynamic, and contextual) were
+relevant to success, as measured by multiple analyses (Gini, ANOVA and
+leave-one-out). Our evaluation demonstrated significant scalability and
+generality: our 270,000 instances were both an order-of-magnitude more
+numerous than similar related work and measurably more diverse. Ultimately,
+our approach was quite accurate, implicating the correct program location
+60--79\% of the time compared to the Python interpreter's 45\% accuracy.
+
 \bibliographystyle{abbrv}

### 2018-08-24 paper/main.tex
graph; citations

@@
 localization either require rich static types (e.g., as in
-Mycroft~\cite{FIXME}, Nate~\cite{learning-to-blame} or Sherrloc~\cite{FIXME}), require
+Mycroft~\cite{MYCROFT-type-errors}, Nate~\cite{learning-to-blame} or Sherrloc~\cite{SHErrLoc}), require
 test cases (e.g., as in Tarantula~\cite{tarantula}), or produce voluminous
@@
     ybar,
-    symbolic x coords={No Context, No Types},
+    symbolic x coords={No Context, No Syntax, No Types},
     xtick=data,
@@
     ylabel=Normalized Model Accuracy,
-    enlarge x limits=0.5,
+    enlarge x limits=0.3,
     legend style={at={(0.5,-0.15)},anchor=north}
@@
 compared to a Top-1 decision tree trained on all features.
-\fixme{will add No Syntax bar}
 }

### 2018-08-24 paper/main.tex
minor fixes

@@
 Let the number of examples be $N$.
-Given a data set $\mathcal{D} \in \{\mathbf{x}^{(i)}, y^{(i)}\}^{N}_{i=1}$,
+Given a dataset $\mathcal{D} \in \{\mathbf{x}^{(i)}, y^{(i)}\}^{N}_{i=1}$,
 where $\mathbf{x}^{(i)} \in \mathcal{R}^{M}$ and $y^{(i)} \in \{0, 1\}$,
@@
 Due to the large amount of data provided in this task, we only consider the
-following models which are scalable to large data set.
+following models which are scalable to large dataset.
 
@@
 At this point we have feature vectors describing every term of every program in
-our data set. The vectors from a random 30\% of the program pairs are set aside for testing and
+our dataset. The vectors from a random 30\% of the program pairs are set aside for testing and
 the other 70\% are used for training.
@@
 
-In our evaluation, we used FIXME as a training set and FIXME as a held-out
+In our evaluation, we used 70\% as a training set and 30\% as a held-out
 evaluation set. We employed cross-validation~\cite{kohavi} to help address
@@
 about this. If this is true, you have to go back and mention this earlier
-where we talk about ground truth definitions and data set selection.}
+where we talk about ground truth definitions and dataset selection.}
 

### 2018-08-24 paper/main.tex
related work

@@
 Our approach shares the overall goal of spectrum-based fault localization:
-pinpointing suspicious statements to reducing debugging effort. However,
+pinpointing suspicious statements to reduce debugging effort. However,
 where spectrum-based traditionally focus on larger, industrial-scale
@@
 localizing faults and providing better error messages, typically through
-the use of type information. \fixme{Sherrloc, Mycroft, Nate, etc.}
+the use of type information. SHErrLoc~\cite{SHErrLoc} identifies a set of
+incompatible type constraints, then chooses the likely type error using the
+Bayesian heuristic that the program is probably mostly correct. We rely on the
+complementary heuristic that faults will share characteristics
+with previously recorded faults, and apply this heuristic to features beyond
+just type information. Mycroft~\cite{MYCROFT-type-errors} modifies existing
+type inference algorithms to produce a ``correcting set'' for a program with
+a type error. Mycroft assumes that the minimal such set is the most
+desirable to the user, and has no way to rank multiple equally
+small sets. We instead use machine learning to agree with human judgements.
+NATE~\cite{learning-to-blame} uses machine learning on static and contextual
+features, including static types, to localize type errors in OCaml
+code. We employ a similar approach to localize Python faults, but we use
+dynamic features as well as static and contextual ones, we handle a wide
+variety of errors beyond type errors, and we evaluate our approach on a set of
+programs both more heterogeneous and almost two orders of magnitude larger than
+the NATE dataset.
 

### 2018-08-24 paper/main.tex
it seems we didn't put LR and MLP in final result

@@
 
-\fixme{how the data is represented}
-
-\fixme{is program pair the correct term?}
-
 To apply machine learning to analyze the data, we have to formulate the
@@
 \textsc{scikit-learn}~\cite{scikit-learn}.
-% Due to the large amount of data points, we did not use \textsc{scikit-learn}'s
-% implementation for logistic regression and multilayer perceptron.
-% We implementation both model using
-% \textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.
+%Due to the size of the dataset, we did not use \textsc{scikit-learn}'s
+%implementation for logistic regression and MLP.
+%We implementation both model using
+%\textsc{keras}~\cite{chollet2015keras} with mini-batch gradient descent.
 
@@
 $10^{-3}$ with multiple of $10^{-2}$ and the minimum samples in each leaf is set to $200$.
-For logistic regression and MLP, binary cross-entropy is set as the loss
-function and Adam \cite{kingma2014adam} optimizer is used.
-Learning rate is set to $0.1$ and the number of epochs are $10$.
-L2 regularization is applied to the weights of both models,
-the weight for L2 term is searched from $10^0$ to $10^{-5}$ with a multiple of $10^{-1}$.
+%For logistic regression and MLP, binary cross-entropy is set as the loss
+%function and Adam \cite{kingma2014adam} optimizer is used.
+%Learning rate is set to $0.1$ and the number of epochs are $10$.
+%L2 regularization is applied to the weights of both models,
+%the weight for L2 term is searched from $10^0$ to $10^{-5}$ with a multiple of $10^{-1}$.
 Other parameters that are not mentioned are kept with the default setting of the

### 2018-08-24 paper/main.tex
paper

@@
 dynamically-typed programs, is that there are often multiple
-logically-valid ways to address a type error~\cite{FIXME} --- only one of
-which may be desired by the programmer~\cite{FIXME}.  Consider the program
+logically-valid ways to address a type error --- only one of
+which may be desired by the programmer~\cite{learning-to-blame}.  Consider the program
 in \figref{motex}, adapted from our dataset of human submissions. The
@@
 crashes (throws an uncaught Python exception) is paired with the next
-program (by the same user) that does not crash, under the assumption that
+program by the same user that does not crash, under the assumption that
 the latter is the fixed version of the former. We discard pairs where the
 difference between crashing and fixed versions is too high (more than a
-stddev above average), since these are most likely to be violations of that
+standard deviation above average), since these are most likely to be violations of that
 assumption (i.e., the program that does not crash is unrelated to the
@@
 labeled training instances.  For these experiments, we employ a random
-forest our machine learning algorithm, finding it to provide the best
+forest as our machine learning algorithm, finding it to provide the best
 accuracy on our dataset with a scalable training and testing time.  In
@@
 
-We use the terms changed involved in the fixes by actual users as our
+We use the terms changed in fixes by actual users as our
 ground truth and training information.  Many PythonTutor users use the site
@@
 that did not crash. We compute a tree-diff~\cite{tree-diff} between the
-originaly, buggy submission and the first fixed submission. We define
-the \emph{ground truth} correct answer to be the set of terms in the
+original, buggy submission and the first fixed submission. For example,
+if the expression \lt{len(\{3,4\})} is changed to \lt{len([3,4])}, then the
+node in the tree corresponding to the set \lt{\{3,4\}} as a whole will
+appear in the diff (since it has been changed to a list), but neither its parent
+node (the enclosing \lt{len} call) nor its children nodes (the integer literals)
+will appear since they were not changed. Changing the number of children of a
+node also causes the node to appear in the diff, e.g. adding a statement
+to a while loop counts as a change to the loop as a whole.
+
+We define the \emph{ground truth} correct answer to be the set of terms in the
 crashing program that also appear in the diff. We discuss the implications
-of this choice in \secref{threats}.
+of these choices in \secref{threats}.
 
@@
 \emph{accurate} if it is contained in the ground truth set. That is, if the
-human user changed lines $X$ and $Y$, a technique (either our algorithm
+human user changed terms $X$ and $Y$, a technique (either our algorithm
 or a baseline) is given credit for returning either $X$ or $Y$. When
 considering ranked lists, a response list is \emph{top-$k$ accurate} if any one
-of the top $k$ answers is contained in the ground truth set. \fixme{Read
-this paragraph and correct it for truth; it is currently simply WRW's best
-guess and I don't actually know what you did.}
+of the top $k$ answers is contained in the ground truth set.
 
@@
 important, merit additional explanation. In practice, we found it to
-encode FIXME, since Python terms with statements in the third child node
-position are FIXME. Similarly, features that determine if the parent node
+encode complex control structures, like large loops, where fixing the crash
+often involved adding or removing whole statements from the loop.
+Similarly, features that determine if the parent node
 is a list or dictionary are relevant because of FIXME. In beginner-written

### 2018-08-25 paper/main.tex
edited

--- 2018-08-24 paper/main.tex
@@
 problem into standard classification problem.
-For each line of a program pair, we extracted the features mentioned in
+The corresponding label, is represented as $1$ and $0$, which means whether
+the term of a program pair (a buggy and a fixed program) is blamed or not.
+This is a binary classification problem.
+For each term of a program pair, we extracted the features mentioned in
 Section \ref{sec-features}.
@@
 \mathcal{R}^{M}$.
-For the corresponding label, is represented as $1$ and $0$, which means whether
-the line blamed or not.
+
+The notations are defined as follows.
 Let the number of examples be $N$.
@@
 Decision tree~(DT) learns a tree of binary predicates over the features and
-recursively partitioning the feature space until a final classification can
-be made.
-%
-Each node in the tree contains a single thresholding $\mathbf{x}_i \leq t$, where
-applying threshold $t \in \mathcal{R}$ on the $i$-th feature of $\mathbf{x}$,
-which determines whether a given input should proceed down the left or right
-subtree.
+recursively partitioning the feature space until specific criteria is met.
+The usual stopping criteria is the number of data in a node falls below a
+certain threshold.
+%
+Each node in the tree contains a single thresholding binary classifier
+$\[\mathbf{x}_i \leq t\]$, $\[\cdot\]$ is the indicator function.
+$i$ is greedly selected to minimize the impurity of the training data.
+The impurity is usually measured by gini index
+\cite{breiman2017classification} or entropy \cite{quinlan1986induction}.
+This thresholding classifier determines that during prediction, whether a
+given input should proceed down the left or right subtree.
 %
@@
 %
-The training of a decision tree is usually done in a top-down greedy manner,
-and there are several implementations of DT such as C4.5
+There are several implementations of DT such as C4.5
 \cite{quinlan2014c4} and CART \cite{breiman2017classification}.
@@
 %
-Random forest (RF)~\cite{breiman2001random} improves generalization by
-learning an \emph{ensemble} of decision trees.
+The greedy nature of DT make it often suffers from overfitting and may
+require a lot of tuning such as pruning the tree.
+Random forest (RF)~\cite{breiman2001random} improves DT by learning an
+\emph{ensemble} of decision trees.
+During training, RF randomly selects a subsets of training data by sampling
+with replacement and use different subset of data to train multiple different
+of trees.
 For prediction, each tree in the forest votes on one of the classes.
 The percentage of vote for each class forms a natural confidence score.
-During training, RF selects a subset of training data by sampling with
-replacement and use the subset to train each of the tree.
-By doing so, RF is able to generate a forest with each tree trained on
-a different subset of data.
+
+By randomly sampling, RF is able to generate a forest with diverse trees
+trained on a different subset of data.
 The diversity of the underlying models tends to make RF less susceptible to
-the overfitting, but it also makes the learned model more difficult to
-interpret.
+the overfitting.
+However, the trade-off is the use more computational power/
+% but it also makes the learned model more difficult to
+%interpret.
 %
@@
 At this point we have feature vectors describing every term of every program in
-our dataset. The vectors from a random 30\% of the program pairs are set aside for testing and
-the other 70\% are used for training.
-Within the training data, we use 3-fold
-cross-validation to select the best parameters for our models, before training
-using those parameters on the full training set and then scoring the models
-using the testing set.
-Each trained model takes in any vector representing a term
-in a buggy program, and returns a confidence score representing how likely it is that
-that term was one of the terms changed between the fixed and buggy versions.
+our dataset.
+We first split the the data by 70-30 into training data and testing data.
+Within the training data, we use 3-fold cross-validation~\cite{kohavi} to
+select the best parameters for our models.
+Then we use the whole training data to train the model with best parameter
+and report its performance on the testing data.
+
+Each trained model takes in the feature vector $\mathbf{x}$ and returns a
+confidence score representing how likely it is that the term was one of the
+terms changed between the fixed and buggy programs.
 Thus we can treat the model as providing a \emph{ranking} over all terms by
@@
 changed) present in the top $k$ results.
-We retrain the models for each value of k, since a model optimized to produce
+We retrain the models for each value of $k$, since a model optimized to produce
 the single best result might be different from one optimized to get a good

### 2018-08-25 paper/main.tex
fix

@@
 Each node in the tree contains a single thresholding binary classifier
-$\[\mathbf{x}_i \leq t\]$, $\[\cdot\]$ is the indicator function.
+$[ \mathbf{x}_i \leq t ]$, $[ \cdot ]$ is the indicator function.
 $i$ is greedly selected to minimize the impurity of the training data.

### 2018-08-25 paper/main.tex
paper

@@
 often involved adding or removing whole statements from the loop.
-Similarly, features that determine if the parent node
-is a list or dictionary are relevant because of FIXME. In beginner-written
-Python we find that defects and typos often in the immediate child position
-of a list or dictionary term because of FIXME. Finally, we note the
+In contrast, features that determine if the parent node
+is a list or dictionary are relevant as strong negative features: we find that
+fixing bugs in beginner-written
+Python rarely requires changing the elements of a list or dictionary literal.
+Finally, we note the
 importance of our \lt{Unknown} type feature for expressions that are never
@@
     }
-\legend{NATE OCaml data, Our Python data}
+\legend{Nate OCaml data, Our Python data}
 \end{axis}
 \end{tikzpicture}
-\caption{Number of natural program clusters in each dataset as a function
+\caption{Number of natural program clusters in the Nate dataset and in a
+random sample of our dataset of the same size, as a function
 of clustering parameter (informally, how ``strict'' or ``tight'' the
-clustering is). \fixme{Talk about how you sample to be fair to the smaller
-datasets.} At all points our dataset demonstrates more diversity than
+clustering is). At all points our dataset demonstrates more diversity than
 that of previous work.}
@@
 % \figref{diversity}, regardless of the choice of threshold, our data breaks into
-% more clusters than the NATE data. This confirms that our data is more
+% more clusters than the Nate data. This confirms that our data is more
 % heterogeneous.
@@
 
-\fixme{We need to say something about how we controlled for number, lest
-reviewers think we have more clusters simply because we have hundreds of
-thousands of datapoints.}
-
 \figref{diversity} shows the results of clustering on our dataset and the
-Nate dataset. At a high level, the structural similarity in the graphs
+Nate dataset. Since our dataset is much larger than Nate's, we
+used a random sample of our data of the same size as the Nate dataset.
+At a high level, the structural similarity in the graphs
 confirms the comparability of cluster counts between the two datasets.  For
@@
 small sets. We instead use machine learning to agree with human judgements.
-NATE~\cite{learning-to-blame} uses machine learning on static and contextual
+Nate~\cite{learning-to-blame} uses machine learning on static and contextual
 features, including static types, to localize type errors in OCaml
@@
 programs both more heterogeneous and almost two orders of magnitude larger than
-the NATE dataset.
+the Nate dataset.
 

### 2018-08-25 paper/main.tex
paper

@@
 Nate~\cite{learning-to-blame} and the natural language processing term
-frequency vector (or ``bag of words'') model~\cite{FIXME}, we represent
+frequency vector (or ``bag of words'') model~\cite{Salton:1986:IMI:576628}, we represent
 each buggy program as a ``bag of abstracted terms''. A \emph{term} is
@@
 \lt{ValueError}, and \lt{KeyError}. Together, these five exceptions make
-up FIXME\% of the errors in our dataset.
+up almost 96\% of the errors in our dataset.
 
@@
 the dataset as a whole. This shows that our algorithm is robust and able to
-give high-accuracy fault localizations on a wide variety of defect types.
+give high-accuracy fault localizations on a variety of defect types.
 Having consistent, rather than defect-type-sensitive, performance is
@@
 code. We employ a similar approach to localize Python faults, but we use
-dynamic features as well as static and contextual ones, we handle a wide
+dynamic features as well as static and contextual ones, we handle a
 variety of errors beyond type errors, and we evaluate our approach on a set of

### 2018-08-25 paper/main.tex
minor changes

@@
 
-To apply machine learning to analyze the data, we have to formulate the
+To apply machine learning for analysis of the data, we have to formulate the
 problem into standard classification problem.
@@
 
-The notations are defined as follows.
+The binary classification problem is defined as follows.
 Let the number of examples be $N$.
@@
 %
-There are several implementations of DT such as C4.5
-\cite{quinlan2014c4} and CART \cite{breiman2017classification}.
+There are several implementations of DT such as C4.5 \cite{quinlan2014c4},
+ID3 \cite{quinlan1986induction} and CART \cite{breiman2017classification}.
 
